---
title: "solar_analysis_filtering"
format: html
editor: visual
---

# SOLAR Results Analysis

After obtaining the IP-Protein associations from the SOLAR analysis. Exploration is needed for interpretation.

## 1. Set up

### 1.1 Load libraries

```{r}
library(stringr)
library(dplyr)
library(tidyr)
library(viridis) #Colour blind friendly palette
library(ggplot2) #Graphing
library(pheatmap) #Heatmap

#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
```

### 1.2 Set Working Directory

```{r}
setwd("/users/k25046756/Twins_Project/data/solar_analysis")
```

### 1.3 Loading Data

```{r}
#Importing the summary results
summary_results<- read.delim("summary_heritability.csv")
```

## 2. Filtering of Significant Associations

### 2.1 FDR Adjusted Filtering

Adjusting with FDR to find those significant based on FDR \< 0.05

```{r}
#Adding in FDR of rhoG and rhoE
summary_results_fdr <- summary_results %>%
  mutate(
    rhoG_FDR = p.adjust(rhoG_p0, method = "fdr"),
    rhoE_FDR = p.adjust(rhoE_p, method = "fdr")
  )

#Genetics
#Experimenting with different h2r thresholds
genetic_0.1_hits <- summary_results_fdr %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05)
#4300

genetic_0.3_hits <- summary_results_fdr %>%
  filter(h2r1 > 0.3 & h2r2 > 0.3 & rhoG_FDR < 0.05)
#4047

genetic_0.5_hits <- summary_results_fdr %>%
  filter(h2r1 > 0.5 & h2r2 > 0.5 & rhoG_FDR < 0.05)
#1068


#Environmental
#Only accounting for rhoE as h2r can be low or high 
environmental_hits <- summary_results_fdr %>%
  filter(rhoE_FDR < 0.05)


#Significant in both genetic and environmental correlations
both_hits <- summary_results_fdr %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05 & rhoE_FDR < 0.05)
```

### 2.2 Further Filtering

Further filtering of genetically correlated pairs with environmental associations and vise versa - environmentally correlated pairs with genetic associations.

```{r}
#Genetics -> how many are also environmental 
gen_env_0.1 <- genetic_0.1_hits %>%
  filter(rhoE_p < 0.05)

gen_env_0.3 <- genetic_0.3_hits %>%
  filter(rhoE_p < 0.05)
#same 130 pairs 

gen_env_0.5 <- genetic_0.5_hits %>%
  filter(rhoE_p < 0.05)
#only 1 IMMUNO_41521 TNFB

#Environmental -> how many are genetics
env_gen <- environmental_hits%>%
  filter(rhoG_p0<0.05)
```

### 2.3 Saving the output environment

Saving to be loaded downstream

```{r}
save.image(file = "solar_analysis_filtering_results.RData")
```

## 3. Protein frequencies

In the 0.1 and 0.3 top hits, find what proteins are most frequent/popular for genetic. Observing whether there are changes in protein frequency patterns.

### 3.1 Genetic h2r 0.1 filtered hits

```{r}
#Count protein frequency in genetic_0.1_hits
protein_frequency_0.1 <- genetic_0.1_hits %>%
  group_by(trait2) %>% #trait2 is the protein column
  summarise(count = n()) %>%
  arrange(desc(count))

#Looking at the top 10
print(head(protein_frequency_0.1, 10))
```

### 3.2 Genetic h2r 0.3 filtered hits

```{r}
#Count protein frequency in genetic_0.3_hits
protein_frequency_0.3 <- genetic_0.3_hits %>%
  group_by(trait2) %>% 
  summarise(count = n()) %>%
  arrange(desc(count))

print(head(protein_frequency_0.3, 10))
```

### 3.3 Environmentally filtered hits

```{r}
# Count how many times each protein appears in your genetic hits
protein_frequency_env <- environmental_hits %>%
  group_by(trait2) %>% # assuming trait2 is your protein
  summarise(count = n()) %>%
  arrange(desc(count))

# Look at the top 10
print(head(protein_frequency_env, 10))
```

## 4. IP-Protein Summary Statistics

Show the distribution of those genetically, environmentally and both significant.

### 4.1 Graphing Correlation Landscape

Visualising the distribution of genetic and environmental correlations.

```{r}
#From summary_results_fdr, seleting those significant for rhoG, rhoE and both
#For purpose of graphing, we redefine the labels
summary_results_fdr_points <- summary_results_fdr %>%
 mutate(Sig_Type = case_when(
  rhoG_FDR < 0.05 & rhoE_FDR < 0.05 ~ "Both Significant",
  rhoG_FDR < 0.05 & rhoE_FDR >= 0.05 ~ "Genetic Only (rhoG)",
  rhoG_FDR >= 0.05 & rhoE_FDR < 0.05 ~ "Environmental Only (rhoE)",
  TRUE ~ "Non-Significant"
 ))

#Plotting a graph to show the rhoG and rhoE distribution of all IP -Protein associations
ggplot(summary_results_fdr_points, aes(x = rhoG, y = rhoE, color = Sig_Type)) +
 #Adding each pair as a point
 geom_point(alpha = 0.5, size = 1) +
 #Adding a cross
 geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
 geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
 #Colour scale
 scale_color_viridis_d(option = "viridis") +
 theme_minimal() +
 #Lablelling
 labs(
  title = "IP-Protein Association Correlation Landscape",
  subtitle = "Based on FDR-adjusted significance",
  x = "Genetic Correlation (rhoG)",
  y = "Environmental Correlation (rhoE)",
  color = "Significance Status"
 )
```

### 4.2 Summary Table

Summary table from the fdr results based on correlation types for quantification.

```{r}
summary_table <- summary_results_fdr %>%
  group_by(Sig_Type) %>%
  summarise(
    n = n(),
    mean_rhoG = mean(rhoG),
    mean_rhoE = mean(rhoE),
    mean_h2r1 = mean(h2r1),
    mean_h2r2 = mean(h2r2)
  )

summary_table
```

## 5. Transforming the data frame

For visualisations, the dataframe, which is in a long format, needs to be converted into a wide format (numerical matrix). The dataframes are normalised and NAs were already filtered.

```{r}
#Create a function that applies to all filtered dataset variables
conv_matrix <- function(df, value_col) {
  
  mat <- df %>%
    select(trait1, trait2, all_of(value_col)) %>% 
    
    pivot_wider(
      names_from = trait2,   #changing proteins column
      values_from = all_of(value_col)
    ) %>%

    column_to_rownames("trait1") %>%
    as.matrix()
  
  #replacing NAs from matrix conversion with 0
  mat[is.na(mat)] <- 0
  mat[is.nan(mat)] <- 0
  mat[is.infinite(mat)] <- 0
  
  return(mat)
}
   
```

## 6. Genetic Hits Analysis

With the genetic hits filtered for h2r \> 0.1.

### 6.1 Heatmap

```{r}
genetic_matrix <- conv_matrix(genetic_0.1_hits, "rhoG")

genetic_heatmap<-pheatmap(
            t(genetic_matrix), #transpose so x axis is IP and protein on Y 
            color = viridis(100),
            cluster_rows = TRUE,
            cluster_cols = TRUE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = TRUE,
            show_colnames = FALSE,
            na_colour = "black",  
            main = "Genetic Correlations (IP vs Proteins)",
            border_color = NA
          )
genetic_heatmap
```

Heatmap appears sparce visually.

### 6.2 Shared Protein Distribution

```{r}
#Counting how many share proteins
ip_gen_sharing <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_gen <- table(ip_gen_sharing$protein_count)

shared_protein_ip_gen
```

## 7. Environmental Hits Analysis

### 7.1 Heatmap

```{r}
environmental_matrix <- conv_matrix (environmental_hits, "rhoE")

env_heatmap<-pheatmap(
            environmental_matrix,
            color = viridis(100),
            cluster_rows = TRUE,
            cluster_cols = TRUE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = FALSE,
            show_colnames = TRUE,
            na_colour = "black",  
            main = "Environmental Correlations (IP vs Proteins)",
            border_color = NA
          )
env_heatmap
```

### 7.2 Shared Protein Distribution

```{r}
#Counting how many share proteins
ip_env_sharing <- environmental_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_env <- table(ip_env_sharing$protein_count)

shared_protein_ip_env
```

## 8. Both Genetic and Environmentally Correlated

All hits are positive. Only CXCL9 protein has IP associations.

```{r}
#Plotting with different graph
both_dual <- both_hits %>%
  mutate(IP = reorder(trait1, abs(rhoG))) %>%
  select(IP, Protein = trait2, rhoG, rhoE)

#Lollipop graph to show both rhoG and rhoE
ggplot(both_dual, aes(x = reorder(IP, abs(rhoG)))) +
  geom_segment(aes(xend = IP, y = rhoG, yend = rhoE), color = "grey85", size = 1) +
  geom_point(aes(y = rhoG, color = "Genetic (ρG)"), size = 3) +
  geom_point(aes(y = rhoE, color = "Environmental (ρE)"), size = 3) +
  coord_flip() +
  scale_color_viridis_d(option = "D", begin = 0.2, end = 0.8)+
  labs(
    title = "Genetic + Environmental Correlations",
    subtitle = "Only CXCL9 had IPs significant in both genetic and environmental correlations.",
    x = "Immunophenotype", y = "Correlation Coefficient with CXCL9",
    color = "Correlation Type"
  ) +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5)
```

Higher genetic correlations than environmental counterpart.
