---
title: "solar_analysis_filtering"
format: html
editor: visual
---a
---

# SOLAR Results Analysis

After obtaining the IP-Protein associations from the SOLAR analysis. Exploration and filtering is needed for interpretation.

## 1. Set up

### 1.1 Load libraries

```{r, message=FALSE}
library(stringr)
library(dplyr)
library(tidyr)
library(tibble)

library(viridis) #Colour blind friendly palette
library(ggplot2) #Graphing
library(pheatmap) #Heatmap
```

### 1.2 Loading Data

```{r}
#Importing the summary results
summary_results<- read.delim("~/Twins_Project/data/solar_analysis/summary_heritability.csv")

#Loading in the dataset for sup.lineage information 
load("/users/k25046756/Twins_Project/data/raw_olink_ip/results/Olink_associations.RData")
```

## 2. Filtering of Significant IP-Protein Associations

### 2.1 FDR Adjusted Filtering

Adjusting with FDR to find those significant based on FDR \< 0.05.

```{r}
#Adding in FDR of rhoG and rhoE
summary_results_fdr <- summary_results %>%
  mutate(
    rhoG_FDR = p.adjust(rhoG_p0, method = "fdr"),
    rhoE_FDR = p.adjust(rhoE_p, method = "fdr")
  )

```

### 2.2 Adding in Lineages from infoF dataframe

Using Sup.Lineage as it contains T cells, NK cells, monocytes, dendritic cells, B cells to determine the IP lineages.

```{r}
#Need to clean the IMMUNO_X.new names to match summary_result_fdr as it doesnt have .new 
infoF_clean <- infoF %>%
  mutate(SET.NAME_clean = gsub("\\.new$", "", SET.NAME)) %>%
  distinct(SET.NAME_clean, .keep_all = TRUE)

#Combining the lineage columns to the summary results dataframe
sum_results_combined <- summary_results_fdr %>%
  left_join(infoF_clean %>% select(SET.NAME_clean, Sup.Lineage), 
            by = c("trait1" = "SET.NAME_clean"))
```

### 2.3 Filtering based on Genetic, Environment or both associations based on the SOLAR outputs

Using the heritability h2r results with rhoG, we can determine genetic hits. For environmental hits, h2r does not need to be significant.

#### 2.3.1 Genetic Hits

```{r}
#Experimenting with different h2r thresholds
genetic_0.1_hits <- sum_results_combined %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05)
#4300

genetic_0.3_hits <- sum_results_combined %>%
  filter(h2r1 > 0.3 & h2r2 > 0.3 & rhoG_FDR < 0.05)
#4047

genetic_0.5_hits <- sum_results_combined %>%
  filter(h2r1 > 0.5 & h2r2 > 0.5 & rhoG_FDR < 0.05)
#1068

```

#### 2.3.2 Environmental hits

```{r}
#Only accounting for rhoE as h2r can be low or high 
environmental_hits <- sum_results_combined %>%
  filter(rhoE_FDR < 0.05)
```

#### 2.3.3 Both hits

```{r}
#Significant in both genetic and environmental correlations
both_hits <- sum_results_combined %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05 & rhoE_FDR < 0.05)
```

### 2.2 Further Filtering

Further filtering of genetically correlated pairs with environmental associations and vise versa - environmentally correlated pairs with genetic associations.

#### 2.2.1 Genetics -\> Environmental

```{r}
#Genetics -> how many are also environmental 
gen_env_0.1 <- genetic_0.1_hits %>%
  filter(rhoE_p < 0.05)

gen_env_0.3 <- genetic_0.3_hits %>%
  filter(rhoE_p < 0.05)
#same 130 pairs 

gen_env_0.5 <- genetic_0.5_hits %>%
  filter(rhoE_p < 0.05)
#only 1 IMMUNO_41521 TNFB
```

#### 2.2.2 Environmental -\> Genetics

```{r}
#Environmental -> how many are genetics
env_gen <- environmental_hits%>%
  filter(rhoG_p0<0.05)
```

### 2.3 Saving the output environment

Saving to be loaded elsewhere.

```{r}
save.image(file = "solar_analysis_filtering_results.RData")
```

## 3. Exploration of the filtered results

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/results/visualisations"
```

### 3.1 Protein Distribution

Analysing how many proteins per IPs are present in each filtering.

#### 3.1.1 Overall Protein Distribution

```{r}
ip_per_protein <- sum_results_combined %>%
  filter(rhoG_FDR < 0.05) %>%  # Significant only
  count(trait2, sort = TRUE, name = "n_IPs")

print(ip_per_protein)
```

#### 3.1.2 Graph for frequency distribution

```{r}
ip_per_protein %>%
  ggplot(aes(x = reorder(trait2, n_IPs), y = n_IPs)) +
  geom_col(aes(fill = n_IPs), alpha = 0.8, color = "white", linewidth = 0.3) +
  #Log for better visualisation
 scale_y_continuous(trans = "pseudo_log", 
                   breaks = c(1, 10, 100, 1000)) +
  #Numerical values next to bar
  geom_text(aes(label = n_IPs), hjust = -0.2, size =1.5, color = "black")+
  #Colour scheme
  scale_fill_viridis_c(option = "viridis")+
  #Axis
  coord_flip() +
  #Labelling
  labs(title = "Overall Distribution of IPs per Protein",
       x = "Protein", y = "Number of IPs ") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 6),     
    axis.title.y = element_text(size = 10),
    axis.text.x = element_text(size = 6),
    axis.title.x = element_text(size = 10),
    plot.title = element_text(size = 12, face = "bold",hjust = 0.5),
    legend.position = "none",
    panel.grid.major.y = element_blank() #No vertical lines
  )

#Saving plot
ggsave(
  filename = "Overall_Protein_Distribution.png", 
  path = output_path,
  width = 10,     
  height = 8
)
```

In the 0.1 and 0.3 top hits, find what proteins are most frequent/popular for genetic. Observing whether there are changes in protein frequency patterns.

### 3.2 Genetic h2r 0.1 filtered hits

```{r}
#Count protein frequency in genetic_0.1_hits
protein_frequency_0.1 <- genetic_0.1_hits %>%
  group_by(trait2) %>% #trait2 is the protein column
  summarise(count = n()) %>%
  arrange(desc(count))

#Looking at the top 10
print(head(protein_frequency_0.1, 10))
```

### 3.3 Genetic h2r 0.3 filtered hits

```{r}
#Count protein frequency in genetic_0.3_hits
protein_frequency_0.3 <- genetic_0.3_hits %>%
  group_by(trait2) %>% 
  summarise(count = n()) %>%
  arrange(desc(count))

print(head(protein_frequency_0.3, 10))
```

### 3.4 Environmentally filtered hits

```{r}
# Count how many times each protein appears in your genetic hits
protein_frequency_env <- environmental_hits %>%
  group_by(trait2) %>% # assuming trait2 is your protein
  summarise(count = n()) %>%
  arrange(desc(count))

# Look at the top 10
print(head(protein_frequency_env, 10))
```

## 4. IP-Protein Summary Statistics

Show the distribution of those genetically, environmentally and both significant.

### 4.1 Graphing Correlation Landscape

Visualising the distribution of genetic and environmental correlations.

```{r}
#From summary_results_fdr, seleting those significant for rhoG, rhoE and both
#For purpose of graphing, we redefine the labels
summary_results_fdr_points <- sum_results_combined %>%
 mutate(Sig_Type = case_when(
  rhoG_FDR < 0.05 & rhoE_FDR >= 0.05 ~ "Genetic (rhoG)",
  rhoG_FDR >= 0.05 & rhoE_FDR < 0.05 ~ "Environmental (rhoE)",
  rhoG_FDR < 0.05 & rhoE_FDR < 0.05 ~ "Both ",
  TRUE ~ "Non-Significant"
 ))

#Plotting a graph to show the rhoG and rhoE distribution of all IP -Protein associations
ggplot(summary_results_fdr_points, aes(x = rhoG, y = rhoE, color = Sig_Type)) +
 #Adding each pair as a point
 geom_point(alpha = 0.6, size = 1) +
 #Adding a cross
 geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
 geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
 #Colour scale
 scale_color_viridis_d(option = "viridis") +
 theme_minimal() +
 #Lablelling
 theme(
   plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
   plot.subtitle = element_text(hjust = 0.5, size = 11),
   legend.position = "right",
   legend.text = element_text(size = 12),
   legend.title = element_text(size = 14, face = "bold"),
   legend.key.size = unit(1, "cm"),
   legend.key.width = unit(1, "cm")) +
 labs(
  title = "IP-Protein Association Correlation Landscape",
  subtitle = "Based on FDR-adjusted significance",
  x = "Genetic Correlation (rhoG)",
  y = "Environmental Correlation (rhoE)",
  color = "Correlation Significance"
 )

#Saving plot
ggsave(
  filename = "Overall_Correlation_dotplot.png", 
  path = output_path,
  width = 12,     
  height = 8
)
```

### 4.2 Individual Lineages

Graphing the indivdiual distribution of each lineage

```{r}
ggplot(summary_results_fdr_points, aes(x = rhoG, y = rhoE, color = Sig_Type)) +
  #Adding each pair as a point
  geom_point(alpha = 0.6, size = 1) +
  #Adding a cross
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  
  #Creating separate graphs for the lineages
  facet_wrap(~Sup.Lineage, ncol = 3) + 
  #Colour scale
  scale_color_viridis_d(option = "viridis") +
  
  theme_bw() + 
  theme(
    plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    legend.position = c(0.85, 0.2), #Positioned in empty space
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold"),
    legend.key.size = unit(1, "cm"),
    legend.key.width = unit(1, "cm"),
    strip.background = element_rect(fill = "white"), 
    strip.text = element_text(face = "bold")
  ) +
  
  labs(
    title = "IP-Protein Association Correlation Landscape",
    subtitle = "Grouped by Lineage",
    x = "Genetic Correlation (rhoG)",
    y = "Environmental Correlation (rhoE)",
    color = "Correlation Signficance"
  )


#Saving plot
ggsave(
  filename = "Overall_Correlation_Lineage_dotplot.png", 
  path = output_path,
  width = 12,     
  height = 8
)
```

### 4.3 Summary Table

Summary table from the fdr results based on correlation types for quantification.

```{r}
summary_table <- summary_results_fdr_points %>%
  mutate(across(where(is.numeric), ~ round(., 3)))
  group_by(Sig_Type) %>%
  summarise(
    n = n(),
    mean_rhoG = mean(rhoG),
    mean_rhoE = mean(rhoE),
    mean_h2r1 = mean(h2r1),
    mean_h2r2 = mean(h2r2)
  )

summary_table
#When using this table, best to round to 3dp
```

## 5. Transforming the data frame

For visualisations, the dataframe, which is in a long format, needs to be converted into a wide format (numerical matrix). The dataframes are normalised and NAs were already filtered.

```{r}
#Create a function that applies to all filtered dataset variables
conv_matrix <- function(df, value_col) {
  
  mat <- df %>%
    select(trait1, trait2, all_of(value_col)) %>% 
    
    pivot_wider(
      names_from = trait2,   #changing proteins column
      values_from = all_of(value_col)
    ) %>%

    column_to_rownames("trait1") %>%
    as.matrix()
  
  #replacing NAs from matrix conversion with 0
  mat[is.na(mat)] <- 0
  mat[is.nan(mat)] <- 0
  mat[is.infinite(mat)] <- 0
  
  return(mat)
}
   
```

## 6. Genetic Hits Analysis

With the genetic hits filtered for h2r \> 0.1.

These are genetic hits and likely instigators of inflammaging. Likely related to ageing.

### 6.1 Heatmap

```{r}
genetic_matrix <- conv_matrix(genetic_0.1_hits, "rhoG")

genetic_heatmap<-pheatmap(
            t(genetic_matrix), #transpose so x axis is IP and protein on Y 
            color = viridis(100),
            cluster_rows = TRUE,
            cluster_cols = TRUE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = TRUE,
            show_colnames = FALSE,
            na_colour = "black",  
            main = "Genetic Correlations (IP vs Proteins)",
            border_color = NA
          )
genetic_heatmap
```

Heatmap appears sparce visually.

### 6.2 Shared Protein Distribution

```{r}
#Counting how many share proteins
ip_gen_sharing <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_gen <- table(ip_gen_sharing$protein_count)

shared_protein_ip_gen
```

## 7. Environmental Hits Analysis

Environmental markers - reactive in response externally. Likely not realted to ageing.

### 7.1 Heatmap

```{r}
environmental_matrix <- conv_matrix (environmental_hits, "rhoE")

env_heatmap<-pheatmap(
            environmental_matrix,
            color = viridis(100),
            cluster_rows = TRUE,
            cluster_cols = TRUE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = FALSE,
            show_colnames = TRUE,
            na_colour = "black",  
            main = "Environmental Correlations (IP vs Proteins)",
            border_color = NA
          )
env_heatmap
```

### 7.2 Shared Protein Distribution

```{r}
#Counting how many share proteins
ip_env_sharing <- environmental_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_env <- table(ip_env_sharing$protein_count)

shared_protein_ip_env
```

## 8. Both Genetic and Environmentally Correlated

All hits are positive. Only CXCL9 protein has IP associations.

```{r}
#Plotting with different graph
both_dual <- both_hits %>%
  mutate(IP = reorder(trait1, abs(rhoG))) %>%
  select(IP, Protein = trait2, rhoG, rhoE)

#Lollipop graph to show both rhoG and rhoE
ggplot(both_dual, aes(x = reorder(IP, abs(rhoG)))) +
  geom_segment(aes(xend = IP, y = rhoG, yend = rhoE), color = "grey85", size = 1) +
  geom_point(aes(y = rhoG, color = "Genetic (ρG)"), size = 3) +
  geom_point(aes(y = rhoE, color = "Environmental (ρE)"), size = 3) +
  coord_flip() +
  scale_color_viridis_d(option = "D", begin = 0.2, end = 0.8)+
  labs(
    title = "Both Genetic + Environmental Correlations",
    subtitle = "Only CXCL9 had IPs significant in both genetic and environmental correlations.",
    x = "Immunophenotype", y = "Correlation Coefficient with CXCL9",
    color = "Correlation Type"
  ) +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5)
```

Higher genetic correlations than environmental counterpart.

## 9. Genetic -\> Environmental

Those that are genetically correlated but also have environmental correlations

### 9.1 Heatmap

```{r}
gen_env_matrix <- conv_matrix (gen_env_0.1, "rhoE")

gen_env_heatmap<-pheatmap(
            gen_env_matrix,
            color = viridis(100),
            cluster_rows = FALSE, #As small, set to false
            cluster_cols = FALSE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = FALSE,
            show_colnames = TRUE,
            na_colour = "black",  
            main = "Both Genetic and Environmental Correlations (IP vs Proteins)",
            border_color = NA
          )
gen_env_heatmap
```

### 9.2 Graph Plot

Negative correlations are present as well, though positive correlations dominate. If they are positive in genetics, then they are positive in environmental. No positive genetic and negative environmental.

```{r}
ggplot(gen_env_0.1, aes(x = rhoG, y = rhoE, color = trait2)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = "Genetically Correlated with Environmental Correlations (h² > 0.1)",
    subtitle = "IP–Protein associations coloured by protein",
    x = "Genetic Correlation (rhoG)",
    y = "Environmental Correlation (rhoE)",
    color = "Protein"
  )
```

## 10. Environmental -\> Genetic

### 10.1 Heatmap

```{r}
env_gen_matrix <- conv_matrix(env_gen, "rhoE")

env_gen_heatmap<-pheatmap(
            env_gen_matrix,
            color = viridis(100),
            cluster_rows = FALSE, #As small, set to false
            cluster_cols = FALSE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = FALSE,
            show_colnames = TRUE,
            na_colour = "black",  
            main = "Both Genetic and Environmental Correlations (IP vs Proteins)",
            border_color = NA
          )
env_gen_heatmap
```

All positive correlations, few overlaps.

### 10.2 Plot Graph

```{r}
ggplot(env_gen, aes(x = rhoE, y = rhoG, color = trait2)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = " Environmentally Correlated with Genetic Correlations ",
    subtitle = "IP–Protein associations coloured by protein",
    x = "Environmental Correlation (rhoE)",
    y = "Genetic Correlation (rhoG)",
    color = "Protein"
  )
```

## 
