---
title: "basic_visualisations"
author: "Sarah Lam"
format: html
editor: visual
---

## 1. Set up

#### 1.1 Data Loading

```{r}
#Loading in the solar results analysis filtering.
load("/users/k25046756/Twins_Project/scripts/solar_analysis/solar_analysis_filtering_results.RData")

#Loading in the dataset for .lineage information 
load("/users/k25046756/Twins_Project/data/raw_olink_ip/results/Olink_associations.RData")
```

As a reminder, genetic_0.1_hits, genetic_0.3_hits, environmental_hits and both_hits are FDR filtered 0.05.

gen_env_0.1 and gen_env_0.3 are genetically correlated with environmental correlations.

env_gen is environmentally correlated with genetic correlations.

#### 1.2 Libraries

```{r}
library(tidyr)
library(dplyr)
library(pheatmap) #static

#install.packages("heatmaply") #cant install. need to figure out how 
#library(heatmaply) #Interactive
library(tibble)
library(viridis) #Colour blind friendly palette
library(ggplot2)

#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
```

For the same heatmap scales for fair visualisations.

but is it right to do this? as scales will be uneven and difficult to tell the colours apart for smaller values?

## 

## 3. Genetics Heatmap

```{r}

genetic_matrix <- conv_matrix(genetic_0.1_hits, "rhoG")

genetic_heatmap<-pheatmap(
            t(genetic_matrix), #transpose so x axis is IP and protein on Y 
            color = viridis(100),
            cluster_rows = TRUE,
            cluster_cols = TRUE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = TRUE,
            show_colnames = FALSE,
            na_colour = "black",  
            main = "Genetic Correlations (IP vs Proteins)",
            border_color = NA
          )
genetic_heatmap
```

Heatmap looks empty as majority of IPs are unique to protein, less IPS share among the proteins.

```{r}
# Count how many proteins each IP trait is associated with
ip_gen_sharing <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

# Look at the distribution
table(ip_gen_sharing$protein_count)

# Count how many proteins each IP trait is associated with
ip_env_sharing <- environmental_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

# Look at the distribution
table(ip_env_sharing$protein_count)
```

small subset sharing proteins...worth looking at -\> master IPs? They come from the same lineage?

```{r}
ip_gen_sharing$trait1 <- as.character(ip_gen_sharing$trait1)
infoF$Trait <- as.character(infoF$Trait)


gen_hub_ids <- ip_gen_sharing %>% 
  filter(protein_count == 9) %>% 
  pull(trait1)


gen_hub_details <- infoF %>% 
  filter(SET.NAME %in% gen_hub_ids) %>%
  select(SET.NAME, Lineage, Sup.Lineage, Phenotype, Full.Subset.Name)

print(gen_hub_details)


ip_env_sharing$trait1 <- as.character(ip_env_sharing$trait1)
infoF$Trait <- as.character(infoF$Trait)

env_hub_ids <- ip_env_sharing %>% 
  filter(protein_count == 6) %>% 
  pull(trait1)


env_hub_details <- infoF %>% 
  filter(SET.NAME %in% env_hub_ids) %>%
  select(SET.NAME, Lineage, Sup.Lineage, Phenotype, Full.Subset.Name)

print(env_hub_details)
```

## 4. Environmental Heatmap

```{r}
environmental_matrix <- conv_matrix (environmental_hits, "rhoE")

env_heatmap<-pheatmap(
            environmental_matrix,
            color = viridis(100),
            cluster_rows = TRUE,
            cluster_cols = TRUE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = FALSE,
            show_colnames = TRUE,
            na_colour = "black",  
            main = "Environmental Correlations (IP vs Proteins)",
            border_color = NA
          )
env_heatmap
```

data visualisation point = need to match the scales with genetics as colour is deceiving on strength. Different heatmap scales

## 5. Both

All hits are positive. Only CXCL9 protein but IP associations.

```{r}
#Plotting with different graph
both_dual <- both_hits %>%
  mutate(IP = reorder(trait1, abs(rhoG))) %>%
  select(IP, Protein = trait2, rhoG, rhoE)

#Lollipop graph to show both rhoG and rhoE
ggplot(both_dual, aes(x = reorder(IP, abs(rhoG)))) +
  geom_segment(aes(xend = IP, y = rhoG, yend = rhoE), color = "grey85", size = 1) +
  geom_point(aes(y = rhoG, color = "Genetic (ρG)"), size = 3) +
  geom_point(aes(y = rhoE, color = "Environmental (ρE)"), size = 3) +
  coord_flip() +
  scale_color_viridis_d(option = "D", begin = 0.2, end = 0.8)+
  labs(
    title = "Genetic + Environmental Correlations",
    subtitle = "Only CXCL9 had IPs significant in both genetic and environmental correlations.",
    x = "Immunophenotype", y = "Correlation Coefficient with CXCL9",
    color = "Correlation Type"
  ) +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5)
```

Overall those IP significnat associations with CXCL9 were more genetic then environmental

## 6. gen_env

```{r}
gen_env_matrix <- conv_matrix (gen_env_0.1, "rhoE")

gen_env_heatmap<-pheatmap(
            gen_env_matrix,
            color = viridis(100),
            cluster_rows = FALSE, #As small, set to false
            cluster_cols = FALSE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = FALSE,
            show_colnames = TRUE,
            na_colour = "black",  
            main = "Both Genetic and Environmental Correlations (IP vs Proteins)",
            border_color = NA
          )
gen_env_heatmap
```

Heatmap doesn't work as only three proteins

there are negative correlations too

```{r}
gen_env_dual <- gen_env_0.1 %>%
  mutate(
    IP = reorder(trait1, abs(rhoG))  # Order by genetic strength
  ) %>%
  select(IP, Protein = trait2, rhoG, rhoE)

#separate plots for the proteins
p1 <- ggplot(gen_env_dual %>% filter(Protein == unique(gen_env_0.1$trait2)[1]), 
             aes(x = reorder(IP, abs(rhoG)))) +
  geom_segment(aes(y = 0, yend = rhoG, color = "Genetic (ρG)"), alpha = 0.7) +
  geom_point(aes(y = rhoG, color = "Genetic (ρG)"), size = 3) +
  geom_segment(aes(y = 0, yend = rhoE, color = "Environmental (ρE)"), alpha = 0.7) +
  geom_point(aes(y = rhoE, color = "Environmental (ρE)"), size = 3) +
  coord_flip() +
  scale_color_manual(values = c("Genetic (ρG)" = "#E74C3C", "Environmental (ρE)" = "#3498DB")) +
  labs(title = unique(gen_env_0.1$trait2)[1]) +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  theme(legend.position = "none")

p2 <- ggplot(gen_env_dual %>% filter(Protein == unique(gen_env_0.1$trait2)[2]), 
             aes(x = reorder(IP, abs(rhoG)))) +
  geom_segment(aes(y = 0, yend = rhoG, color = "Genetic (ρG)"), alpha = 0.7) +
  geom_point(aes(y = rhoG, color = "Genetic (ρG)"), size = 3) +
  geom_segment(aes(y = 0, yend = rhoE, color = "Environmental (ρE)"), alpha = 0.7) +
  geom_point(aes(y = rhoE, color = "Environmental (ρE)"), size = 3) +
  coord_flip() +
  scale_color_manual(values = c("Genetic (ρG)" = "#E74C3C", "Environmental (ρE)" = "#3498DB")) +
  labs(title = unique(gen_env_0.1$trait2)[2]) +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  theme(legend.position = "none")

p3 <- ggplot(gen_env_dual %>% filter(Protein == unique(gen_env_0.1$trait2)[3]), 
             aes(x = reorder(IP, abs(rhoG)))) +
  geom_segment(aes(y = 0, yend = rhoG, color = "Genetic (ρG)"), alpha = 0.7) +
  geom_point(aes(y = rhoG, color = "Genetic (ρG)"), size = 3) +
  geom_segment(aes(y = 0, yend = rhoE, color = "Environmental (ρE)"), alpha = 0.7) +
  geom_point(aes(y = rhoE, color = "Environmental (ρE)"), size = 3) +
  coord_flip() +
  scale_color_manual(values = c("Genetic (ρG)" = "#E74C3C", "Environmental (ρE)" = "#3498DB")) +
  labs(title = unique(gen_env_0.1$trait2)[3]) +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  theme(legend.position = "none")

p1
p2
p3
```

## 7. Env_gen

```{r}
env_gen_matrix <- conv_matrix(env_gen, "rhoE")

env_gen_heatmap<-pheatmap(
            env_gen_matrix,
            color = colorRampPalette(brewer.pal(10, "RdYlBu"))(100),
            cluster_rows = FALSE, #As small, set to false
            cluster_cols = FALSE,
            fontsize_row = 6,
            fontsize_col = 6,
            show_rownames = FALSE,
            show_colnames = TRUE,
            na_colour = "black",  
            main = "Both Genetic and Environmental Correlations (IP vs Proteins)",
            border_color = NA
          )
env_gen_heatmap
```

Heatmap doesnt work as only two proteins

# Network Graphs

# Genetic associations

New method: Finding the genetic associations/correlations between IP and IP . Then with protein and protein.

Instead of running solar to find the genetic correlations

1.  construct a matrix. Rows as IPs, columns as proteins, entries as pG
2.  IP IP = correlation between rows (Ip profiles across proteins)
3.  Protein protein -\> already ran solar but correlation between columns

# IP-Protein Network graph

```{r}
#Matrix conversion from genetic_0.1_hits
#Using RhoG for the biological connection / proximity as it is already filtered
#Based on shared proteins

g01_long <- genetic_0.1_hits %>%
  select(trait1, trait2, rhoG)%>%
  mutate(
    IP = if_else(grepl("^IMMUNO_",trait1), trait1, trait2),
    protein = if_else(grepl("^IMMUNO_", trait1), trait2, trait1)
  ) %>%
  filter(grepl("^IMMUNO_", IP)) %>% 
  select(IP, protein, rhoG)

IP_protein_mat <- g01_long %>%
  pivot_wider(
    names_from  = protein,
    values_from = rhoG
  ) %>%
  column_to_rownames("IP") %>%
  as.matrix()


# number of non-NA proteins per IP
n_proteins_per_IP <- rowSums(!is.na(IP_protein_mat))

#Filter IPs that have 2+ protein associations, the rest are unique to one protein
min_proteins <- 2
keep_IPs <- names(n_proteins_per_IP)[n_proteins_per_IP >= min_proteins]

IP_protein_mat_filt <- IP_protein_mat[keep_IPs,]

```

```{r}
edges <- genetic_0.1_hits %>%
  mutate(
    weight = abs(rhoG),           # Edge strength
    sign = ifelse(rhoG > 0, "pos", "neg")  # Positive/negative correlation
  ) %>%
  select(from = trait1, to = trait2, weight, rhoG, sign)


all_nodes <- unique(c(edges$from, edges$to))
nodes <- data.frame(
  name = all_nodes,
  type = ifelse(grepl("^IMMUNO_", all_nodes), "IP", "Protein")
)

g_ip_prot <- graph_from_data_frame(edges, vertices = nodes, directed = FALSE)


cl_ip_prot <- cluster_louvain(g_ip_prot, weights = E(g_ip_prot)$weight)
V(g_ip_prot)$module <- membership(cl_ip_prot)
V(g_ip_prot)$degree <- degree(g_ip_prot)


tg_ip_prot <- as_tbl_graph(g_ip_prot)

ggraph(tg_ip_prot, layout = "fr") +
  geom_edge_link(aes(width = weight, color = rhoG), 
                 alpha = 0.7) +
  scale_edge_width(range = c(0.3, 2)) +
  scale_edge_color_gradient2(low = "blue", mid = "grey80", high = "red") +
  geom_node_point(aes(color = factor(module), shape = type), 
                  size = 4, alpha = 0.9) +
  geom_node_text(aes(label = ifelse(degree > 5, name, NA)), 
                 repel = TRUE, size = 2.5) +
  scale_shape_manual(values = c("IP" = 16, "Protein" = 15)) +
  theme_void() +
  labs(title = "IP-Protein Bipartite Genetic Network",
       subtitle = "Modules colored, ρG edge weights, IP=circle Protein=square")

```

### Genetic correlations between IP-IP

```{r}
#IP-IP associations through pearsons correlation for genetic similarities based on proteins
#Pearsons based on pG patterns across proteins
IP_IP_cor <- cor(t(IP_protein_mat_filt),
                 use   = "pairwise.complete.obs",
                 method = "pearson")


#Dimensions are 923 so 923 hub IPs -> potential master regulators 
dim(IP_IP_cor)
rownames(IP_IP_cor)[1:5]

```

```{r}

#from correlation matrix, create edges
IP_edges <- IP_IP_cor %>%
  as.data.frame() %>%
  rownames_to_column("IP1") %>%
  pivot_longer(-IP1, names_to = "IP2", values_to = "cor") %>%
  filter(IP1 < IP2, cor > 0.7, !is.na(cor))  #Can change threhold and is only positive

#changed the correlation to 0.7 and 0.5 was messy
g_IP <- graph_from_data_frame(IP_edges, directed = FALSE)

#Modules
cl_ip <- cluster_louvain(g_IP, weights = IP_edges$cor)

plot(cl_ip, g_IP,
     vertex.size  = 3,
     vertex.label.cex = 0.8, 
     vertex.label.dist= 0.5,
     edge.width   = 1)
```

hard to visualise

```{r}
#Tidygraph
tg_ip <- tbl_graph(
  nodes = data.frame(
    name=V(g_IP)$name, 
    module=membership(cl_ip)),
  edges = IP_edges %>% rename(weight = cor), 
  directed=FALSE)

#Plotting with ggraph
ggraph(tg_ip, layout="fr") +
  geom_edge_link(aes(color = factor(to), width=log(weight + 1.1 )),
                 alpha=0.5, 
                 start_cap = circle(3,'mm'),
                 end_cap = circle(3, 'mm')) +
  scale_edge_width(range = c(0.5, 2.5)) +
  geom_node_point(aes(color=factor(module)), size=2, alpha = 0.5) +
  geom_node_text(aes(label=name), repel=TRUE, size=2) +
  theme_void() +
  labs(title="IP Genetic Similarity Modules")
```

filter by lineage as graph crashes too much memory

# Protein protein associations

```{r}
protein_protein_cor <- cor(IP_protein_mat_filt ,
                           use = "pairwise.complete.obs",
                           method = "pearson")
dim(protein_protein_cor)

```

```{r}
protein_edges <- protein_protein_cor %>%
  as.data.frame() %>%
  rownames_to_column("protein1") %>%
  pivot_longer(-protein1, names_to = "protein2", values_to = "cor") %>%
  filter(protein1 < protein2,  cor > 0.7, !is.na(cor))

g_prot <- graph_from_data_frame(protein_edges, directed = FALSE)


#Modules
cl_prot <- cluster_louvain(g_prot, weights = protein_edges$cor)

plot(cl_prot, g_prot,
     vertex.size  = 3,
     vertex.label.cex = 0.8, 
     vertex.label.dist= 0.5,
     edge.width   = 1)
```

```{r}
#Tidygraph
tg_prot <- tbl_graph(
  nodes = data.frame(
    name=V(g_prot)$name, 
    module=membership(cl_prot)),
  edges = protein_edges %>% rename(weight = cor), 
  directed=FALSE)%>%
  mutate(importance = centrality_degree())

#Plotting with ggraph
ggraph(tg_prot, layout="fr") +
  geom_edge_link(aes(color = factor(to), width=log(weight + 1.1 )),
                 alpha=0.5, 
                 start_cap = circle(3,'mm'),
                 end_cap = circle(3, 'mm')) +
  scale_edge_width(range = c(0.5, 2.5)) +
  geom_node_point(aes(color=factor(module), size = importance), alpha = 0.5) +
  scale_size_continuous(range = c(2, 7)) +
  geom_node_text(aes(label=name), repel=TRUE, size=2) +
  theme_void() +
  labs(title="Protein Genetic Similarity Modules")
```

# infoF lineage exploration

```{r}
table(infoF$Sup.Lineage)
```
