---
title: "Enrichment_Analysis"
format: html
editor: visual
---

# Enrichment Analysis

Idnetifying enrichment or depletion in the Genetic and Environmental filtered dataset to determine statistical outcomes. Using Fisher's exact test enables two sided (enriched or deleted) assessment for lineages of IPs

## 1. Set up

### 1.1 Load libraries

```{r, message=FALSE}
library(dplyr)
```

### 1.2 Loading Data

```{r, message=FALSE}
load("solar_analysis_filtering_results.RData")
```

### 1.3 Output path

```{r}
output_path <- "~/Twins_Project/notebooks"
```

## 2 IP info

Obtaining sup and sub lineages for information in each subset

```{r}
#Get all IPs from data
all_background_ips <- sum_results_combined %>% 
  distinct(trait1, Sup.Lineage, Sub.Lineage)

#Get IPs from each
gen_hit_ips <- genetic_0.1_hits %>% distinct(trait1) %>% pull(trait1)
env_hit_ips <- env_2x_hits %>% distinct(trait1) %>% pull(trait1)

```

## 3 Fishers

Creating a function to build the matrix for both. FDR also applied.

```{r}
run_enrichment <- function(hit_list, background_df, target_column) {
  
  results <- lapply(unique(background_df[[target_column]]), function(lineage) {
    if(is.na(lineage)) return(NULL) 
    
    #Building 2x2 matrix
    
    #Number of IPs in gen/env
    in_hits_is_lin     <- sum(background_df$trait1 %in% hit_list & background_df[[target_column]] == lineage)
    
    in_hits_NOT_lin    <- sum(background_df$trait1 %in% hit_list & background_df[[target_column]] != lineage)
    
    #Ips that didnt make it into gen/env from total
    NOT_hits_is_lin    <- sum(!(background_df$trait1 %in% hit_list) & background_df[[target_column]] == lineage)
    
    NOT_hits_NOT_lin   <- sum(!(background_df$trait1 %in% hit_list) & background_df[[target_column]] != lineage)
    
    #Table/matrix comparison
    contingency_matrix <- matrix(c(in_hits_is_lin, in_hits_NOT_lin, 
                                   NOT_hits_is_lin, NOT_hits_NOT_lin), nrow = 2)
    
    #Run two sided test
    test <- fisher.test(contingency_matrix)
    
    #Creating the df
    data.frame(
      Sup_or_Sub = target_column, #For sub and sup identification
      Lineage = lineage,
      Hits_Count = in_hits_is_lin,
      Total_in_all_results = in_hits_is_lin + NOT_hits_is_lin,
      Odds_Ratio = test$estimate,
      P_Value = test$p.value
    )
  })
  
  #Apply FDR 
  final_table <- bind_rows(results) %>%
    mutate(FDR = p.adjust(P_Value, method = "fdr")) %>%
    mutate(Status = case_when(
      FDR < 0.05 & Odds_Ratio > 1 ~ "Enriched",
      FDR < 0.05 & Odds_Ratio < 1 ~ "Depleted",
      TRUE ~ "Not Significant"
    )) %>%
    arrange(Sup_or_Sub, FDR)
  
  return(final_table)
}
```

## 4 Output

### Genetic

```{r}
gen_sup_enrichment <- run_enrichment(gen_hit_ips, all_background_ips, "Sup.Lineage")
gen_sub_enrichment <- run_enrichment(gen_hit_ips, all_background_ips, "Sub.Lineage")

#Bind Sup and sub lineages
genetic_enrichment_results <- bind_rows(gen_sup_enrichment, gen_sub_enrichment)
print(genetic_enrichment_results)


```

```{r}
write.csv(genetic_enrichment_results, file = file.path(output_path, "gen_ip_lin_enrichment.csv"), row.names = FALSE)

```

### Environmental

```{r}
env_sup_enrichment <- run_enrichment(env_hit_ips, all_background_ips, "Sup.Lineage")
env_sub_enrichment <- run_enrichment(env_hit_ips, all_background_ips, "Sub.Lineage")


#Bind Sup and sub lineages
environmental_enrichment_results <- bind_rows(env_sup_enrichment, env_sub_enrichment)
print(environmental_enrichment_results)


```

```{r}
write.csv(environmental_enrichment_results, file = file.path(output_path, "env_2x_ip_lin_enrichment.csv"), row.names = FALSE)
```
