---
title: "IP_IP_Genetic_Correlation"
format: html
editor: visual
---

# Visualising Genetic Correlations

From the SOLAR analysis significant genetic results, we want to obtain IP-IP

## 1. Set Up

### 1.1 Load Libraries

```{r, message = FALSE}
library(tidyr)
library(dplyr)
library(tibble)

#Graphs
library(viridis) #Colour blind friendly palette
library(ggplot2)

#Panelling
library(patchwork) 
library(cowplot)
library(png)
library(grid)

#Sankey
library(networkD3)
library(htmlwidgets)
library(htmltools)
library(webshot2)

#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
library(tidyverse)

#Chord
library(circlize)
```

### 1.2 Data Loading

```{r}
#Loading in the solar results analysis filtering.
load("~/Twins_Project/notebooks/solar_analysis_filtering_results.RData")

```

As a reminder, genetic_0.1_hits, genetic_0.3_hits, environmental_hits and both_hits are FDR filtered 0.05.

gen_env_0.1 and gen_env_0.3 are genetically correlated with environmental correlations.

env_gen is environmentally correlated with genetic correlations.

sum_results_combined contains the FDR filtered plus lineage columns.

## 1.3 Save path

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/visualisations/ip_ip"
```

# IP-IP Genetic Associations

New method: Finding the genetic associations/correlations between IP and IP .

Ideally, I would run IP-IP SOLAR on all IPs to find the genetic correlations but it is too computationally intensive. Instead, using the already genetically filtered genetic_0.1_hits where we know IP-Protein associations were heritable, we can take the selected IPs (4300) then that has been narrowed down to identify only IPs that share 2+ proteins, disregarding those unique to just one, and build IP profiles across proteins. Find IP-IP correlations between rows. However this doesn't mean IP-IP are directly genetically correlated as they could be connected to the protein through different genetic pathways. This can only infer/estimate but not a direct result.

If I use this way to find the genetic associatiosn with IP-IP then wouldn't I also need to apply this logic/step to protein-protein associations as well?

-   Genetic Correlation with **Pearsons**

    -   As we used SOLAR to prove those associations were genetically driven, we can use Pearson to map the observed strength of the associations.

-   Genetic Correlations building on the SOLAR protein-protein network results with **Jaccard similarity**

    -   if IP 1 is associated with protein x and IP 2 is associated with protein Y and the current network shows protein x and y have high pG, then IP 1 and 2 are genetically correlated by proxy.

    -   But this method relies on the SOLAR results which may not be representative?

    -   This would involve (1) creating IP-Protein mapping matrix and 1 if SOLAR associated, 0 if not , (2) using exisiting protein-protein matrix , (3) multiply them ip-protein matrix x protein-protein matrix

    -   I need to find IPs that share the SAME proteins sets.

    -   However, this may be down to sharing the same cell lineage which is a known relationship.

With finding the IP-IP matrix, there is potential to explore the cell lineages.

## 2 Matrix

Conversion of the genetic_0.1_hits is necessary from a long to wide format with rows as IPs, columns as proteins and entries as RhoG (pG) . There will be lots of NAs as the majority of IPs do not significnatly associate with 2+ proteins.

Problem is that not all IPs share multiple proteins, 1464 IPs are unique to one protein but the rest (2836) are significantly correlated to 2+ proteins.

```{r}
#Matrix conversion from genetic_0.1_hits
#Using RhoG for the biological connection / proximity as it is already filtered
#Based on shared proteins
g01_long <- genetic_0.1_hits %>%
  select(trait1, trait2, rhoG)%>%
  mutate(
    IP = if_else(grepl("^IMMUNO_",trait1), trait1, trait2),
    protein = if_else(grepl("^IMMUNO_", trait1), trait2, trait1)
  ) %>%
  filter(grepl("^IMMUNO_", IP)) %>% 
  select(IP, protein, rhoG)

IP_protein_mat <- g01_long %>%
  pivot_wider(
    names_from  = protein,
    values_from = rhoG
  ) %>%
  column_to_rownames("IP") %>%
  as.matrix()

#number of non-NA proteins per IP
n_proteins_per_IP <- rowSums(!is.na(IP_protein_mat))

#Filter IPs that have 2+ protein associations, the rest are unique to one protein
#Filtering for 2+ proteins allows us to identify master regulations and identify where one instigator IP controls multiple responder proteins
min_proteins <- 2
keep_IPs <- names(n_proteins_per_IP)[n_proteins_per_IP >= min_proteins]

IP_protein_mat_filt <- IP_protein_mat[keep_IPs,]

```

## 3 Jaccard Similarity

### 3.1 Shared Protein Distribution

Displaying the distribution of IPs sharing Protein counts

```{r}
#Counting how many share proteins
ip_gen_sharing <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_gen <- table(ip_gen_sharing$protein_count)

shared_protein_ip_gen
```

### 3.2 Jaccard Similairty

Calculating the same sharing with the Jaccard similarity index

```{r}
#List of all unique ip-protein pairs 
ip_links <- g01_long %>%
  select(IP, protein) %>%
  distinct()

#Counting how many we have
ip_totals <- ip_links %>%
  count(IP, name = "total_proteins")

#those that are sharing 
ip_sharing <- ip_links %>%
  inner_join(ip_links, by = "protein", relationship = "many-to-many") %>%
  rename(IP_A = IP.x, IP_B = IP.y) %>%
  filter(IP_A < IP_B) %>% #keep different IP pairs with the same protein - no redundancy 
  group_by(IP_A, IP_B) %>% #counting how mnay common proteins
  summarise(shared_count = n(), .groups = "drop")

#Jaccard Score caluculations
ip_jaccard <- ip_sharing %>%
  left_join(ip_totals, by = c("IP_A" = "IP")) %>%
  left_join(ip_totals, by = c("IP_B" = "IP")) %>%
  mutate(
    jaccard_score = shared_count / (total_proteins.x + total_proteins.y - shared_count)
  ) %>%
  arrange(desc(jaccard_score))

```

But jaccard index measures overlap not absolute strength

Those that share more proteins than those that only share one matter more for jaccard score.

Thus, weighting and filtering are applied to quantify similarity with more nuance. Weighting so IP pairs with overall more proteins shared and in common are favoured then IP pairs with overall less proteins shared though in common.

```{r}
weight_ip_jaccard <- ip_jaccard %>%
  filter(shared_count > 1) %>% #Only keep pairs sharing 2+ proteins
  mutate(weighted = jaccard_score * shared_count) %>%
  arrange(desc(weighted))
```

### 3.3 Lineage Tracing

From the Jaccard similarity results, if they have the same cell lineage then it is a known relationship so they must share the same proteins.

If they derive from different lineages then there is a possibility that this is an instigator as they come from different pathways/parts of the immune system.

#### Filtering by lineage relationships

```{r}
#Need to select Ip and Sub.Lineage 
lineage_ip <- genetic_0.1_hits %>%
  select(IP=trait1, Sup.Lineage, Sub.Lineage) %>%
  distinct()

#joining the lineage 
ip_lineage_check <- weight_ip_jaccard %>%
  left_join(lineage_ip, by = c("IP_A" = "IP")) %>%
  rename(Sup_A = Sup.Lineage, Sub_A = Sub.Lineage) %>%
  
  left_join(lineage_ip, by = c("IP_B" = "IP")) %>%
  rename(Sup_B = Sup.Lineage, Sub_B = Sub.Lineage) %>%
  
  mutate(Relationship = if_else(Sup_A == Sup_B, "Intra-Lineage", "Cross-Lineage")) %>%
  filter(shared_count > 1) %>% #Applying filtering 
  arrange(desc(weighted))

head(ip_lineage_check, 20)
```

All of the IPs that have high scores are of the same lineage (intra-lineage). This is natural as strongest genetic signals are from the same family tree. Lineage driven, structured by immune lineage.

High Jaccard + High weight (9 proteins shared -\> hub of IPs). Maybe multiple pathways triggered.

```{r}
#How many unique lineages in total

cat("Unique Sub.Lineages\n")
ip_lineage_check %>%
  select(Sub_A, Sub_B) %>%
  unlist() %>%
  unique() %>%
  sort() %>%
  print()

cat("\nUnique Sub.Lineages Intra-Lineages\n")
ip_lineage_check %>%
  filter(Relationship == "Intra-Lineage") %>%
  select(Sub_A) %>% # Since A and B are the same in Intra, we only need one
  distinct() %>%
  pull(Sub_A) %>%
  sort() %>%
  print()

cat("\nUnique Sub.Lineages Cross-Lineages\n")
ip_lineage_check %>%
  filter(Relationship == "Cross-Lineage") %>%
  select(Sub_A, Sub_B) %>%
  unlist() %>%
  unique() %>%
  sort() %>%
  print()
```

## 4 Graphs - Overall Counts

### 4.1 IP-IP Shared Protein By Lineage Relationship

```{r}
ip_bar_shared <- ggplot(ip_lineage_check, aes(x = as.factor(shared_count))) +
  geom_bar(aes(fill = as.factor(shared_count)), color = "white", alpha = 0.8,show.legend = FALSE) +
  #Label
  geom_text(stat = "count", 
            aes(label = after_stat(count)), 
            vjust = -0.5, size = 3, fontface = "bold") +
  
  #Colour
  scale_fill_viridis_d(option = "viridis") +
  
  #Facet Panel
  facet_wrap(~Relationship, ncol = 1) +
  #Scales
  scale_y_log10(breaks = c(1, 10, 100, 1000, 10000, 100000),
                limits = c(1, 150000),
                labels = function(x) format(x, big.mark = ",", scientific = FALSE)
  ) +
  #Labels
  labs(
    title = "IP-IP Pairs By Protein Shared Frequency",
    x = "Number of Shared Proteins",
    y = "Frequency (IP-IP Pairs)"
  ) +
  
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(face = "bold", size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    )

#Saving plot
ggsave(
  filename = "IP_IP_Prot_Shared.png", 
  plot = ip_bar_shared,
  path = output_path,
  width = 8,     
  height = 8
)

```

### 4.2 Density of Jaccard Scores

```{r}
#Need to group to find frequency
density_freq_plot_data <- ip_lineage_check %>%
  group_by(Relationship, shared_count, jaccard_score) %>%
  summarise(freq = n(), .groups = "drop")

#Plot
density_jaccard_plot <- ggplot(density_freq_plot_data, aes(x = as.factor(shared_count), y = jaccard_score)) +
  #Point
  geom_point(aes(fill = freq, size = freq), shape = 21, color = "black", alpha = 0.8) +
  #Log Colours
  scale_fill_viridis_c(trans = "log10", option = "viridis", name = "Frequency (log)",breaks = c(1, 10, 100, 1000, 10000), guide = "legend") +
  
  scale_size_continuous(trans = "log10", range = c(2, 10), name = "Frequency (log)", breaks = c(1, 10, 100, 1000, 10000)) +
  #Panel
  facet_wrap(~Relationship, ncol = 1) +
  
  labs(
    title = "Jaccard Score Density",
    x = "Number of Shared Proteins",
    y = "Jaccard Similarity Score"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 12),
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold", size = 12),
    legend.position = "right"

  )

#Saving plot
ggsave(
  filename = "IP_IP_Jac_Score.png", 
  plot = density_jaccard_plot,
  path = output_path,
  width = 8,     
  height = 8
)
```

### 4.3 Combining the previous two graphs for paneling

```{r}
ip_bar_shared <- ip_bar_shared + 
  labs(tag = "A") + 
  theme(plot.tag.position = c(0.02, 0.95), plot.tag = element_text(size = 20, face = "bold"))

density_jaccard_plot <- density_jaccard_plot + 
  labs(tag = "B") + 
  theme(plot.tag.position = c(-0.02, 0.95), plot.tag = element_text(size = 20, face = "bold"))

combined_panels <- (ip_bar_shared | plot_spacer() | density_jaccard_plot) +
  plot_layout(widths = c(1, 0.25, 1))

ggsave(
  filename = "IP_IP_Panels_Combined.png", 
  plot = combined_panels,
  path = output_path,
  width = 18,     
  height = 8,
  dpi = 600
)
```

## 5 Intra Lineage

### 5.1 Filtering

```{r}
intra_lineage <- ip_lineage_check %>%
  filter(Relationship == "Intra-Lineage") %>%
  filter(shared_count > 1) %>% #Filtered so no unique pairs
  arrange(desc(weighted))

head(intra_lineage, 20)
```

We know from genetics_0.1_hits that there are hubs for the IPs and the number of proteins they share.

### 5.2 Modules

Identify what modules based on Louvain clustering and their specific lineages.

```{r}
#Filter for  hub 
ip_intra_hub <- intra_lineage %>%
  filter(shared_count >= 7) %>%
  select(IP_A, IP_B, weighted)

#Graph
graph_obj <- graph_from_data_frame(ip_intra_hub, directed = FALSE)

#Louvain clustering
communities <- cluster_louvain(graph_obj)

#Getting the IP clusters for numerical information
ip_clusters <- data.frame(
  IP = names(membership(communities)),
  Cluster = as.numeric(membership(communities))
)

#Obtaining lineage
final_clusters <- ip_clusters %>%
  left_join(lineage_ip, by = "IP")

#Inspecting the summary 
cluster_summary <- final_clusters %>%
  group_by(Cluster, Sub.Lineage) %>%
  summarise(n = n(), .groups = "drop") %>%
  pivot_wider(names_from = Sub.Lineage, values_from = n, values_fill = 0)

print(cluster_summary)
```

Even in each super lineage, sub lineages do not cross talk. No mixing of cells. This infers stability.

### 5.3 Heatmap for Cluster

```{r}
#Convert to long format from wide
cluster_summary_long <- cluster_summary %>%
  pivot_longer(
    cols = -Cluster, 
    names_to = "Sub_Lineage", 
    values_to = "Count"
  )

#Heatmap for 7+ protein sharing hub as above
cluster_7_heatmap<-ggplot(cluster_summary_long, aes(x = as.factor(Cluster), y = Sub_Lineage, fill = Count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "#440154FF") +#Colour scale
  labs(title = "Intra-Lineage Clusters (7+ Shared Proteins)",
       x = "Cluster (Louvain)",
       y = "Sub-Lineage") +
  #Aesthetic
  theme_bw() +
  theme(
    plot.title = element_text(size = 12, face = "bold",hjust = 0.5, margin = margin(b = 10)),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(face = "bold", size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 8),     
    axis.title.y = element_text(size = 10,face = "bold"),
    axis.text.x = element_text(size = 8),
    axis.title.x = element_text(size = 10,face = "bold")
    ) 

#Saving plot
ggsave(
  filename = "Intra_Cluster.png", 
  plot = cluster_7_heatmap,
  path = output_path,
  width = 8,     
  height = 6
)
```

## 6 Cross lineage

### 6.1 Filtering

```{r}
cross_lineage <- ip_lineage_check %>%
  filter(Relationship == "Cross-Lineage") %>%
  filter(shared_count > 1) %>% #Filtered so no unqiue pairs
  arrange(desc(weighted))

head(cross_lineage, 20)
```

There are a few cross lineage. Top cross have weighted scores of 3.2 which is low (shared proteins 4 out of 5) though these are both T cells. Regardless of weight, checking jaccard score 1.0. These have 3 proteins shared with cross lineages of B cells and T cells. These have high jaccard + low weight -\> specific links, maybe only one pathway triggered.

Specific sub.lineages has highest connection to others? How many sub connects to sub

### 6.2 Bridge Analysis

Finding out what proteins these top cross lineage pairs have

Top 3 proteins shared with a score of 1

```{r}
#Filter for 3+ as these
cross_lineage_best <- cross_lineage %>%
  filter(shared_count >= 3 & jaccard_score==1) 
```

Determining the unique lineages in the crosses.

```{r}
cross_lineage_best %>%
  select(Sup_A, Sup_B) %>%
  unlist() %>%      
  unique() %>%   
  sort()
```

Only B cells and T cells.

Obtaining the proteins they are paired to and the counts of pairs.

```{r}
#Obtaining protein names
shared_protein_map <- cross_lineage_best %>%
  inner_join(genetic_0.1_hits, by = c("IP_A" = "trait1")) %>%
  inner_join(genetic_0.1_hits, by = c("IP_B" = "trait1", "trait2" = "trait2")) %>%
  select(IP_A, IP_B, Protein = trait2)

#Counting which proteins have the most cross-lineage IP-IP pairs
bridge_protein_counts <- shared_protein_map %>%
  group_by(Protein) %>%
  summarise(bridge_count = n()) %>%
  arrange(desc(bridge_count))

print(bridge_protein_counts)
```

There are CXCL10, CXCL9, CXCL11 (chemokine triad) and IL_12B. CXCL10 and CXCL9 persist in all IP-IP pairs. This is followed by CXCL11 then IL12B

Finding out what specific cross lineage (Sub.Lineages) bridges there are

```{r}
#Creating a df that quantifies the pairs
protein_lineage_bridge <- shared_protein_map %>%
  inner_join(cross_lineage_best, by = c("IP_A", "IP_B")) %>%
  group_by(Protein, Sub_A, Sub_B) %>%
  summarise(pair_numbers = n(), .groups = "drop") %>%
  arrange(desc(pair_numbers))

#Make sure no redundant lineage pairs 
#Organise all A columns = T cells, all B columns = B cells
protein_lineage_bridge <- cross_lineage_best %>%
  mutate(
    #Check if need to swap
    should_swap = (Sup_A == "B cells"),
    #Swapping
    temp_Sup_A = Sup_A, temp_Sub_A = Sub_A,
    temp_Sup_B = Sup_B, temp_Sub_B = Sub_B,
    Sup_A = if_else(should_swap, temp_Sup_B, temp_Sup_A),
    Sub_A = if_else(should_swap, temp_Sub_B, temp_Sub_A),
    Sup_B = if_else(should_swap, temp_Sup_A, temp_Sup_B),
    Sub_B = if_else(should_swap, temp_Sub_A, temp_Sub_B)
  ) %>%
  inner_join(shared_protein_map, by = c("IP_A", "IP_B")) %>%
  group_by(Protein, Sup_A, Sub_A, Sup_B, Sub_B) %>%
  summarise(pair_numbers = n(), .groups = "drop") %>% #Number of pairs for all 3 proteins
  arrange(desc(pair_numbers))
  

print(protein_lineage_bridge)
```

### 

### Interactive Sankey for Top Cross Lineage

This is interactive. Once html opened, you can hover on the connections to find out how many IP-IP pairs displayed that connection.

```{r}
#Ordering of T cell -> Sub_A -> Protein -> Sub_B -> B cell
sankey_data <- protein_lineage_bridge %>%
  #Need to add in T cell and B cell into the df
  mutate(
    Sup_A = "T cells",
    Sup_B = "B cells" 
  ) %>%
  select(Sup_A, Sub_A, Protein, Sub_B, Sup_B, pair_numbers) 

#Create links for each flow
l1 <- sankey_data %>% select(Sup_A, Sub_A, pair_numbers) %>% rename(source=Sup_A, target=Sub_A, value=pair_numbers)
l2 <- sankey_data %>% select(Sub_A, Protein, pair_numbers) %>% rename(source=Sub_A, target=Protein, value=pair_numbers)
l3 <- sankey_data %>% select(Protein, Sub_B, pair_numbers) %>% rename(source=Protein, target=Sub_B, value=pair_numbers)
l4 <- sankey_data %>% select(Sub_B, Sup_B, pair_numbers) %>% rename(source=Sub_B, target=Sup_B, value=pair_numbers)

#Binding in order
links <- bind_rows(l1, l2, l3, l4)

#Nodes
nodes <- data.frame(name = unique(c(links$source, links$target)))
nodes$id <- 0:(nrow(nodes)-1)

#Colours
num_nodes <- nrow(nodes)
viridis_colors <- viridis::viridis(num_nodes)
js_array <- paste0('["', paste(viridis_colors, collapse = '", "'), '"]')

#IDs
links <- links %>%
  left_join(nodes, by = c("source" = "name")) %>% rename(source_id = id) %>%
  left_join(nodes, by = c("target" = "name")) %>% rename(target_id = id)


#Plotting
ip_sankey <- sankeyNetwork(Links = links, Nodes = nodes, Source = "source_id", Target = "target_id",
              Value = "value", NodeID = "name", units = "Interactions",
              fontSize = 18, nodeWidth = 30,
              width = 1200,
              height = 600
              )


#Modification for the colours
ip_sankey <- htmlwidgets::onRender(
  ip_sankey,
  paste0('
    function(el, x) {
      var colors = ', js_array, ';
      d3.select(el).selectAll(".node rect")
        .style("fill", function(d, i) {return colors[i]; });
    }
  ')
)

#Format and Title
sankey_file_path <- file.path(output_path, "Cross_Lineage_Sankey.html")

save_html(
  tagList(
    tags$h2("Cross-Lineage Shared Protein Flow", style = "text-align:center; font-family: Arial;"),
    tags$div(style = "width: 1200px; margin: auto;", ip_sankey)
  ), 
  file = sankey_file_path
)

#HTML version
browseURL(sankey_file_path)



#Static Version
#To obtain static, conduct locally rather than HPC
webshot(
  url = sankey_file_path, 
  file = file.path(output_path, "Cross_Lineage_Sankey_Static.png"),
  vwidth = 1300, 
  vheight = 700,
  delay = 2
)

```

# Graphing

OVERALL GOAL: These top IP-IP pairs, need to compare with which proteins are they sharing

## Chord Diagram

### Aggregating the data

```{r}
#Aggregating the data
chord_df <- ip_lineage_check %>%
  #Make sure no redundant pairs
  mutate(
    Sub_1 = pmin(Sub_A, Sub_B),
    Sub_2 = pmax(Sub_A, Sub_B)
  ) %>%
  group_by(Sub_1, Sub_2, Sup_A, Sup_B, Relationship) %>%
  summarise(value = sum(weighted), .groups = "drop")%>% #can filter for stronger signals with  filter(value > quantile(value, 0.8)) 
  rename(Sub_A = Sub_1, Sub_B = Sub_2)
```

### Intra lineage

```{r}

#Saving plot
intra_file_name <- file.path(output_path, "Intra_Lineage_chord.png")
png(intra_file_name, width = 8, height = 8, units = "in", res = 300)
par(mar = c(0, 0, 0, 0))

#Intra df for chord
intra_df <- chord_df %>% filter(Relationship == "Intra-Lineage")

#Assigning colours for unique sub lineages
all_sub_lineages <- sort(unique(c(intra_df$Sub_A, intra_df$Sub_B)))
sub_colors <- viridis(length(all_sub_lineages), option = "viridis")
names(sub_colors) <- all_sub_lineages

#Circle parameters
circos.par(gap.after = 2, 
           start.degree = 180,
           canvas.xlim = c(-1.1, 1.1), 
           canvas.ylim = c(-1.1, 1.1)) 

#Order so lineages are together
order_circle <- c(
  "CD4 T cells", "CD8 T cells", "DP T cells", 
  "Memory CD8 T cells", "Naive CD8 T cells", 
  "Monocytes", "B cells/A", "B cells/D", 
  "B cells/E", "B cells/G", "B cells/M"
)

#Diagram
chordDiagram(
  x = intra_df %>% select(Sub_A, Sub_B, value),
  order = order_circle ,
  grid.col = sub_colors,
  transparency = 0.6,
  annotationTrack = "grid",
  preAllocateTracks = list(track.height = 0.07)
)

#Labelling
circos.track(track.index = 1, panel.fun = function(x, y) 
  #Every sub lineage labelled
  {circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
              #Ensure it is readable, not upside down
              facing = "clockwise", 
              niceFacing = TRUE, 
              adj = c(0, 0.5), 
              cex = 1) # size
}, bg.border = NA)

#Title 
title(main = "Intra-Lineage Connections", 
      cex.main = 1.5, 
      font = 2,
      line =-3)

#Save with dev.off
dev.off()
#Clear before next use
circos.clear()
```

### Cross lineage

This is shared count \>= 3 & jaccard_score = 1

```{r}
#Saving plot
cross_file_name <- file.path(output_path, "Cross_Lineage_chord.png")
png(cross_file_name, width = 8, height = 8, units = "in", res = 300)
par(mar = c(0, 0, 0, 0))

#Cross df
cross_chord_df <- cross_lineage_best%>% 
  group_by(Sub_A, Sub_B) %>% 
  summarise(value = sum(weighted), .groups = "drop")

#Assigning colours for unique sub lineages
all_nodes <- sort(unique(c(cross_chord_df$Sub_A, cross_chord_df$Sub_B)))
cross_colors <- viridis(length(all_nodes), option = "viridis")
names(cross_colors) <- all_nodes

#Circle parameters
circos.par(gap.after = 2, 
           start.degree = 90,
           canvas.xlim = c(-1.1, 1.1), 
           canvas.ylim = c(-1.1, 1.1)) 

#Diagram
chordDiagram(
  x = cross_chord_df %>% select(Sub_A, Sub_B, value),
  grid.col = cross_colors,
  transparency = 0.6,
  annotationTrack = "grid",
  preAllocateTracks = list(track.height = 0.07)
)

#Labelling
circos.track(track.index = 1, panel.fun = function(x, y) 
  #Every sub lineage labelled
  {circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
              #Ensure it is readable, not upside down
              facing = "clockwise", 
              niceFacing = TRUE, 
              adj = c(0, 0.5), 
              cex = 1) # size
}, bg.border = NA)

#Title 
title(main = "Cross-Lineage Connections",
      cex.main = 1.5, 
      font = 2,
      line =-3)

#Save with dev.off
dev.off()
#Clear before next use
circos.clear()
```

### Paneling the chord plots together

```{r}

#File path, reading saved images
img_intra <- readPNG(file.path(output_path, "Intra_Lineage_chord.png"))
img_cross <- readPNG(file.path(output_path, "Cross_Lineage_chord.png"))

#Convert to graphical object 
img_intra_plot <- rasterGrob(img_intra)
img_cross_plot  <- rasterGrob(img_cross)

#wrap ana annotate
combined_chord <- wrap_elements(img_intra_plot) + wrap_elements(img_cross_plot) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 30, face = "bold"))

#Saving
ggsave(
  filename = "Combined_Chord_Plots.png",
  plot = combined_chord,
  path = output_path,
  width = 16, 
  height = 8)
```

### Paneling all graphs from this quarto document

```{r}
#Function to load the png
load_png <- function(path) {
  img <- readPNG(path)
  rasterGrob(img, interpolate = TRUE)
}

#Loading the images
p_row1   <- load_png(file.path(output_path, "IP_IP_Panels_Combined.png"))
p_intra  <- load_png(file.path(output_path, "Intra_Lineage_chord.png"))
p_cross  <- load_png(file.path(output_path, "Cross_Lineage_chord.png"))
p_cluster <- load_png(file.path(output_path, "Intra_Cluster.png"))
p_sankey  <- load_png(file.path(output_path, "Cross_Lineage_Sankey_Static.png"))

#Combining + annotate
final_combined <- 
  wrap_elements(p_row1) / 
  (wrap_elements(p_intra) + wrap_elements(p_cross)) / 
  (wrap_elements(p_cluster) + wrap_elements(p_sankey) + plot_layout(widths = c(1.2, 2))) +
  plot_annotation(tag_levels = list(c("", "C", "D", "E", "F"))) & 
  theme(plot.tag = element_text(size = 24, face = "bold"))

#Saving
ggsave(
  filename = "IP_IP_Multi_Final_Panel.png",
  plot = final_combined,
  path = output_path,
  width = 18,
  height = 24
)
```

# 
