---
title: "basic_visualisations"
author: "Sarah Lam"
format: html
editor: visual
---

# Visualising Genetic Correlations

From the SOLAR analysis significant genetic results, we want to obtain IP-IP and Protein-Protein genetic correlations. This could be represented in a network graph.

# 1. Set up

## 1.1 Libraries

```{r, message=FALSE}
library(tidyr)
library(dplyr)
library(tibble)

#Graphs
library(viridis) #Colour blind friendly palette
library(ggplot2)

#Heatmap 
library(pheatmap) #static
library(ComplexHeatmap)


#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
library(tidyverse)

library(circlize)
```

## 1.2 Data Loading

```{r}
#Loading in the solar results analysis filtering.
load("~/Twins_Project/scripts/solar_analysis/solar_analysis_filtering_results.RData")

#Loading in the dataset for .lineage information 
load("~/Twins_Project/data/raw_olink_ip/results/Olink_associations.RData")
```

As a reminder, genetic_0.1_hits, genetic_0.3_hits, environmental_hits and both_hits are FDR filtered 0.05.

gen_env_0.1 and gen_env_0.3 are genetically correlated with environmental correlations.

env_gen is environmentally correlated with genetic correlations.

sum_results_combined contains the FDR filtered plus lineage columns.

## 1.3 Save path

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/visualisations"
```

# 2. IP-Protein Genetics Network graph

Starting with the genetic hits as early instigators have a genetic basis, where they are likely from ageing.

Filtering IPs that share proteins and graphing for visualisations.

Using the already genetically significant filtered genetic_0.1_hits (h2r1 \> 0.1 & h2r2 \> 0.1 & rhoG_FDR \< 0.05), we know the IP-Protein associations are genetically linked.

## 2.1 Matrix

Conversion of the genetic_0.1_hits is necessary from a long to wide format with rows as IPs, columns as proteins and entries as RhoG (pG) . There will be lots of NAs as the majority of IPs do not significnatly associate with 2+ proteins.

Problem is that not all IPs share multiple proteins, 1464 IPs are unique to one protein but the rest (2836) are significantly correlated to 2+ proteins.

```{r}
#Matrix conversion from genetic_0.1_hits
#Using RhoG for the biological connection / proximity as it is already filtered
#Based on shared proteins
g01_long <- genetic_0.1_hits %>%
  select(trait1, trait2, rhoG)%>%
  mutate(
    IP = if_else(grepl("^IMMUNO_",trait1), trait1, trait2),
    protein = if_else(grepl("^IMMUNO_", trait1), trait2, trait1)
  ) %>%
  filter(grepl("^IMMUNO_", IP)) %>% 
  select(IP, protein, rhoG)

IP_protein_mat <- g01_long %>%
  pivot_wider(
    names_from  = protein,
    values_from = rhoG
  ) %>%
  column_to_rownames("IP") %>%
  as.matrix()

#number of non-NA proteins per IP
n_proteins_per_IP <- rowSums(!is.na(IP_protein_mat))

#Filter IPs that have 2+ protein associations, the rest are unique to one protein
#Filtering for 2+ proteins allows us to identify master regulations and identify where one instigator IP controls multiple responder proteins
min_proteins <- 2
keep_IPs <- names(n_proteins_per_IP)[n_proteins_per_IP >= min_proteins]

IP_protein_mat_filt <- IP_protein_mat[keep_IPs,]

```

## 2. Network Graph Genetic_0.1_hits

```{r}
#Building the edges and nodes from the genetic protein hits
gen_edges <- genetic_0.1_hits %>%
  filter(abs(rhoG) > 0.5) %>%  #filtered 0.5 for visuals
  mutate(weight = abs(rhoG)) %>% #abs value strength
  select(from = trait1, to = trait2, weight, rhoG)

#Matrix to proteins as nodes and correlations as edges
g_gen_s <- graph_from_data_frame(gen_edges, directed = FALSE)

#Clustering Louvain used to find modules
cl_gen_s <- cluster_louvain(g_gen_s, weights = E(g_gen_s)$weight)

#Building the graph 
#Tidygraph
tg_gen <- tbl_graph(
  nodes = data.frame(
    name = V(g_gen_s)$name, 
    module = membership(cl_gen_s)  #from the clustering results
  ),
  edges = gen_edges %>%          
    rename(weight = weight),       
  directed = FALSE
)%>%
  activate(nodes) %>% #for proteins not the edges
  mutate(Centrality = centrality_degree()) #Proteins correlated with lots more other proteins -> number of edges connected to each node
#allows to idnetify 'hub' proteins with lots of connections and 'specialist' proteins connected to another or few


#Set colours
n_mods <- length(unique(V(tg_gen)$module))
viridis_colours <- viridis(n_mods)

#Plotting with ggraph as better visually
#fr fruchterman-reingold -> related proteins closer together
ggraph(tg_gen, layout = "fr") +
  #Nodes and edges
  geom_edge_link(aes(width = weight, edge_colour = rhoG),alpha = 0.8) +
  geom_node_point(aes(color = as.factor(module), size = Centrality), alpha = 1)+
  #Colour
scale_edge_colour_gradient(low = "black", high = "grey80") +
  scale_color_manual(values = viridis_colours, name = "Module")+
  geom_node_text(aes(label = name), repel = TRUE, size = 2.5, 
                 max.overlaps = 12) +
  #Aesthetics, layouts
  theme_void() +
  theme(
   plot.background = element_rect(fill = "white", color = NA),
   panel.background = element_rect(fill = "white", color = NA),
   plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
   plot.subtitle = element_text(hjust = 0.5, size = 11),
   #space for legend
   plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
   legend.position = "right",
   legend.box = "vertical",
   legend.text = element_text(size = 6),
   legend.title = element_text(size = 8, face = "bold"))+
  labs(title = "IP-Protein Genetic Correlation Network",
       subtitle = "Based on SOLAR analysis results (H2r > 0.1)",
       edge_color = "rhoG",
       color = "Module") +
  guides(edge_width = "none")
```

Maybe try grouping based on lineage sub.lineages

Other diagrams: circular chord diagram

# 3. IP-IP Genetic Associations

New method: Finding the genetic associations/correlations between IP and IP .

Ideally, I would run IP-IP SOLAR on all IPs to find the genetic correlations but it is too computationally intensive. Instead, using the already genetically filtered genetic_0.1_hits where we know IP-Protein associations were heritable, we can take the selected IPs (4300) then that has been narrowed down to identify only IPs that share 2+ proteins, disregarding those unique to just one, and build IP profiles across proteins. Find IP-IP correlations between rows. However this doesn't mean IP-IP are directly genetically correlated as they could be connected to the protein through different genetic pathways. This can only infer/estimate but not a direct result.

If I use this way to find the genetic associatiosn with IP-IP then wouldn't I also need to apply this logic/step to protein-protein associations as well?

-   Genetic Correlation with **Pearsons**

    -   As we used SOLAR to prove those associations were genetically driven, we can use Pearson to map the observed strength of the associations.

-   Genetic Correlations building on the SOLAR protein-protein network results with **Jaccard similarity**

    -   if IP 1 is associated with protein x and IP 2 is associated with protein Y and the current network shows protein x and y have high pG, then IP 1 and 2 are genetically correlated by proxy.

    -   But this method relies on the SOLAR results which may not be representative?

    -   This would involve (1) creating IP-Protein mapping matrix and 1 if SOLAR associated, 0 if not , (2) using exisiting protein-protein matrix , (3) multiply them ip-protein matrix x protein-protein matrix

    -   I need to find IPs that share the SAME proteins sets.

    -   However, this may be down to sharing the same cell lineage which is a known relationship.

With finding the IP-IP matrix, there is potential to explore the cell lineages.

## 3.1 Pearson's Correlation

Creates a large IP-IP matrix

```{r}
#IP-IP associations through pearsons correlation for genetic similarities based on proteins
#Pearsons based on pG patterns across proteins
IP_IP_cor <- cor(t(IP_protein_mat_filt),
                 use   = "pairwise.complete.obs",
                 method = "pearson")


#Dimensions are 923 so 923 hub IPs -> potential master regulators 
dim(IP_IP_cor)
```

Something is wrong with the matrix as lots of 1s when it should not be

```{r}
#from correlation matrix, create edges
IP_edges <- IP_IP_cor %>%
  as.data.frame() %>%
  rownames_to_column("IP1") %>%
  pivot_longer(-IP1, names_to = "IP2", values_to = "cor") %>%
  filter(IP1 < IP2, cor > 0.5, !is.na(cor))  #Can change threhold and is only positive # 0.25? But with too many pairs to visualise

#changed the correlation to 0.7 and 0.5 was messy
g_IP <- graph_from_data_frame(IP_edges, directed = FALSE)

#Modules
cl_ip <- cluster_louvain(g_IP, weights = IP_edges$cor)

#Tidygraph
tg_ip <- tbl_graph(
  nodes = data.frame(
    name=V(g_IP)$name, 
    module=membership(cl_ip)),
  edges = IP_edges %>% rename(weight = cor), 
  directed=FALSE)

#Plotting with ggraph
ggraph(tg_ip, layout="fr") +
  geom_edge_link(aes(color = factor(to), width=log(weight + 1.1 )),
                 alpha=0.5, 
                 start_cap = circle(3,'mm'),
                 end_cap = circle(3, 'mm')) +
  scale_edge_width(range = c(0.5, 2.5)) +
  geom_node_point(aes(color=factor(module)), size=2, alpha = 0.5) +
  geom_node_text(aes(label=name), repel=TRUE, size=2) +
  theme_void() +
  labs(title="IP Genetic Similarity Modules")
```

hard to visualise

filter by lineage as graph crashes too much memory

## 3.2 Jaccard Similarity

### 3.2.1 Shared Protein Distribution

Displaying the distribution again from the previous qmd

```{r}
#Counting how many share proteins
ip_gen_sharing <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_gen <- table(ip_gen_sharing$protein_count)

shared_protein_ip_gen
```

### 3.2.2 Jaccard Similairty

Calculating the same sharing with the Jaccard similarity index

```{r}
#List of all unique ip-protein pairs 
ip_links <- g01_long %>%
  select(IP, protein) %>%
  distinct()

#Counting how many we have
ip_totals <- ip_links %>%
  count(IP, name = "total_proteins")

#those that are sharing 
ip_sharing <- ip_links %>%
  inner_join(ip_links, by = "protein", relationship = "many-to-many") %>%
  rename(IP_A = IP.x, IP_B = IP.y) %>%
  filter(IP_A < IP_B) %>% #keep different IP pairs with the same protein
  group_by(IP_A, IP_B) %>% #counting how mnay common proteins
  summarise(shared_count = n(), .groups = "drop")

#Jaccard Score caluculations
ip_jaccard <- ip_sharing %>%
  left_join(ip_totals, by = c("IP_A" = "IP")) %>%
  left_join(ip_totals, by = c("IP_B" = "IP")) %>%
  mutate(
    jaccard_score = shared_count / (total_proteins.x + total_proteins.y - shared_count)
  ) %>%
  arrange(desc(jaccard_score))

```

But jaccard index measures overlap not absolute strength

Those that share more proteins than those that only share one matter more for jaccard score.

Thus, weighting and filtering are applied to quantify similarity with more nuance. Weighting so IP pairs with overall more proteins shared and in common are favoured then IP pairs with overall less proteins shared though in common.

```{r}
weight_ip_jaccard <- ip_jaccard %>%
  filter(shared_count > 1) %>% #Only keep pairs sharing 2+ proteins
  mutate(weighted = jaccard_score * shared_count) %>%
  arrange(desc(weighted))
```

### 3.2.3 Lineage Tracing

From the Jaccard similarity results, if they have the same cell lineage then it is a known relationship so they must share the same proteins.

If they derive from different lineages then there is a possibility that this is an instigator as they come from different pathways/parts of the immune system.

#### 3.2.3.1 Filtering by lineage relationships

```{r}
#Need to select Ip and Sub.Lineage 
lineage_ip <- genetic_0.1_hits %>%
  select(IP=trait1, Sup.Lineage, Sub.Lineage) %>%
  distinct()

#joining the lineage 
ip_lineage_check <- weight_ip_jaccard %>%
  left_join(lineage_ip, by = c("IP_A" = "IP")) %>%
  rename(Sup_A = Sup.Lineage, Sub_A = Sub.Lineage) %>%
  
  left_join(lineage_ip, by = c("IP_B" = "IP")) %>%
  rename(Sup_B = Sup.Lineage, Sub_B = Sub.Lineage) %>%
  
  mutate(Relationship = if_else(Sub_A == Sub_B, "Intra-Lineage", "Cross-Lineage")) %>%
  filter(shared_count > 1) %>% #Applying filtering 
  arrange(desc(weighted))

head(ip_lineage_check, 20)
```

All of the IPs that have high scores are of the same lineage (intra-lineage). This is natural as strongest genetic signals are from the same family tree. Lineage driven, structured by immune lineage.

High Jaccard + High weight (9 proteins shared -\> hub of IPs). Maybe multiple pathways triggered.

```{r}
#How many unqiue lineages in cross and intra
cat("Unique Sub.Lineages\n")
ip_lineage_check %>%
  select(Sub_A, Sub_B) %>%
  unlist() %>%
  unique() %>%
  sort() %>%
  print()

cat("\nUnique Sub.Lineages Intra-Lineages\n")
ip_lineage_check %>%
  filter(Relationship == "Intra-Lineage") %>%
  select(Sub_A) %>% # Since A and B are the same in Intra, we only need one
  distinct() %>%
  pull(Sub_A) %>%
  sort() %>%
  print()

cat("\nUnique Sub.Lineages Cross-Lineages\n")
ip_lineage_check %>%
  filter(Relationship == "Cross-Lineage") %>%
  select(Sub_A, Sub_B) %>%
  unlist() %>%
  unique() %>%
  sort() %>%
  print()
```

**NOTE: Could investigate specific sub.lineage instead or the broader sup.lineage**

#### 3.2.3.2 CROSS Lineage

Finding the top cross-lineage IPs

```{r}
cross_lineage <- ip_lineage_check %>%
  filter(Relationship == "Cross-Lineage") %>%
  filter(shared_count > 1) %>% 
  arrange(desc(weighted))

head(cross_lineage, 20)
```

There are a few cross lineage. Top cross have weighted scores of 3.2 which is low (shared proteins 4 out of 5) though these are both T cells

Regardless of weight, checking jaccard score 1.0. These have 3 proteins shared with cross lineages of B cells and T cells. These have high jaccard + low weight -\> specific links, maybe only one pathway triggered.

Specific sub.lineages has highest connection to others? How many sub connects to sub

## 3.3 Graphing

I need to graph (1) intra vs cross lineage realationships (2) Top : shared proteins count, total protein in x and y (3) Jaccard weighted scores?

### 

Focus on cross lineage bridges.

OVERALL GOAL: These top IP-IP pairs, need to compare with which proteins are they sharing

### 3.3. Lineage 

Sankey diagram for sup -\> sub -\> top shared

or

lineage a -\> PROTEIN SHARED -\> lineage B?

### 3.3. Cross lineage chord diagram based 

```{r}
#Using circlize
#Aggregating the data
chord_df <- ip_lineage_check %>%
  group_by(Sub_A, Sub_B, Sup_A, Sup_B, Relationship) %>%
  summarise(value = sum(weighted), .groups = "drop") #can filter for stronger signals with  filter(value > quantile(value, 0.8)) 

#Relationship 
chord_df <- chord_df %>%
  mutate(link_col = if_else(Relationship == "Intra-Lineage", 
                            "#440154FF",  
                            "#21908CFF")) 
#Clearing 
circos.clear()

nodes <- chord_df %>%
  select(Sub = Sub_A, Sup = Sup_A) %>%
  bind_rows(chord_df %>% select(Sub = Sub_B, Sup = Sup_B)) %>%
  distinct() %>%
  arrange(Sup, Sub)

#PLOT
circos.clear()
circos.par(gap.after = 2, start.degree = 90)

chordDiagram(
  x = chord_df %>% select(Sub_A, Sub_B, value),
  order = nodes$Sub,
  col = chord_df$link_col,          
  transparency = 0.3,     #Overlapping
  annotationTrack = "grid", 
  preAllocateTracks = list(track.height = 0.1)
)

#Lables
circos.track(track.index = 1, 
             panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter,
              CELL_META$ylim[1],
              CELL_META$sector.index, 
              facing = "clockwise", 
              niceFacing = TRUE, 
              adj = c(0, 0.5), 
              cex = 0.6)
},
bg.border = NA)

#Legend
legend("bottomleft", title = "Weighted Jaccard Score", 
       legend = c("Low", "High"), 
       fill = c("#440154FF","#21908CFF"), bty = "n")

circos.clear()
```

## jaccard score

```{r}
#Using circlize
#Aggregating the data
chord_df <- ja %>%
  group_by(Sub_A, Sub_B, Sup_A, Sup_B, Relationship) %>%
  summarise(value = sum(weighted), .groups = "drop") #can filter for stronger signals with  filter(value > quantile(value, 0.8)) 

#Relationship 
chord_df <- chord_df %>%
  mutate(link_col = if_else(Relationship == "Intra-Lineage", 
                            "#440154FF",  
                            "#21908CFF")) 
#Clearing 
circos.clear()

nodes <- chord_df %>%
  select(Sub = Sub_A, Sup = Sup_A) %>%
  bind_rows(chord_df %>% select(Sub = Sub_B, Sup = Sup_B)) %>%
  distinct() %>%
  arrange(Sup, Sub)

#PLOT
circos.clear()
circos.par(gap.after = 2, start.degree = 90)

chordDiagram(
  x = chord_df %>% select(Sub_A, Sub_B, value),
  order = nodes$Sub,
  col = chord_df$link_col,          
  transparency = 0.3,     #Overlapping
  annotationTrack = "grid", 
  preAllocateTracks = list(track.height = 0.1)
)

#Lables
circos.track(track.index = 1, 
             panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter,
              CELL_META$ylim[1],
              CELL_META$sector.index, 
              facing = "clockwise", 
              niceFacing = TRUE, 
              adj = c(0, 0.5), 
              cex = 0.6)
},
bg.border = NA)

#Legend
legend("bottomleft", title = "Weighted Jaccard Score", 
       legend = c("Low", "High"), 
       fill = c("#440154FF","#21908CFF"), bty = "n")

circos.clear()
```
