---
title: "basic_visualisations"
author: "Sarah Lam"
format: html
editor: visual
---

# Visualising Genetic Correlations

From the SOLAR analysis significant genetic results, we want to obtain IP-IP and Protein-Protein genetic correlations. This could be represented in a network graph.

# 1. Set up

## 1.1 Libraries

```{r, message=FALSE}
library(tidyr)
library(dplyr)
library(tibble)

#Graphs
library(viridis) #Colour blind friendly palette
library(ggplot2)

#Heatmap 

#library(ComplexHeatmap)


#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
library(tidyverse)

library(circlize)
```

## 1.2 Data Loading

```{r}
#Loading in the solar results analysis filtering.
load("~/Twins_Project/scripts/solar_analysis/solar_analysis_filtering_results.RData")

#Loading in the dataset for .lineage information 
load("~/Twins_Project/data/raw_olink_ip/results/Olink_associations.RData")
```

As a reminder, genetic_0.1_hits, genetic_0.3_hits, environmental_hits and both_hits are FDR filtered 0.05.

gen_env_0.1 and gen_env_0.3 are genetically correlated with environmental correlations.

env_gen is environmentally correlated with genetic correlations.

sum_results_combined contains the FDR filtered plus lineage columns.

## 1.3 Save path

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/visualisations"
```

# 2. IP-Protein Genetics Network graph

Starting with the genetic hits as early instigators have a genetic basis, where they are likely from ageing.

Filtering IPs that share proteins and graphing for visualisations.

Using the already genetically significant filtered genetic_0.1_hits (h2r1 \> 0.1 & h2r2 \> 0.1 & rhoG_FDR \< 0.05), we know the IP-Protein associations are genetically linked.

## 2.1 Matrix

Conversion of the genetic_0.1_hits is necessary from a long to wide format with rows as IPs, columns as proteins and entries as RhoG (pG) . There will be lots of NAs as the majority of IPs do not significnatly associate with 2+ proteins.

Problem is that not all IPs share multiple proteins, 1464 IPs are unique to one protein but the rest (2836) are significantly correlated to 2+ proteins.

```{r}
#Matrix conversion from genetic_0.1_hits
#Using RhoG for the biological connection / proximity as it is already filtered
#Based on shared proteins
g01_long <- genetic_0.1_hits %>%
  select(trait1, trait2, rhoG)%>%
  mutate(
    IP = if_else(grepl("^IMMUNO_",trait1), trait1, trait2),
    protein = if_else(grepl("^IMMUNO_", trait1), trait2, trait1)
  ) %>%
  filter(grepl("^IMMUNO_", IP)) %>% 
  select(IP, protein, rhoG)

IP_protein_mat <- g01_long %>%
  pivot_wider(
    names_from  = protein,
    values_from = rhoG
  ) %>%
  column_to_rownames("IP") %>%
  as.matrix()

#number of non-NA proteins per IP
n_proteins_per_IP <- rowSums(!is.na(IP_protein_mat))

#Filter IPs that have 2+ protein associations, the rest are unique to one protein
#Filtering for 2+ proteins allows us to identify master regulations and identify where one instigator IP controls multiple responder proteins
min_proteins <- 2
keep_IPs <- names(n_proteins_per_IP)[n_proteins_per_IP >= min_proteins]

IP_protein_mat_filt <- IP_protein_mat[keep_IPs,]

```

## 2. Genetic_0.1_hits

### Network Graph

```{r}
#Building the edges and nodes from the genetic protein hits
gen_edges <- genetic_0.1_hits %>%
  filter(abs(rhoG) > 0.5) %>%  #filtered 0.5 for visuals
  mutate(weight = abs(rhoG)) %>% #abs value strength
  select(from = trait1, to = trait2, weight, rhoG)

#Matrix to proteins as nodes and correlations as edges
g_gen_s <- graph_from_data_frame(gen_edges, directed = FALSE)

#Clustering Louvain used to find modules
cl_gen_s <- cluster_louvain(g_gen_s, weights = E(g_gen_s)$weight)

#Building the graph 
#Tidygraph
tg_gen <- tbl_graph(
  nodes = data.frame(
    name = V(g_gen_s)$name, 
    module = membership(cl_gen_s)  #from the clustering results
  ),
  edges = gen_edges %>%          
    rename(weight = weight),       
  directed = FALSE
)%>%
  activate(nodes) %>% #for proteins not the edges
  mutate(Centrality = centrality_degree()) #Proteins correlated with lots more other proteins -> number of edges connected to each node
#allows to idnetify 'hub' proteins with lots of connections and 'specialist' proteins connected to another or few


#Set colours
n_mods <- length(unique(V(tg_gen)$module))
viridis_colours <- viridis(n_mods)

#Plotting with ggraph as better visually
#fr fruchterman-reingold -> related proteins closer together
ggraph(tg_gen, layout = "fr") +
  #Nodes and edges
  geom_edge_link(aes(width = weight, edge_colour = rhoG),alpha = 0.8) +
  geom_node_point(aes(color = as.factor(module), size = Centrality), alpha = 1)+
  #Colour
scale_edge_colour_gradient(low = "black", high = "grey80") +
  scale_color_manual(values = viridis_colours, name = "Module")+
  geom_node_text(aes(label = name), repel = TRUE, size = 2.5, 
                 max.overlaps = 12) +
  #Aesthetics, layouts
  theme_void() +
  theme(
   plot.background = element_rect(fill = "white", color = NA),
   panel.background = element_rect(fill = "white", color = NA),
   plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
   plot.subtitle = element_text(hjust = 0.5, size = 11),
   #space for legend
   plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
   legend.position = "right",
   legend.box = "vertical",
   legend.text = element_text(size = 6),
   legend.title = element_text(size = 8, face = "bold"))+
  labs(title = "IP-Protein Genetic Correlation Network",
       subtitle = "Based on SOLAR analysis results (H²r > 0.1)",
       edge_color = "rhoG",
       color = "Module") +
  guides(edge_width = "none")

```
