---
title: "protein_protein_analysis"
format: html
editor: visual
---

# Protein Protein Genetic Correlations

With the SOLAR analysis results performed for all proteins protein associations, identify the genetic and environmental correlations. The SOLAR ran for protein-protein associations used the twins to identify whether they share the same proteins

## 1. Set Up

### 1.1 Load Libraries

```{r, message = FALSE}
library(tidyr)
library(dplyr)

#graphing
library(viridis) #Colour blind friendly palette
library(ComplexHeatmap)

#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
library(ggforce) #Circular network modules

```

### 1.2 Load data

Loading the collated csv from the solar analysis results

```{r}
prot_data <- read.delim("~/Twins_Project/data/protein_protein_solar/prot_summary_heritability_COMPLETE.csv")
```

### 1.3 Save path

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/visualisations"
```

### 1.4 Set Random Seed

```{r}
set.seed(123)
```

## 2. Filtering

Using FDR filtering

```{r}
#Use FDR adjustment
prot_data_fdr <- prot_data %>%
  mutate(
    rhoG_FDR = p.adjust(rhoG_p0, method = "fdr"),
    rhoE_FDR = p.adjust(rhoE_p, method = "fdr")
  )

#Filter FDR <0.05 for genetic hits
prot_gen_0.1_hits <- prot_data_fdr %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05)
```

I only filtered for genetic hits. Perhaps can do environmental and both.

Can find more about the top pairs and correlations.

## 3. Saving the output environment

Saving to be loaded elsewhere.

```{r}
save.image(file = "prot_prot_filtering_results.RData")
```

## 4. Graphing

#### 4.1 Network Graphs

PPI network. Relationship between proteins. Identify clusters/modules and key proteins.

```{r}
#Building the edges and nodes from the genetic protein hits
prot_edges <- prot_gen_0.1_hits %>%
  filter(abs(rhoG) > 0.5) %>%  #filtered 0.5 for visuals
  mutate(weight = abs(rhoG)) %>% #abs value strength
  select(from = trait1, to = trait2, weight, rhoG)

#Matrix to proteins as nodes and correlations as edges
g_prot_s <- graph_from_data_frame(prot_edges, directed = FALSE)

#Clustering Louvain used to find modules
cl_prot_s <- cluster_louvain(g_prot_s, weights = E(g_prot_s)$weight)

#Building the graph 
#Tidygraph
tg_prot <- tbl_graph(
  nodes = data.frame(
    name = V(g_prot_s)$name, 
    module = membership(cl_prot_s)  #from the clustering results
  ),
  edges = prot_edges %>%          
    rename(weight = weight),       
  directed = FALSE
)%>%
  activate(nodes) %>% #for proteins not the edges
  mutate(Centrality = centrality_degree()) #Proteins correlated with lots more other proteins -> number of edges connected to each node
#allows to idnetify 'hub' proteins with lots of connections and 'specialist' proteins connected to another or few


#Set colours
n_mods <- length(unique(V(tg_prot)$module))
viridis_colours <- viridis(n_mods)

#Plotting with ggraph as better visually
#fr fruchterman-reingold -> related proteins closer together
ggraph(tg_prot, layout = "fr") +
  #Nodes and edges
  geom_edge_link(aes(width = weight, edge_colour = rhoG),alpha = 0.8) +
  geom_node_point(aes(color = as.factor(module), size = Centrality), alpha = 1)+
  #Colour
scale_edge_colour_gradient(low = "black", high = "grey80") +
  scale_color_manual(values = viridis_colours, name = "Module")+
  geom_node_text(aes(label = name), repel = TRUE, size = 2.5, 
                 max.overlaps = 12) +
  #Aesthetics, layouts
  theme_void() +
  theme(
   plot.background = element_rect(fill = "white", color = NA),
   panel.background = element_rect(fill = "white", color = NA),
   plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
   plot.subtitle = element_text(hjust = 0.5, size = 11),
   #space for legend
   plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
   legend.position = "right",
   legend.box = "vertical",
   legend.text = element_text(size = 6),
   legend.title = element_text(size = 8, face = "bold"))+
  labs(title = "Protein-Protein Genetic Correlation Network",
       subtitle = "Based on SOLAR analysis results (H2r > 0.1)",
       edge_color = "rhoG",
       color = "Module") +
  guides(edge_width = "none")

#Saving plot
ggsave(
  filename = "Protein_protein_solar_networkgraph.png", 
  path = output_path,
  width = 10,     
  height = 6
)
```

The lines/edges represent genetic corerelation RhoG (pG) -\> shared genetic influence / relationship between two proteins.

The size of the points represent centrality which is how many connections/neighbours a proteins has -\> how integrated the protein is in the network.

Colours represent modules -\> how groups of proteins are highly correlated with each other.

SCF and CCL20 are the only protein pair to have a negative rhoG. When searching this up, seems like research supports this that they are environmentally correlated in inflammation.

-   filtering of the genetic and those that are environmetally?

-   instead of representing the colours, add in circles

```{r}

n_mods <- length(unique(V(tg_prot)$module))
viridis_colours <- viridis(n_mods)

#Graphing
ggraph(tg_prot, layout = "fr") +
  geom_mark_hull(aes(x = x, y = y,fill = as.factor(module)), 
                 concavity = 5, #Hulls  - curves
                 radius = unit(0.3, "cm"),  #Smoothing
                 alpha = 0.7, 
                 size = 0.05,
                 expand = unit(0.5, "cm"))+
  #Edges
  geom_edge_link(aes(width = weight, edge_colour = rhoG), alpha = 0.9) +
  #nodes
  geom_node_point(aes(size = Centrality), color = "grey20", alpha = 0.9) +
  
  #Aesthetics
  scale_edge_colour_gradient(low = "black", high = "grey90") +
  #scale_fill_manual(values = viridis_colours, guide = "none") + 
  scale_fill_viridis_d(name = "Module Group") +
  geom_node_text(aes(label = name), repel = TRUE, size = 2, max.overlaps = 10) +
  theme_void() +
    theme(
   plot.background = element_rect(fill = "white", color = NA),
   panel.background = element_rect(fill = "white", color = NA),
   plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
   plot.subtitle = element_text(hjust = 0.5, size = 11),
   #space for legend
   plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
   legend.position = "right",
   legend.box = "vertical",
   legend.text = element_text(size = 7),
   legend.title = element_text(size = 9, face = "bold"))+
  labs(
    title = "Protein-Protein Genetic Correlation Network",
    subtitle = "Based on SOLAR analysis results (H2r > 0.1)",
    edge_colour = "rhoG",
    color = "Module Group",
    size = "Centrality"
  ) +
  guides(edge_width = "none")

#Saving plot
ggsave(
  filename = "Protein_protein_solar_circ_networkgraph.png", 
  path = output_path,
  width = 10,     
  height = 6
)

```

Size of points = centrality so module 5 appears to have the largest hub/connection in the network followed by module 4. Though module 1 contains CXCL10 and CXCL9 but they don't have as many connections.

If I make this interactive, can adjust the filtering of rhoG \>0.5 but will risk overcrowded visualisations.

# Pearsons

If i was to do pearsons....

This is based on sharing IPS, not from a separate solar analysis.

```{r}
#Loading in the solar results analysis filtering.
load("~/Twins_Project/scripts/solar_analysis/solar_analysis_filtering_results.RData")

```

## Matrix

Conversion of the genetic_0.1_hits is necessary from a long to wide format with rows as IPs, columns as proteins and entries as RhoG (pG) . There will be lots of NAs as the majority of IPs do not significnatly associate with 2+ proteins.

Problem is that not all IPs share multiple proteins, 1464 IPs are unique to one protein but the rest (2836) are significantly correlated to 2+ proteins.

```{r}
#Matrix conversion from genetic_0.1_hits
#Using RhoG for the biological connection / proximity as it is already filtered
#Based on shared proteins
g01_long <- genetic_0.1_hits %>%
  select(trait1, trait2, rhoG)%>%
  mutate(
    IP = if_else(grepl("^IMMUNO_",trait1), trait1, trait2),
    protein = if_else(grepl("^IMMUNO_", trait1), trait2, trait1)
  ) %>%
  filter(grepl("^IMMUNO_", IP)) %>% 
  select(IP, protein, rhoG)

IP_protein_mat <- g01_long %>%
  pivot_wider(
    names_from  = protein,
    values_from = rhoG
  ) %>%
  column_to_rownames("IP") %>%
  as.matrix()

#number of non-NA proteins per IP
n_proteins_per_IP <- rowSums(!is.na(IP_protein_mat))

#Filter IPs that have 2+ protein associations, the rest are unique to one protein
#Filtering for 2+ proteins allows us to identify master regulations and identify where one instigator IP controls multiple responder proteins
min_proteins <- 2
keep_IPs <- names(n_proteins_per_IP)[n_proteins_per_IP >= min_proteins]

IP_protein_mat_filt <- IP_protein_mat[keep_IPs,]
```

## Pearsons

```{r}
protein_protein_cor <- cor(IP_protein_mat_filt ,
                           use = "pairwise.complete.obs",
                           method = "pearson")
dim(protein_protein_cor)

```

## Network graph

```{r}
protein_edges <- protein_protein_cor %>%
  as.data.frame() %>%
  rownames_to_column("protein1") %>%
  pivot_longer(-protein1, names_to = "protein2", values_to = "cor") %>%
  filter(protein1 < protein2,  cor > 0.25, !is.na(cor))

g_prot <- graph_from_data_frame(protein_edges, directed = FALSE)

#Modules
cl_prot <- cluster_louvain(g_prot, weights = protein_edges$cor)

#Tidygraph
tg_prot <- tbl_graph(
  nodes = data.frame(
    name=V(g_prot)$name, 
    module=membership(cl_prot)),
  edges = protein_edges %>% rename(weight = cor), 
  directed=FALSE)%>%
  mutate(importance = centrality_degree())

#Plotting with ggraph
ggraph(tg_prot, layout="fr") +
 geom_edge_link(aes(width=log(weight + 1)), color = "grey80", alpha=0.5) +
  geom_node_point(aes(color=factor(module), size = importance), alpha = 0.8) +
  scale_color_viridis_d(option = "viridis") + 
  scale_size_continuous(range = c(2, 7)) +
  geom_node_text(aes(label=name), repel=TRUE, size=2) +
  theme_void() +
  labs(title="Protein Genetic Similarity Modules")
```
