---
title: "protein_protein_analysis"
format: html
editor: visual
---

# Protein Protein Genetic Correlations

With the SOLAR analysis results performed for all proteins protein associations, identify the genetic and environmental correlations. The SOLAR ran for protein-protein associations used the twins to identify whether they share the same proteins.

## 1. Set Up

### 1.1 Load Libraries

```{r, message = FALSE}
library(tidyr)
library(dplyr)
```

### 1.2 Load data

Loading the collated csv from the solar analysis results

```{r}
prot_data <- read.delim("~/Twins_Project/data/protein_protein_solar/prot_summary_heritability_COMPLETE.csv")
```

### 1.3 Save path

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/visualisations/protein_protein"
```

### 1.4 Set Random Seed

As we are going to generate clusters, set random seed/

```{r}
set.seed(123)
```

## 2. Filtering

Using FDR filtering. Filter for both genetic and environmental hits.

```{r}
#Use FDR adjustment
prot_data_fdr <- prot_data %>%
  mutate(
    rhoG_FDR = p.adjust(rhoG_p0, method = "fdr"),
    rhoE_FDR = p.adjust(rhoE_p, method = "fdr")
  )

#Filter FDR <0.05 for genetic hits
prot_gen_0.1_hits <- prot_data_fdr %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05)

#Filter FDR <0.05 for environrmntal hits
prot_env_hits<- prot_data_fdr  %>%
  filter( rhoE_FDR < 0.05)
```

## 3. Saving the output environment

Saving to be loaded elsewhere.

```{r}
save.image(file = "prot_prot_filtering_results.RData")
```

## 4. Investigations

### Genetic

```{r}
prot_gen_0.1_hits  %>%
  summarise (
    mean_h2r1 = mean(h2r1),
    sd_h2r1   = sd(h2r1),
    max_h2r1 = max(h2r1),
    mean_h2r2 = mean(h2r2),
    sd_h2r2= sd(h2r2),
    max_h2r2 = max(h2r2),
    mean_rhoG = mean(abs(rhoG)),
    sd_rhoG= sd(abs(rhoG)),
    mean_rhoE = mean(abs(rhoE)),
    sd_rhoE= sd(abs(rhoE))
  )

```

Obtaining heritability of proteins overall. As proteins are in both trait columns, there is a need to create a new df.

```{r}
#Combining all unique proteins
unique_heritability_gen <- prot_gen_0.1_hits %>%
  select(trait1, h2r1) %>%
  rename(protein = trait1, h2 = h2r1) %>%
  bind_rows(
    prot_gen_0.1_hits %>% 
    select(trait2, h2r2) %>% 
    rename(protein = trait2, h2 = h2r2)
  ) %>%
  distinct(protein, .keep_all = TRUE) #Duplicate removal

#Calculating stats
genfinal_h2_summary <- unique_heritability_gen %>%
  summarise(
    mean_h2 = mean(h2),
    sd_h2 = sd(h2),
    max_h2 = max(h2)
  )

genfinal_h2_summary
```

### Environmental

```{r}
prot_env_hits  %>%
  summarise (
    mean_h2r1 = mean(h2r1),
    sd_h2r1   = sd(h2r1),
    max_h2r1 = max(h2r1),
    mean_h2r2 = mean(h2r2),
    sd_h2r2= sd(h2r2),
    max_h2r2 = max(h2r2),
    mean_rhoG = mean(abs(rhoG)),
    sd_rhoG= sd(abs(rhoG)),
    mean_rhoE = mean(abs(rhoE)),
    sd_rhoE= sd(abs(rhoE))
  )

```

```{r}
#Combining all unique proteins
unique_heritability_env <- prot_env_hits %>%
  select(trait1, h2r1) %>%
  rename(protein = trait1, h2 = h2r1) %>%
  bind_rows(
    prot_gen_0.1_hits %>% 
    select(trait2, h2r2) %>% 
    rename(protein = trait2, h2 = h2r2)
  ) %>%
  distinct(protein, .keep_all = TRUE) #Duplicate removal

#Calculating stats
envfinal_h2_summary <- unique_heritability_env %>%
  summarise(
    mean_h2 = mean(h2),
    sd_h2 = sd(h2),
    max_h2 = max(h2)
  )

envfinal_h2_summary
```
