---
title: "protein_protein_analysis"
format: html
editor: visual
---

# Protein Protein Genetic Correlations

With the SOLAR analysis results performed for all proteins protein associations, identify the genetic and environmental correlations. The SOLAR ran for protein-protein associations used the twins to identify whether they share the same proteins

## 1. Set Up

### 1.1 Load Libraries

```{r, message = FALSE}
library(tidyr)
library(dplyr)

#graphing
library(viridis) #Colour blind friendly palette
library(ComplexHeatmap)

#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
library(ggforce) #Circular network modules

```

### 1.2 Load data

Loading the collated csv from the solar analysis results

```{r}
prot_data <- read.delim("~/Twins_Project/data/protein_protein_solar/prot_summary_heritability_COMPLETE.csv")
```

### 1.3 Save path

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/visualisations/protein_protein"
```

### 1.4 Set Random Seed

As we are going to generate clusters, set random seed/

```{r}
set.seed(123)
```

## 2. Filtering

Using FDR filtering

```{r}
#Use FDR adjustment
prot_data_fdr <- prot_data %>%
  mutate(
    rhoG_FDR = p.adjust(rhoG_p0, method = "fdr"),
    rhoE_FDR = p.adjust(rhoE_p, method = "fdr")
  )

#Filter FDR <0.05 for genetic hits
prot_gen_0.1_hits <- prot_data_fdr %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05)
```

I only filtered for genetic hits. Perhaps can do environmental and both.

Can find more about the top pairs and correlations.

## 3. Saving the output environment

Saving to be loaded elsewhere.

```{r}
save.image(file = "prot_prot_filtering_results.RData")
```

## 4. Investigations

```{r}
prot_gen_0.1_hits  %>%
  summarise (
    mean_h2r1 = mean(h2r1),
    sd_h2r1   = sd(h2r1),
    max_h2r1 = max(h2r1),
    mean_h2r2 = mean(h2r2),
    sd_h2r2= sd(h2r2),
    max_h2r2 = max(h2r2),
    mean_rhoG = mean(abs(rhoG)),
    sd_rhoG= sd(abs(rhoG)),
    mean_rhoE = mean(abs(rhoE)),
    sd_rhoE= sd(abs(rhoE))
  )

```

Obtaining heritability of proteins overall. As proteins are in both trait columns, there is a need to create a new df.

```{r}
#Combining all unique proteins
unique_heritability <- prot_gen_0.1_hits %>%
  select(trait1, h2r1) %>%
  rename(protein = trait1, h2 = h2r1) %>%
  bind_rows(
    prot_gen_0.1_hits %>% 
    select(trait2, h2r2) %>% 
    rename(protein = trait2, h2 = h2r2)
  ) %>%
  distinct(protein, .keep_all = TRUE) #Duplicate removal

#Calculating stats
final_h2_summary <- unique_heritability %>%
  summarise(
    mean_h2 = mean(h2),
    sd_h2 = sd(h2),
    max_h2 = max(h2)
  )

final_h2_summary
```

## 5. Graphing

#### 5.1 Network Graphs

PPI network. Relationship between proteins. Identify clusters/modules and key proteins.

```{r}
#Building the edges and nodes from the genetic protein hits
prot_edges <- prot_gen_0.1_hits %>%
  filter(abs(rhoG) > 0.5) %>%  #filtered 0.5 for visuals
  mutate(weight = abs(rhoG)) %>% #abs value strength
  select(from = trait1, to = trait2, weight, rhoG)

#Matrix to proteins as nodes and correlations as edges
g_prot_s <- graph_from_data_frame(prot_edges, directed = FALSE)

#Clustering Louvain used to find modules
cl_prot_s <- cluster_louvain(g_prot_s, weights = E(g_prot_s)$weight)


#Building the graph 
#Tidygraph
tg_prot <- tbl_graph(
  nodes = data.frame(
    name = V(g_prot_s)$name, 
    module = membership(cl_prot_s)  #from the clustering results
  ),
  edges = prot_edges %>%          
    rename(weight = weight),       
  directed = FALSE
)%>%
  activate(nodes) %>% #for proteins not the edges
  mutate(Centrality = centrality_degree()) #Proteins correlated with lots more other proteins -> number of edges connected to each node
#allows to idnetify 'hub' proteins with lots of connections and 'specialist' proteins connected to another or few


#Set colours
n_mods <- length(unique(V(tg_prot)$module))
viridis_colours <- viridis(n_mods)

#Plotting with ggraph as better visually
#fr fruchterman-reingold -> related proteins closer together
ggraph(tg_prot, layout = "fr") +
  #Nodes and edges
  geom_edge_link(aes(width = weight, edge_colour = rhoG),alpha = 0.8) +
  geom_node_point(aes(color = as.factor(module), size = Centrality), alpha = 1)+
  #Colour
scale_edge_colour_gradient(low = "black", high = "grey80") +
  scale_color_manual(values = viridis_colours, name = "Module")+
  geom_node_text(aes(label = name), repel = TRUE, size = 2.5, 
                 max.overlaps = 12) +
  #Aesthetics, layouts
  theme_void() +
  theme(
   plot.background = element_rect(fill = "white", color = NA),
   panel.background = element_rect(fill = "white", color = NA),
   plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
   plot.subtitle = element_text(hjust = 0.5, size = 11),
   #space for legend
   plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
   legend.position = "right",
   legend.box = "vertical",
   legend.text = element_text(size = 6),
   legend.title = element_text(size = 8, face = "bold"))+
  labs(title = "Protein-Protein Genetic Correlation Network",
       subtitle = "Based on SOLAR analysis results (H² > 0.1)",
       edge_color = "rhoG",
       color = "Module") +
  guides(edge_width = "none")

#Saving plot
ggsave(
  filename = "Prot_Network_Graph.png", 
  path = output_path,
  width = 10,     
  height = 6
)
```

The lines/edges represent genetic corerelation RhoG (pG) -\> shared genetic influence / relationship between two proteins.

The size of the points represent centrality which is how many connections/neighbours a proteins has -\> how integrated the protein is in the network.

Colours represent modules -\> how groups of proteins are highly correlated with each other.

SCF and CCL20 are the only protein pair to have a negative rhoG. When searching this up, seems like research supports this that they are environmentally correlated in inflammation.

-   filtering of the genetic and those that are environmetally?

-   instead of representing the colours, add in circles

```{r}

n_mods <- length(unique(V(tg_prot)$module))
viridis_colours <- viridis(n_mods)

#Graphing
ggraph(tg_prot, layout = "fr") +
  geom_mark_hull(aes(x = x, y = y,fill = as.factor(module)), 
                 concavity = 5, #Hulls  - curves
                 radius = unit(0.3, "cm"),  #Smoothing
                 alpha = 0.8, 
                 size = 0.05,
                 expand = unit(0.5, "cm"))+
  #Edges
  geom_edge_link(aes(width = weight, 
                     linetype = ifelse(rhoG < 0, "Negative", "Positive")), 
                 edge_colour = "grey50", 
                 alpha = 0.8) +
  #Negative and positive different line types 
  scale_edge_linetype_manual(values = c("Negative" = "dashed", "Positive" = "solid"),
                             name = "Correlation Type") +
  #Thickness of line is strength 
  scale_edge_width(range = c(1, 5), name = "ρG Correlation Strength") +
  #nodes
  geom_node_point(aes(size = Centrality), color = "grey20", alpha = 0.9) +
  
  #Aesthetics
  scale_fill_viridis_d(name = "Module Group") +
  
  #Label
  geom_node_text(aes(label = name), repel = TRUE, size = 3, fontface = "bold",force = 2,point.padding = unit(0.3, "lines"),max.overlaps = Inf) +
  
  theme_void() +
    theme(
   plot.background = element_rect(fill = "white", color = NA),
   panel.background = element_rect(fill = "white", color = NA),
   plot.title = element_text(hjust = 0.5,face = "bold", size = 16),
   #space for legend
   plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
   legend.position = "right",
   legend.box = "vertical",
   legend.text = element_text(size = 10),
   legend.title = element_text(size = 12, face = "bold"))+
  labs(
    title = "Protein-Protein Genetic Correlation Network",
    #edge_colour = "Genetic Strength (ρG)",
    color = "Module Group",
    size = "Centrality"
  ) +
  guides(guides(edge_width = guide_legend(title = "ρG Correlation Strength")))

#Saving plot
ggsave(
  filename = "Prot_Circ_Network_Graph.png", 
  path = output_path,
  width = 14,     
  height = 8
)

```

Size of points = centrality so module 5 appears to have the largest hub/connection in the network followed by module 4. Though module 1 contains CXCL10 and CXCL9 but they don't have as many connections.

If I make this interactive, can adjust the filtering of rhoG \>0.5 but will risk overcrowded visualisations.
