---
title: "solar_analysis_filtering"
format: html
editor: visual
---

# SOLAR Results Analysis

After obtaining the IP-Protein associations from the SOLAR analysis. Exploration and filtering is needed for interpretation.

## 1. Set up

### 1.1 Load libraries

```{r, message=FALSE}
library(stringr)
library(dplyr)
library(tidyr)
library(tibble)

#Plotting
library(viridis) #Colour blind friendly palette
library(ggplot2) #Graphing
library(ggrepel) #Volcano
library(patchwork) #Combine Plots

#Heatmap
library(ComplexHeatmap)
library(circlize)
```

### 1.2 Loading Data

```{r}
#Importing the summary results
summary_results<- read.delim("~/Twins_Project/data/solar_analysis/summary_heritability.csv")

#Loading in the dataset for sup.lineage information 
load("~/Twins_Project/data/raw_olink_ip/results/Olink_associations.RData")
```

### 1.3 Save path

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/visualisations"
```

## 2. Filtering of Significant IP-Protein Associations

### 2.1 FDR Adjusted Filtering

Adjusting with FDR to find those significant based on FDR \< 0.05.

```{r}
#Adding in FDR of rhoG and rhoE
summary_results_fdr <- summary_results %>%
  mutate(
    rhoG_FDR = p.adjust(rhoG_p0, method = "fdr"),
    rhoE_FDR = p.adjust(rhoE_p, method = "fdr")
  )

```

### 2.2 Adding in Lineages from infoF dataframe

Using Sup.Lineage as it contains T cells, NK cells, monocytes, dendritic cells, B cells to determine the IP lineages.

```{r}
#Need to clean the IMMUNO_X.new names to match summary_result_fdr as it doesnt have .new
infoF_clean <- infoF %>%
  mutate(SET.NAME_clean = gsub("\\.new$", "", SET.NAME)) %>%
  distinct(SET.NAME_clean, .keep_all = TRUE)

#Combining the lineage columns to the summary results dataframe
sum_results_combined <- summary_results_fdr %>%
  left_join(infoF_clean %>% select(SET.NAME_clean, Lineage, Sup.Lineage, Sub.Lineage), 
            by = c("trait1" = "SET.NAME_clean"))
```

### 2.3 Filtering based on Genetic, Environment or both associations based on the SOLAR outputs

Using the heritability h2r results with rhoG, we can determine genetic hits. For environmental hits, h2r does not need to be significant.

#### 2.3.1 Genetic Hits

```{r}
#Experimenting with different h2r thresholds
genetic_0.1_hits <- sum_results_combined %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05)
#4300

genetic_0.3_hits <- sum_results_combined %>%
  filter(h2r1 > 0.3 & h2r2 > 0.3 & rhoG_FDR < 0.05)
#4047

genetic_0.5_hits <- sum_results_combined %>%
  filter(h2r1 > 0.5 & h2r2 > 0.5 & rhoG_FDR < 0.05)
#1068

```

#### 2.3.2 Environmental hits

Only accounting for rhoE as h2r can be low or high

environmental_hits \<- sum_results_combined %\>%

filter(rhoE_FDR \< 0.05)

```{r}
#Only accounting for rhoE as h2r can be low or high 
environmental_hits <- sum_results_combined %>%
  filter(rhoE_FDR < 0.05)
```

#### 2.3.3 Both hits

```{r}
#Significant in both genetic and environmental correlations
both_hits <- sum_results_combined %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05 & rhoE_FDR < 0.05)
```

### 2.2 Further Filtering

Further filtering of genetically correlated pairs with environmental associations and vise versa - environmentally correlated pairs with genetic associations.

#### 2.2.1 Genetics -\> Environmental

```{r}
#Genetics -> how many are also environmental 
gen_env_0.1 <- genetic_0.1_hits %>%
  filter(rhoE_p < 0.05)

gen_env_0.3 <- genetic_0.3_hits %>%
  filter(rhoE_p < 0.05)
#same 130 pairs 

gen_env_0.5 <- genetic_0.5_hits %>%
  filter(rhoE_p < 0.05)
#only 1 IMMUNO_41521 TNFB
```

#### 2.2.2 Environmental -\> Genetics

```{r}
#Environmental -> how many are genetics
env_gen <- environmental_hits%>%
  filter(rhoG_p0<0.05)
```

### 2.3 Saving the output environment

Saving to be loaded elsewhere.

```{r}
save.image(file = "solar_analysis_filtering_results.RData")
```

## 3. Exploration of the overall IP-Protein Results (non-filtered)

sum_results_combined exploration. This is the general counts and distribution.

### 3.1 Protein Distribution

Analysing how many proteins per IPs are present in each filtering.

#### 3.1.1 Overall Protein Distribution

```{r}
sum_results_combined %>%
  summarise(
    total_rows = n(),
    unique_IPs = n_distinct(trait1),
    unique_proteins = n_distinct(trait2)
  )
```

#### 3.1.2 IP to Protein Frequency Distribution

```{r}
ip_per_protein <- sum_results_combined %>%
  count(trait2, sort = TRUE, name = "n_IPs")
print(ip_per_protein)
```

```{r}
ip_per_protein %>%
  ggplot(aes(x = reorder(trait2, n_IPs), y = n_IPs)) +
  geom_col(aes(fill = n_IPs), alpha = 0.8, color = "white", linewidth = 0.3) +
  #Log for better visualisation
 scale_y_continuous(trans = "pseudo_log", 
                   breaks = c(1, 10, 100, 1000, 2000,4000)) +
  #Numerical values next to bar
  geom_text(aes(label = n_IPs), hjust = -0.2, size =1.5, color = "black", fontface = "bold")+
  #Colour scheme
  scale_fill_viridis_c(option = "viridis")+
  #Axis
  coord_flip() +
  #Labelling
  labs(title = "Overall Distribution of IPs per Protein",
       x = "Protein", y = "Number of IPs ") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 6),     
    axis.title.y = element_text(size = 10),
    axis.text.x = element_text(size = 6),
    axis.title.x = element_text(size = 10),
    plot.title = element_text(size = 12, face = "bold",hjust = 0.5),
    legend.position = "none",
    panel.grid.major.y = element_blank() #No vertical lines
  )

#Saving plot
ggsave(
  filename = "Overall_Protein_Distribution.png", 
  path = output_path,
  width = 10,     
  height = 8
)
```

In the 0.1 and 0.3 top hits, find what proteins are most frequent/popular for genetic. Observing whether there are changes in protein frequency patterns.

45 proteins

#### 3.1.3 Top 15 Proteins

Better Visualisation

```{r}
ip_per_protein %>%
  slice_max(n_IPs, n = 15, with_ties = FALSE) %>%
  ggplot(aes(x = reorder(trait2, n_IPs), y = n_IPs)) +
  geom_col(aes(fill = n_IPs), alpha = 0.8, color = "white", linewidth = 0.3) +

  #Numerical values next to bar
  geom_text(aes(label = n_IPs), hjust = -0.2, size =3, color = "black", fontface = "bold")+
  #Colour scheme
  scale_fill_viridis_c(option = "viridis")+
  #Axis
  coord_flip() +
  scale_y_continuous(
    limits = c(0, 3800),               
    breaks = seq(0, 4000, by = 500),
    expand = expansion(mult = c(0, 0.1)))+
  #Labelling
  labs(title = "Top 15 Protein by IP Distribution ",
       x = "Protein", y = "Number of IPs ") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),     
    axis.title.y = element_text(size = 12,face = "bold"),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12,face = "bold"),
    plot.title = element_text(size = 14, face = "bold",hjust = 0.5, margin = margin(b = 10)),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    panel.grid.major.y = element_blank() #No vertical lines
  )

#Saving plot
ggsave(
  filename = "Top_15_Protein_Distribution.png", 
  path = output_path,
  width = 10,     
  height = 8
)
```

### 3.2 IP sharing count

#### 3.2.1 Distribution

```{r}
protein_per_ip <- sum_results_combined %>%
  count(trait1, sort = TRUE, name = "n_Proteins")

print(protein_per_ip )
```

#### 3.2.2 Lineage Investigation

```{r}
lineage_summary <- sum_results_combined %>%
  group_by(Lineage, Sup.Lineage, Sub.Lineage) %>%
  summarise(
    n_unique_IPs = n_distinct(trait1),
    n_unique_Proteins = n_distinct(trait2),
    n_Pairs = n() 
  ) %>%
  arrange(desc(n_Pairs))

print(lineage_summary)
```

27 unique lineages across pairs.

## 4. IP-Protein Post-Filtering Summary Statistics

Show the distribution of those genetically, environmentally and both significant.

### 4.1 Graphing Correlation Landscape

Visualising the distribution of genetic and environmental correlations.

```{r}
#From summary_results_fdr, seleting those significant for rhoG, rhoE and both
#For purpose of graphing, we redefine the labels
summary_results_fdr_points <- sum_results_combined %>%
 mutate(Sig_Type = case_when(
  rhoG_FDR < 0.05 & rhoE_FDR >= 0.05 ~ "Genetic (ρG)",
  rhoG_FDR >= 0.05 & rhoE_FDR < 0.05 ~ "Environmental (ρE)",
  rhoG_FDR < 0.05 & rhoE_FDR < 0.05 ~ "Both ",
  TRUE ~ "Non-Significant"
 ))

#Plotting a graph to show the rhoG and rhoE distribution of all IP -Protein associations
rho_dis_overall<- ggplot(summary_results_fdr_points, aes(x = rhoG, y = rhoE, color = Sig_Type)) +
   #Adding each pair as a point
   geom_point(alpha = 0.6, size = 1) +
   #Adding a cross
   geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
   geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
   #Colour scale
   scale_color_manual(values = c(
  "Both "           = "#440154FF", 
  "Genetic (ρG)"    = "#21908CFF", 
  "Environmental (ρE)" = "#FDE725FF", 
  "Non-Significant"  = "grey80"      
))
   theme_minimal() +
   #Lablelling
   theme(
    # plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
     legend.position = "none",
     plot.subtitle = element_text(hjust = 0.5, face = "bold",size = 14))+
    # legend.position = "right",
     #legend.text = element_text(size = 12),
   #  legend.title = element_text(size = 14, face = "bold"),
    # legend.key.size = unit(1, "cm"),
     #legend.key.width = unit(1, "cm")) +
   labs(
   # title = "Overall Correlation Landscape",
   subtitle = "Overall Correlation Landscape",
    x = "Genetic Correlation (ρG)",
    y = "Environmental Correlation (ρE)",
    color = "Correlation Significance"
   )

#Saving plot
ggsave(
  filename = "Overall_Correlation_dotplot.png", 
  plot = rho_dis_overall,
  path = output_path,
  width = 12,     
  height = 8
)
```

### 4.2 Individual Lineages

Graphing the indivdiual distribution of each lineage

```{r}
indiv_lin_rho_dis <- ggplot(summary_results_fdr_points, aes(x = rhoG, y = rhoE, color = Sig_Type)) +
    #Adding each pair as a point
    geom_point(alpha = 0.6, size = 1) +
    #Adding a cross
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    
    #Creating separate graphs for the lineages
    facet_wrap(~Sup.Lineage, ncol = 3) + 
    #Colour scale
     scale_color_manual(values = c(
  "Both "           = "#440154FF", 
  "Genetic (ρG)"    = "#21908CFF", 
  "Environmental (ρE)" = "#FDE725FF", 
  "Non-Significant"  = "grey80"   ))+
    #For legend dots
    guides(color = guide_legend(override.aes = list(size = 2.5))) +
    theme_bw() + 
    theme(
      plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 14),
      legend.position = c(0.85, 0.2), #Positioned in empty space
      legend.text = element_text(size = 14),
      legend.title = element_text(size = 14, face = "bold"),
      legend.key.size = unit(1.2, "cm"),
      legend.key.width = unit(1, "cm"),
      strip.background = element_rect(fill = "white"), 
      strip.text = element_text(face = "bold")
    ) +
    labs(
     # title = "By Lineage",
      subtitle = "By Lineage",
      x = "Genetic Correlation (ρG)",
      y = "Environmental Correlation (ρE)",
      color = "Correlation Signficance"
    )


#Saving plot
ggsave(
  filename = "Overall_Correlation_Lineage_dotplot.png", 
  plot = indiv_lin_rho_dis ,
  path = output_path,
  width = 12,     
  height = 8
)
```

### 4.3 Collating Both Graphs

```{r}
#Using Patchwork
comb_overall_rho <- (rho_dis_overall / indiv_lin_rho_dis ) + 
  plot_annotation(
    title = "IP-Protein Association Correlation Analysis",
    tag_levels = 'A') &
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.tag = element_text(size = 20, face = "bold")
  )

print(comb_overall_rho)

ggsave(
  filename = "combined_ip_prot_corr.png", 
  plot = comb_overall_rho ,
  path = output_path,
  width = 12,     
  height = 16
)
```

### 4.3 Volcano Plot on Genetic

```{r}
volcano_data <- sum_results_combined %>%
  mutate(
    log_p = -log10(rhoG_p0),
    #Showing lineage of significnat
    sig_lineage = if_else(rhoG_FDR < 0.05, as.character(Sup.Lineage), "Not Significant")) %>%
    mutate(sig_lineage = factor(sig_lineage, 
                              levels = c("Not Significant", setdiff(unique(sig_lineage), "Not Significant")))) %>%
  arrange(sig_lineage)

#Aesthetics - Colours
unique_lineage <- setdiff(unique(volcano_data$sig_lineage), "Not Significant")
lineage_colors <- viridis::viridis(length(unique_lineage))
names(lineage_colors) <- unique_lineage
final_colors <- c(lineage_colors, "Not Significant" = "grey80")

#Plot
volcano_plot <- ggplot(volcano_data, aes(x = rhoG, y = log_p, color = sig_lineage)) +
  geom_point(alpha = 0.75, size = 0.6) +
  scale_color_manual(values = final_colors) +
  
  #Lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_hline(yintercept = -log10(max(volcano_data$rhoG_p0[volcano_data$rhoG_FDR < 0.05], na.rm = TRUE)), 
             linetype = "dotted", color = "red") +
  
  #Lables for the top 10 proteins
  geom_text_repel(data = volcano_data %>% 
                      filter(rhoG_FDR < 0.05) %>% 
                      arrange(desc(log_p)) %>%
                      distinct(trait2, .keep_all = TRUE) %>%
                      slice_max(order_by = log_p, n = 15),
                    aes(label = trait2), 
                    size = 3.5,
                    fontface = "bold", 
                    color = "black",
                    show.legend = FALSE) +
  
  theme_classic() +
  labs(title = " Overall Genetic Correlation Landscape (ρG)",
       subtitle = "Significant hits. Not taking into account of heritability.",
       x = "Genetic Correlation (ρG)",
       y = "-log10(p-value)",
       color = "Lineage Group") +
  theme(legend.position = "right",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10, face = "bold"))

print(volcano_plot)

#save
ggsave(filename =  "volcano_genetic_landscape.png",
       path = output_path,
       width = 10, 
       height = 8)
```

Top genetic hits can be bidirectional? so in graph can point out the two sides? this is not filtered for genetic

```{r}
sum_results_combined %>%
  filter(trait2 == "CXCL10") %>%
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%
  group_by(Direction) %>%
  tally()
```

```{r}
genetic_0.1_hits %>%
  filter(trait2 == "CXCL10") %>%
  count(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)"))
```

Interestingly for CXCL10/CXCL9 +(chemokines?) others, bidirection even filtered. But others like ADA are majority one direction dominant .

```{r}
lineage_direction <- genetic_0.1_hits %>%
  filter(trait2 == "CXCL10") %>%
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%
  group_by(Sup.Lineage, Direction) %>%
  tally() %>%
  arrange(Sup.Lineage, desc(n))

print(lineage_direction)

```

### 4.4 Summary Table

Summary table from the fdr results based on correlation types for quantification.

```{r}
summary_table <- summary_results_fdr_points %>%
  mutate(across(where(is.numeric), ~ round(., 3))) %>%
  group_by(Sig_Type) %>%
  summarise(
    n = n(),
    mean_rhoG = mean(rhoG),
    mean_rhoE = mean(rhoE),
    mean_h2r1 = mean(h2r1),
    mean_h2r2 = mean(h2r2)
  )

summary_table
#When using this table, best to round to 3dp
```

## 5. Transforming the data frame

For visualisations, the dataframe, which is in a long format, needs to be converted into a wide format (numerical matrix). The dataframes are normalised and NAs were already filtered.

```{r}
#Create a function that applies to all filtered dataset variables
conv_matrix <- function(df, value_col) {
  
  mat <- df %>%
    select(trait1, trait2, all_of(value_col)) %>% 
    
    pivot_wider(
      names_from = trait2,   #changing proteins column
      values_from = all_of(value_col)
    ) %>%

    column_to_rownames("trait1") %>%
    as.matrix()
  
  #replacing NAs from matrix conversion with 0
  mat[is.na(mat)] <- 0
  mat[is.nan(mat)] <- 0
  mat[is.infinite(mat)] <- 0
  
  return(mat)
}
```

## 6. Genetic Hits Analysis

With the genetic hits filtered for h2r \> 0.1.

These are genetic hits and likely instigators of inflammaging. Likely related to ageing.

### 6.1 Analysis

```{r}
genetic_0.1_hits %>%
  summarise(
    total_rows = n(),
    unique_IPs = n_distinct(trait1),
    unique_proteins = n_distinct(trait2)
  )
```

#### 6.1.1 Shared Protein Distribution

```{r}
#Counting how many share proteins
ip_gen_sharing <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_gen <- table(ip_gen_sharing$protein_count)

shared_protein_ip_gen
```

### 6.1 Heatmap

```{r}
genetic_matrix <- conv_matrix(genetic_0.1_hits, "rhoG")

#Lineage
ip_names <- colnames(t(genetic_matrix))

anno_col <- genetic_0.1_hits %>%
  select(trait1, Lineage) %>%
  distinct() %>%
  filter(trait1 %in% ip_names) %>%
  column_to_rownames("trait1")

lineage_split <- anno_col$Lineage

col_fun = colorRamp2(c(-0.6, 0, 0.6), 
                     c("#440154FF", "#e8e8e8", "#21908CFF"))

#Graph
genetic_heatmap<-  Heatmap(t(genetic_matrix), 
          name = "ρG", 
          col = col_fun,
          column_split = lineage_split,
          column_title_side = "bottom",
          column_title_gp = gpar(fontsize = 8, fontface = "bold"),
          column_title_rot = 90,
          cluster_rows = TRUE,
          cluster_columns = TRUE,
          show_column_dend = FALSE,
          show_column_names = FALSE,
          show_row_names = TRUE,
          row_names_gp = gpar(fontsize = 8),
          na_col = "grey90",
          border = FALSE
            )

 #Saving 
#png(filename = file.path(output_path, "gen_heat.png"), 
 #   width = 16, 
  #  height = 8, 
   # units = "in", 
    #res = 300)

draw(genetic_heatmap, column_title = "Genetic Architecture of IP-Protein Pairs")

#dev.off()
```

### 

## 7. Environmental Hits Analysis

Environmental markers - reactive in response externally. Likely not realted to ageing.

### 7.1 Heatmap

"#21908CFF"

```{r}
environmental_matrix <- conv_matrix (environmental_hits, "rhoE")

ip_names_env <- colnames(t(environmental_matrix))

anno_env <- environmental_hits %>%
  select(trait1, Lineage) %>%
  distinct() %>%
  filter(trait1 %in% ip_names_env) %>%
  column_to_rownames("trait1")

lineage_env <- anno_env$Lineage

col_env = colorRamp2(c(-0.6, 0, 0.6), 
                         c("#440154FF", "#e8e8e8", "#21908CFF"))

env_heatmap <- Heatmap(t(environmental_matrix), 
          name = "ρE", 
          col = col_env,
          column_split = lineage_env,
          column_title_side = "bottom",
          column_title_gp = gpar(fontsize = 8, fontface = "bold"),
          column_title_rot = 90,
          cluster_rows = TRUE,
          cluster_columns = TRUE,
          show_column_dend = FALSE,
          show_column_names = FALSE,
          show_row_names = TRUE,
          row_names_gp = gpar(fontsize = 8),
          na_col = "grey90",
          border = FALSE)

draw(env_heatmap, column_title = "Environmental Architecture of IP-Protein Pairs")

```

### 7.2 Shared Protein Distribution

```{r}
#Counting how many share proteins
ip_env_sharing <- environmental_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_env <- table(ip_env_sharing$protein_count)

shared_protein_ip_env
```

## 8. Both Genetic and Environmentally Correlated

All hits are positive. Only CXCL9 protein has IP associations.

Overall CXCL9 had the phenotype pP correlation.

Deriving from genetic pG and environmental pE but rhoG is stronger/more dominant than rhoE.

```{r}
#Plotting with different graph
both_dual <- both_hits %>%
  mutate(IP = reorder(trait1, abs(rhoG))) %>%
  select(IP, Protein = trait2, rhoG, rhoE)

#For the graph mean + sd visuals
stats_rhoG <- both_hits %>%
  summarise(mean = mean(rhoG), sd = sd(rhoG)) %>%
  mutate(min = mean - sd, max = mean + sd)
stats_rhoE <- both_hits %>%
  summarise(mean = mean(rhoE), sd = sd(rhoE)) %>%
  mutate(min = mean - sd, max = mean + sd)

#Dumbell graph to show both rhoG and rhoE
ggplot(both_dual, aes(x = reorder(IP, abs(rhoG)))) +
  
  #Mean +sd
  geom_rect(aes(xmin = -Inf, xmax = Inf, 
                ymin = stats_rhoG$min, ymax = stats_rhoG$max), 
            fill = "#440154FF", alpha = 0.02) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, 
                ymin = stats_rhoE$min, ymax = stats_rhoE$max), 
            fill = "#21908CFF", alpha = 0.02) +
  geom_hline(yintercept = stats_rhoG$mean, color = "#440154FF", linetype = "solid", alpha = 0.4) +
  geom_hline(yintercept = stats_rhoE$mean, color = "#21908CFF", linetype = "solid", alpha = 0.4) +
  #label for mean +sd
  annotate("text", x = 1, y = stats_rhoG$mean, label = "MEAN", 
           angle = 90, size = 3, color = "#440154FF", vjust = -1, hjust = -0.2,fontface = "bold") +
  annotate("text", x = 1, y = stats_rhoG$max+0.01, label = "STDEV", 
           angle = 90, size = 3, color = "#440154FF", vjust = -1, hjust = -0.2,fontface = "bold") +
           
  
  geom_segment(aes(xend = IP, y = rhoG, yend = rhoE), color = "grey85", size = 1) +
  geom_point(aes(y = rhoG, color = "Genetic (ρG)"), size = 5) +
  geom_point(aes(y = rhoE, color = "Environmental (ρE)"), size = 5) +
  scale_y_continuous(expand = c(0, 0), limits = c(0.2, 0.55)) + #x axis
  coord_flip() +
  scale_color_manual(values = c("#21908CFF","#440154FF"))+
  labs(
    title = "Both Genetic and Environmental Correlations",
    x = "IP", y = "Correlation Coefficient (ρ) with CXCL9",
    color = "Correlation Type"
  ) +
  theme_minimal() +
    theme(
    legend.position = "bottom",
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 11),
    plot.title = element_text(face = "bold", hjust = 0.5,margin = margin(b = 20)),
    axis.title = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 9),
    axis.line = element_line(color = "black"))


#Saving plot
ggsave(
  filename = "dumbell_both_gen_env.png", 
  path = output_path,
  width = 10,     
  height = 8
)


print(stats_rhoG$mean)
print(stats_rhoG$sd)
print(stats_rhoE$mean)
print(stats_rhoE$sd)

```

CXCL9 more geneticaly correlated than environmental counterpart though both significant. Heterogeneity of immune cell influence. MEAN and SD purely for trend, no biological context. Effect size are significantly higher in genetic than environmental . They are all from CD8 T cell lineage.

## 9. Genetic -\> Environmental

Those that are genetically correlated but also have environmental correlations

### 9.1 Heatmap

```{r}
gen_env_matrix <- conv_matrix (gen_env_0.1, "rhoE")

ip_names_gen_env <- colnames(t(gen_env_matrix))

anno_gen_env<- gen_env_0.1 %>%
  select(trait1, Lineage) %>%
  distinct() %>%
  filter(trait1 %in% ip_names_gen_env) %>%
  column_to_rownames("trait1")

lineage_gen_env <- anno_gen_env$Lineage

col_gen_env = colorRamp2(c(-0.6, 0, 0.6), 
                          c("#440154FF", "#e8e8e8", "#21908CFF"))


gen_env_heatmap<-Heatmap(t(gen_env_matrix),
          name = "rhoE",
          col = col_gen_env,
          column_split = lineage_gen_env,
          column_title_side = "bottom",
          column_title_gp = gpar(fontsize = 8, fontface = "bold"),
          column_title_rot = 90,
          cluster_rows = FALSE, 
          cluster_columns = FALSE,
          show_column_dend = FALSE,
          show_column_names = FALSE,
          show_row_names = TRUE,
          row_names_gp = gpar(fontsize = 8),
          na_col = "grey90",
          border = FALSE
          )

draw(gen_env_heatmap, 
     column_title = "Genetically with Environmental IP-Protein Correlations ")
```

### 9.2 Graph Plot

Negative correlations are present as well, though positive correlations dominate. If they are positive in genetics, then they are positive in environmental. No positive genetic and negative environmental.

```{r}
ggplot(gen_env_0.1, aes(x = rhoG, y = rhoE, color = trait2)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = "Genetically Correlated with Environmental Correlations (h² > 0.1)",
    subtitle = "IP–Protein associations coloured by protein",
    x = "Genetic Correlation (rhoG)",
    y = "Environmental Correlation (rhoE)",
    color = "Protein"
  )
```

### 9.3 Dumbbell

```{r}


#Need df with lineag e
gen_env_plot_data <- gen_env_0.1 %>%
  arrange(Sup.Lineage, rhoG) %>%
  mutate(Plot_Label = factor(trait1, levels = unique(trait1)))

#Plot
ggplot(gen_env_plot_data) +
  geom_tile(aes(y = Plot_Label, x = -0.6, fill = Sup.Lineage), 
            width = 0.05, height = 1) +
  #connection line
  geom_segment(aes(y = Plot_Label, yend = Plot_Label, 
                   x = rhoG, xend = rhoE), 
               color = "grey80", size = 0.8) +
  #Points
  geom_point(aes(y = Plot_Label, x = rhoG, color = "Genetic (rhoG)"), size = 2.5) +
  geom_point(aes(y = Plot_Label, x = rhoE, color = "Environmental (rhoE)"), size = 2.5) +
  
  #Separate the proteins by graph
  facet_grid(trait2 ~ ., scales = "free_y", space = "free_y") +
  #0 line
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  #Aesthetics
  scale_color_manual(values = c("Genetic (rhoG)" = "#440154FF", 
                                "Environmental (rhoE)" = "#21908CFF")) +
  scale_fill_manual(values = c("B cells" = "#47c16eFF", 
                               "monocytes" = "#8FD744FF", 
                               "T cells" = "#FDE725FF")) +
  theme_minimal() +
  labs(
    title = "Genetically Correlated Environmental Hits",
    subtitle = "Connected dots represent the same IP-Protein pair",
    x = "Correlation Coefficient (ρ)",
    y = "IP",
    color = "Correlation Type"
  ) +


  theme(
    strip.text.y = element_text(angle = 0, face = "bold", hjust = 0),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(size = 11, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.spacing = unit(1, "lines")
    legend.position = "bottom")

ggsave(
  filename = "gen_env.png", 
  path = output_path,
  width = 10,     
  height = 12
)
```

```         
```

## 10. Environmental -\> Genetic

### 10.1 Heatmap

```{r}
env_gen_matrix <- conv_matrix(env_gen, "rhoE")

ip_names_env_gen <- colnames(t(env_gen_matrix))

anno_env_gen <- env_gen %>%
  select(trait1, Lineage) %>%
  distinct() %>%
  filter(trait1 %in% ip_names_env_gen) %>%
  column_to_rownames("trait1")

lineage_env_gen <- anno_env_gen$Lineage

col_env_gen = colorRamp2(c(-0.6, 0, 0.6), 
                          c("#440154FF", "#e8e8e8", "#21908CFF"))

env_gen_heatmap<-Heatmap(t(env_gen_matrix),
          name = "rhoE",
          col = col_env_gen,
          column_split = lineage_env_gen,
          column_title_side = "bottom",
          column_title_gp = gpar(fontsize = 8, fontface = "bold"),
          column_title_rot = 90,
          cluster_rows = FALSE, 
          cluster_columns = FALSE,
          show_column_dend = FALSE,
          show_column_names = FALSE,
          show_row_names = TRUE,
          row_names_gp = gpar(fontsize = 8),
          na_col = "grey90",
          border = FALSE
          )

draw(env_gen_heatmap, 
     column_title = "Environmental then Genetic IP-Protein Correlations ")
```

All positive correlations, few overlaps.

### 10.2 Plot Graph

```{r}
ggplot(env_gen, aes(x = rhoE, y = rhoG, color = trait2)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = " Environmentally Correlated with Genetic Correlations ",
    subtitle = "IP–Protein associations coloured by protein",
    x = "Environmental Correlation (rhoE)",
    y = "Genetic Correlation (rhoG)",
    color = "Protein"
  )
```

## 
