---
title: "solar_analysis_filtering"
format: html
editor: visual
---

# SOLAR Results Analysis

After obtaining the IP-Protein associations from the SOLAR analysis. Exploration and filtering is needed for interpretation.

## 1. Set up

### 1.1 Load libraries

```{r, message=FALSE}
library(stringr)
library(dplyr)
library(tidyr)
library(tibble)
library(tidyverse)

#Plotting
library(viridis) #Colour blind friendly palette
library(ggplot2) #Graphing
library(ggrepel) #Volcano
library(patchwork) #Combine Plots

#Heatmap
library(ComplexHeatmap)

#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
library(ggforce) #Circular network modules

#install.packages("visNetwork")
library(visNetwork) #interactive network
```

### 1.2 Loading Data

```{r}
#Importing the summary results
summary_results<- read.delim("~/Twins_Project/data/solar_analysis/summary_heritability.csv")

#Loading in the dataset for sup.lineage information 
load("~/Twins_Project/data/raw_olink_ip/results/Olink_associations.RData")
```

```{r}
output_path <- "~/Twins_Project/visualisations"
```

## 2. Filtering of Significant IP-Protein Associations

### 2.1 FDR Adjusted Filtering

Adjusting with FDR to find those significant based on FDR \< 0.05.

```{r}
#Adding in FDR of rhoG and rhoE
summary_results_fdr <- summary_results %>%
  mutate(
    rhoG_FDR = p.adjust(rhoG_p0, method = "fdr"),
    rhoE_FDR = p.adjust(rhoE_p, method = "fdr")
  )

```

### 2.2 Adding in Lineages from infoF dataframe

Using Sup.Lineage as it contains T cells, NK cells, monocytes, dendritic cells, B cells to determine the IP lineages.

```{r}
#Need to clean the IMMUNO_X.new names to match summary_result_fdr as it doesnt have .new
infoF_clean <- infoF %>%
  mutate(SET.NAME_clean = gsub("\\.new$", "", SET.NAME)) %>%
  distinct(SET.NAME_clean, .keep_all = TRUE)

#Combining the lineage columns to the summary results dataframe
sum_results_combined <- summary_results_fdr %>%
  left_join(infoF_clean %>% select(SET.NAME_clean, Lineage, Sup.Lineage, Sub.Lineage), 
            by = c("trait1" = "SET.NAME_clean"))

#Just for visuals, relabelling of default df
sum_results_combined <- sum_results_combined %>%
  mutate(Sup.Lineage = recode(Sup.Lineage,
    "dendritic cells" = "DC",
    "monocytes"       = "Monocytes"))
```

### 2.3 Filtering based on Genetic, Environment or both associations based on the SOLAR outputs

Using the heritability h2r results with rhoG, we can determine genetic hits. For environmental hits, h2r does not need to be significant.

#### Genetic Hits

```{r}
#Experimenting with different h2r thresholds
genetic_0.1_hits <- sum_results_combined %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05)
#4300

genetic_0.3_hits <- sum_results_combined %>%
  filter(h2r1 > 0.3 & h2r2 > 0.3 & rhoG_FDR < 0.05)
#4047

genetic_0.5_hits <- sum_results_combined %>%
  filter(h2r1 > 0.5 & h2r2 > 0.5 & rhoG_FDR < 0.05)
#1068

```

#### Environmental hits

Only accounting for rhoE as h2r can be low or high

environmental_hits \<- sum_results_combined %\>%

filter(rhoE_FDR \< 0.05)

```{r}
#Only accounting for rhoE as h2r can be low or high 
environmental_hits <- sum_results_combined %>%
  filter(rhoE_FDR < 0.05)
```

#### Both hits

```{r}
#Significant in both genetic and environmental correlations
both_hits <- sum_results_combined %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05 & rhoE_FDR < 0.05)
```

### 2.4 Further Filtering

Further filtering of genetically correlated pairs with environmental associations and vise versa - environmentally correlated pairs with genetic associations.

#### Genetics -\> Environmental

```{r}
#Genetics -> how many are also environmental 
gen_env_0.1 <- genetic_0.1_hits %>%
  filter(rhoE_p < 0.05)

gen_env_0.3 <- genetic_0.3_hits %>%
  filter(rhoE_p < 0.05)
#same 130 pairs 

gen_env_0.5 <- genetic_0.5_hits %>%
  filter(rhoE_p < 0.05)
#only 1 IMMUNO_41521 TNFB
```

#### Environmental -\> Genetics

```{r}
#Environmental -> how many are genetics
env_gen <- environmental_hits%>%
  filter(rhoG_p0<0.05)
```

### 2.5 Saving the output environment

Saving to be loaded elsewhere.

```{r}
save.image(file = "solar_analysis_filtering_results.RData")
```

## 3. Exploration of the overall cohort

sum_results_combined exploration. This is the general counts and distribution. IP-Protein Results (non-filtered).

### 3.1 Baseline ρP Phenotype Correlation

To validate the total correlations observed between IP-protein associations.

```{r}
rhoP_plot<- ggplot(sum_results_combined, aes(x = rhoP)) +
  geom_histogram(bins = 50, fill = "#440154FF", color = "white") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", linewidth = 1) +
  #Labels
  labs(title = "Distribution of Phenotypic Correlations (ρP)",
       x = "Phenotypic Correlation (ρP)",
       y = "Frequency") +
  #Aesthetics
 theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),     
    axis.title.y = element_text(size = 12,face = "bold"),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12,face = "bold"),
    plot.title = element_text(size = 16, face = "bold",hjust = 0.5, margin = margin(b = 10)),
    plot.title.position = "plot",
    legend.position = "none",
    axis.line = element_line(color = "black"),
    panel.grid.major.y = element_blank() #No vertical lines
  )



#Saving plot
ggsave(
  filename = "overall_cohort/Phenotypic_Correlation_Distribution.png", 
  plot = rhoP_plot,
  path = output_path,
  width = 12,     
  height = 8
)
```

### 3.2 Protein Distribution

Analysing how many proteins per IPs are present in each filtering.

#### Overall Protein Distribution

```{r}
sum_results_combined %>%
  summarise(
    total_rows = n(),
    unique_IPs = n_distinct(trait1),
    unique_proteins = n_distinct(trait2)
  )
```

#### Unique IP to Protein Frequency Distribution

```{r}
ip_per_protein <- sum_results_combined %>%
  distinct(trait1, trait2) %>% #Should be unique but sanity checks
  count(trait2, sort = TRUE, name = "n_IPs")

print(ip_per_protein)
```

#### Graph

```{r}
ip_per_protein %>%
  ggplot(aes(x = reorder(trait2, n_IPs), y = n_IPs)) +
  geom_col(aes(fill = n_IPs), alpha = 0.8, color = "white", linewidth = 0.3) +

  #Numerical values next to bar
  geom_text(aes(label = n_IPs), hjust = -0.2, size =1.5, color = "black", fontface = "bold")+
  #Colour scheme
  scale_fill_viridis_c(option = "viridis")+
  #Axis
  coord_flip() +
  #Labelling
  labs(title = "Overall Distribution of IPs per Protein",
       x = "Protein", y = "Number of IPs ") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 6),     
    axis.title.y = element_text(size = 10),
    axis.text.x = element_text(size = 6),
    axis.title.x = element_text(size = 10),
    plot.title = element_text(size = 16, face = "bold",hjust = 0.5),
    legend.position = "none",
    panel.grid.major.y = element_blank() #No vertical lines
  )

#Saving plot
ggsave(
  filename = "overall_cohort/Overall_Protein_Distribution.png", 
  path = output_path,
  width = 10,     
  height = 8
)
```

In the 0.1 and 0.3 top hits, find what proteins are most frequent/popular for genetic. Observing whether there are changes in protein frequency patterns.

45 proteins altogether.

#### Top 15 Proteins

Better Visualisation.

```{r}
top_15_proteins <- ip_per_protein %>%
  slice_max(n_IPs, n = 15, with_ties = FALSE) %>%
  ggplot(aes(x = reorder(trait2, n_IPs), y = n_IPs)) +
  geom_col(aes(fill = n_IPs), alpha = 0.8, color = "white", linewidth = 0.3) +

  #Numerical values next to bar
  geom_text(aes(label = n_IPs), hjust = -0.2, size =3, color = "black", fontface = "bold")+
  #Colour scheme
  scale_fill_viridis_c(option = "viridis")+
  #Axis
  coord_flip() +
  scale_y_continuous(
    limits = c(0, 3800),               
    breaks = seq(0, 4000, by = 500),
    expand = expansion(mult = c(0, 0.1)))+
  #Labelling
  labs(title = "Top 15 Proteins by Number of IP Connections",
       x = "Protein", y = "Number of IPs ") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),     
    axis.title.y = element_text(size = 12,face = "bold"),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12,face = "bold"),
    plot.title = element_text(size = 16, face = "bold",hjust = 0.5, margin = margin(b = 10)),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    panel.grid.major.y = element_blank() #No vertical lines
  )

#Saving plot
ggsave(
    filename = "overall_cohort/Overall_Top_15_Protein_Distribution.png", 
    plot = top_15_proteins,
  path = output_path,
  width = 10,     
  height = 8
)
```

### 3.2 IP sharing count

#### Distribution

```{r}
protein_per_ip <- sum_results_combined %>%
  distinct(trait1, trait2) %>% #Should be unique but sanity checks
  count(trait1, sort = TRUE, name = "n_Prot")

print(protein_per_ip )
```

#### IPs to Proteins sharing number

```{r}
#Table
shared_protein_ip <- table(protein_per_ip$n_Prot)

print(shared_protein_ip)
```

#### Graph

```{r}
#DF conversion of the table
ip_shared_prot_df <- as.data.frame(shared_protein_ip)
colnames(ip_shared_prot_df ) <- c("Num_Proteins", "Count_IPs")


ip_shared_prot_plot <- ggplot(ip_shared_prot_df, aes(x = Num_Proteins, y = Count_IPs)) +
  geom_col(aes(fill = Count_IPs), alpha = 0.8, color = "white", linewidth = 0.3) +

  #Numerical values next to bar
  geom_text(aes(label = Count_IPs), hjust = -0.2, size =3, color = "black", fontface = "bold")+
  #Colour scheme
  scale_fill_viridis_c(option = "viridis")+
  #Axis
  coord_flip() +

  labs(title = "Protein Connectivity to IP Distribution",
       x = "Number of Proteins Connected to a Single IP", y = "Number of IPs ") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),     
    axis.title.y = element_text(size = 10,face = "bold"),
    axis.text.x = element_text(size = 8),
    axis.title.x = element_text(size = 10,face = "bold"),
    plot.title = element_text(size = 16, face = "bold",hjust = 0.5, margin = margin(b = 10)),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    panel.grid.major.y = element_blank() #No vertical lines
  )

plot(ip_shared_prot_plot)
#Saving plot
ggsave(
  filename = "overall_cohort/Overall_IP_Prot_Distribution.png", 
  path = output_path,
  width = 10,     
  height = 8
)
```

### Plot Combination

```{r}

spacer <-("
  #1111#
  222333
")

overall_ip_prot_dis_panels <- (rhoP_plot)/
  (top_15_proteins + ip_shared_prot_plot)+
    plot_layout(heights = c(1, 1))

  overall_ip_prot_dis_panels<- rhoP_plot +top_15_proteins+ip_shared_prot_plot +plot_layout(design = spacer)+
  
plot_annotation(
    tag_levels = 'A') &
  theme(
    plot.tag = element_text(size = 25, face = "bold") ,
    plot.title = element_text(size = 28, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 18)
  )

print (overall_ip_prot_dis_panels)
#Saving 
ggsave(
  filename = "overall_cohort/Overall_IP_Prot_Final_Panel.png",
  plot = overall_ip_prot_dis_panels,
  path = output_path,
  width = 24,
  height = 20
)
```

### 3.3 Lineage Investigation

```{r}
lineage_summary <- sum_results_combined %>%
  group_by(Lineage, Sup.Lineage, Sub.Lineage) %>%
  summarise(
    n_unique_IPs = n_distinct(trait1),
    n_unique_Proteins = n_distinct(trait2),
    n_Pairs = n() 
  ) %>%
  arrange(desc(n_Pairs))

print(lineage_summary)
```

27 unique lineages across pairs.

Sub Lineage investigations show order dominance of T cells -\> B cells -\> monocytes -\> NK cells -\> DC

```{r}
sum_results_combined %>%
  group_by(Sup.Lineage) %>%
  summarise(unique_ip_count = n_distinct(trait1)) %>%
  arrange(desc(unique_ip_count))
```

### 3.4 Heritability

Counting how many pairs passed h2 \> 0.1 for stability.

```{r}
sum_results_combined %>%
  count(h2r1 >0.1 & h2r2 >0.1)
```

## 4. Post-Filtering Summary Statistics

Show the distribution of those genetically, environmentally and both significant.

### 4.1 Graphing Correlation Landscape

Visualising the distribution of genetic and environmental correlations.

```{r}
#From summary_results_fdr, seleting those significant for rhoG, rhoE and both
#For purpose of graphing, we redefine the labels
summary_results_fdr_points <- sum_results_combined %>%
  mutate(
    is_genetic = h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05,
    is_env     = rhoE_FDR < 0.05
  ) %>%
  mutate(Sig_Type = case_when(
    is_genetic & is_env ~ "Both",
    is_genetic         ~ "Genetic (ρG)",
    is_env             ~ "Environmental (ρE)",
    TRUE               ~ "Non-Significant"
  ))%>%
  arrange(factor(Sig_Type, levels = c("Non-Significant", "Genetic (ρG)", "Environmental (ρE)", "Both"))) #for plotting, move non sig points to the back

#Plotting a graph to show the rhoG and rhoE distribution of all IP -Protein associations
rho_dis_overall<- ggplot(summary_results_fdr_points, aes(x = rhoG, y = rhoE, color = Sig_Type)) +
   #Adding each pair as a point
   geom_point(alpha = 0.6, size = 1) +
   #Adding a cross
   geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
   geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
   #Colour scale
   scale_color_manual(values = c(
  "Both"           = "#decb21", 
  "Genetic (ρG)"    = "#21908CFF", 
  "Environmental (ρE)" = "#440154FF", 
  "Non-Significant"  = "grey90"      
))+
   theme_bw() + 
   #Lablelling
   theme(
    # plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
     legend.position = "none",
     plot.subtitle = element_text(hjust = 0.5, face = "bold",size = 14),
     axis.title.y = element_text(size = 12,face = "bold"),
     axis.title.x = element_text(size = 12,face = "bold"))+
    # legend.position = "right",
     #legend.text = element_text(size = 12),
   #  legend.title = element_text(size = 14, face = "bold"),
    # legend.key.size = unit(1, "cm"),
     #legend.key.width = unit(1, "cm")) +
   labs(
   # title = "Overall Correlation Landscape",
   subtitle = "Overall Correlation Landscape",
    x = "Genetic Correlation (ρG)",
    y = "Environmental Correlation (ρE)",
    color = "Correlation Significance"
   )

#Saving plot
ggsave(
  filename = "overall_cohort/Overall_Correlation_dotplot.png", 
  plot = rho_dis_overall,
  path = output_path,
  width = 12,     
  height = 8
)
```

### 4.2 Individual Lineages

Graphing the individual distribution of each lineage

```{r}

indiv_lin_rho_dis <- ggplot(summary_results_fdr_points, aes(x = rhoG, y = rhoE, color = Sig_Type)) +
    #Adding each pair as a point
    geom_point(alpha = 0.6, size = 1) +
    #Adding a cross
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    
    #Creating separate graphs for the lineages
    facet_wrap(~Sup.Lineage, ncol = 3) + 
    #Colour scale
     scale_color_manual(values = c(
  "Both"           = "#decb21", 
  "Genetic (ρG)"    = "#21908CFF", 
  "Environmental (ρE)" = "#440154FF",  
  "Non-Significant"  = "grey90"   ))+
    #For legend dots
    guides(color = guide_legend(override.aes = list(size = 2.5))) +
    theme_bw() + 
    theme(
      plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 14),
      legend.position = c(0.85, 0.2), #Positioned in empty space
      legend.text = element_text(size = 14),
      legend.title = element_text(size = 14, face = "bold"),
      legend.key.size = unit(1.2, "cm"),
      legend.key.width = unit(1, "cm"),
      strip.background = element_rect(fill = "white"), 
      strip.text = element_text(face = "bold"),
      axis.title.y = element_text(size = 12,face = "bold"),
      axis.title.x = element_text(size = 12,face = "bold"),
    ) +
    labs(
     # title = "By Lineage",
      subtitle = "By Super-Lineage",
      x = "Genetic Correlation (ρG)",
      y = "Environmental Correlation (ρE)",
      color = "Correlation Signficance"
    )

#Saving plot
ggsave(
  filename = "overall_cohort/Overall_Correlation_Lineage_dotplot.png", 
  plot = indiv_lin_rho_dis ,
  path = output_path,
  width = 12,     
  height = 8
)
```

### 4.3 Collating Both Graphs

```{r}
#Using Patchwork
comb_overall_rho <- (rho_dis_overall / indiv_lin_rho_dis ) + 
  plot_annotation(
    tag_levels = 'A') &
  theme(
    plot.title = element_text(size = 25, face = "bold", hjust = 0.5),
    plot.tag = element_text(size = 25, face = "bold")
  )

print(comb_overall_rho)

ggsave(
  filename = "overall_cohort/Overall_Combined_Correlations_Dotplot.png", 
  plot = comb_overall_rho ,
  path = output_path,
  width = 12,     
  height = 16
)
```

### 4.4 Bidirectionality

```{r}
sum_results_combined %>%
  filter(trait2 == "CXCL10") %>% #can filter lineage, trait 
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%
  group_by(Direction) %>%
  tally()
```

### 4.5 Summary Table

Summary table from the fdr results based on correlation types for quantification.

```{r}
summary_table <- summary_results_fdr_points %>%
  mutate(across(where(is.numeric), ~ round(., 3))) %>%
  group_by(Sig_Type) %>%
  summarise(
    n = n(),
    mean_rhoG = mean(rhoG),
    mean_rhoE = mean(rhoE),
    mean_h2r1 = mean(h2r1),
    mean_h2r2 = mean(h2r2)
  )

summary_table
#When using this table, best to round to 3dp
```

## 5. Transforming the data frame

For visualisations, the dataframe, which is in a long format, needs to be converted into a wide format (numerical matrix). The dataframes are normalised and NAs were already filtered.

```{r}
#Create a function that applies to all filtered dataset variables
conv_matrix <- function(df, value_col) {
  
  mat <- df %>%
    select(trait1, trait2, all_of(value_col)) %>% 
    
    pivot_wider(
      names_from = trait2,   #changing proteins column
      values_from = all_of(value_col)
    ) %>%

    column_to_rownames("trait1") %>%
    as.matrix()
  
  #replacing NAs from matrix conversion with 0
  mat[is.na(mat)] <- 0
  mat[is.nan(mat)] <- 0
  mat[is.infinite(mat)] <- 0
  
  return(mat)
}
```

## 6. Genetic Hits Analysis

With the genetic hits filtered for h2r \> 0.1.

These are genetic hits and likely instigators of inflammaging. Likely related to ageing.

### 6.1 Summary IP to Protein

#### Number of IPs and Proteins

```{r}
genetic_0.1_hits %>%
  summarise(
    total_rows = n(),
    unique_IPs = n_distinct(trait1),
    unique_proteins = n_distinct(trait2)
  )
```

```{r}
genetic_0.1_hits  %>%
  summarise (
    mean_h2r1 = mean(h2r1),
    sd_h2r1   = sd(h2r1),
    max_h2r1 = max(h2r1),
    mean_h2r2 = mean(h2r2),
    sd_h2r2= sd(h2r2),
    max_h2r2 = max(h2r2),
    mean_rhoG = mean(abs(rhoG)),
    sd_rhoG= sd(abs(rhoG)),
    mean_rhoE = mean(abs(rhoE)),
    sd_rhoE= sd(abs(rhoE))
  )

```

### 6.2 Protein Distribution

```{r}
ip_per_protein_gen <- genetic_0.1_hits%>%
  distinct(trait1, trait2) %>% #Should be unique but sanity checks
  count(trait2, sort = TRUE, name = "n_IPs")

print(ip_per_protein_gen)
```

```{r}
gen_top15_prot <-ip_per_protein_gen %>%
  slice_max(n_IPs, n = 15, with_ties = FALSE) %>%
  ggplot(aes(x = reorder(trait2, n_IPs), y = n_IPs)) +
  geom_col(aes(fill = n_IPs), alpha = 0.8, color = "white", linewidth = 0.3) +

  #Numerical values next to bar
  geom_text(aes(label = n_IPs), hjust = -0.2, size =3, color = "black", fontface = "bold")+
  #Colour scheme
  scale_fill_viridis_c(option = "viridis")+
  #Axis
  coord_flip() +
  scale_y_continuous(
    limits = c(0, 1500),               
    expand = expansion(mult = c(0, 0.1)))+
  #Labelling
  labs(title = "Top 15 Proteins by Number of IP Connections",
       x = "Protein", y = "Number of IPs ") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),     
    axis.title.y = element_text(size = 12,face = "bold"),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12,face = "bold"),
    plot.title = element_text(size = 14, face = "bold",hjust = 0.5, margin = margin(b = 10)),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    panel.grid.major.y = element_blank() #No vertical lines
  )

#Saving plot
ggsave(
  filename = "genetic_hits/Gen_Top_15_Protein_Distribution.png", 
  plot = gen_top15_prot,
  path = output_path,
  width = 10,     
  height = 8
)
```

### 6.3 IP Shared Protein Distribution

```{r}
#Counting how many share proteins
ip_gen_sharing <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_gen <- table(ip_gen_sharing$protein_count)

shared_protein_ip_gen
```

```{r}
#DF conversion of the table
ip_shared_prot_df_gen <- as.data.frame(shared_protein_ip_gen)
colnames(ip_shared_prot_df_gen ) <- c("Num_Proteins", "Count_IPs")


ip_shared_prot_plot_gen <- ggplot(ip_shared_prot_df_gen, aes(x = Num_Proteins, y = Count_IPs)) +
  geom_col(aes(fill = Count_IPs), alpha = 0.8, color = "white", linewidth = 0.3) +

  #Numerical values next to bar
  geom_text(aes(label = Count_IPs), hjust = -0.2, size =3, color = "black", fontface = "bold")+
  #Colour scheme
  scale_fill_viridis_c(option = "viridis")+
  #Axis
  coord_flip() +

  labs(title = "Protein Connectivity to IP Distribution",
       x = "Number of Proteins Connected to a Single IP", y = "Number of IPs ") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),     
    axis.title.y = element_text(size = 10,face = "bold"),
    axis.text.x = element_text(size = 8),
    axis.title.x = element_text(size = 10,face = "bold"),
    plot.title = element_text(size = 12, face = "bold",hjust = 0.5, margin = margin(b = 10)),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    panel.grid.major.y = element_blank() #No vertical lines
  )

plot(ip_shared_prot_plot_gen)
#Saving plot
ggsave(
  filename = "genetic_hits/Gen_IP_Prot_Distribution.png", 
  plot = ip_shared_prot_plot_gen,
  path = output_path,
  width = 10,     
  height = 8
)
```

#### Investigating 9 Protein Hub

```{r}
#Obtaining the hub of 5 IPs 
gen_hub_9_prot <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  filter(protein_count == 9) %>%
  pull(trait1)

#Convert to df
gen_hub_9_prot_df <- genetic_0.1_hits %>%
  filter(trait1 %in% gen_hub_9_prot)

```

```{r}
#Lineage investigation
gen_hub_9_prot_df %>%
  group_by(trait1) %>%
  summarise(lineage = first(Sub.Lineage))
```

They are all CD8 T cells

```{r}
#Protein investigation
gen_hub_9_prot_df %>%
  group_by(trait2) %>%
  summarise(ip_count = n_distinct(trait1)) %>%
  arrange(desc(ip_count))
```

All share the same proteins.

```{r}
#Heritability values
gen_hub_9_prot_df %>%
  summarise(
    avg_h2_IP = mean(h2r1),
    sd_h2_IP = sd(h2r1),
    avg_h2_Protein = mean(h2r2),
    sd_h2_Protein = sd(h2r2),
    avg_rhoG = mean(rhoG),
    sd_rhoG = sd(rhoG)
    
  )
```

#### Investigating 8 Protein Hub

```{r}
#Obtaining the hub of 5 IPs 
gen_hub_8_prot <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  filter(protein_count == 8) %>%
  pull(trait1)

#Convert to df
gen_hub_8_prot_df <- genetic_0.1_hits %>%
  filter(trait1 %in% gen_hub_8_prot)
```

```{r}
#Lineage investigation
gen_hub_8_prot_df %>%
  group_by(trait1) %>%
  summarise(lineage = first(Sub.Lineage))
```

```{r}
#Protein investigation
gen_hub_8_prot_df %>%
  group_by(trait2) %>%
  summarise(ip_count = n_distinct(trait1)) %>%
  arrange(desc(ip_count))
```

### 6.4 Lineage Distribution

```{r}
#Lineage count
genetic_0.1_hits %>%
  count(Lineage, sort = TRUE)
```

```{r}
#Sup lineage 
genetic_0.1_hits %>%
  group_by(Sup.Lineage) %>%
  summarise(unique_ip_count = n_distinct(trait1)) %>%
  arrange(desc(unique_ip_count))
```

### 6.5 Bidirectionality

#### Overall

```{r}
#overall count
genetic_0.1_hits %>%
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%
  count(Direction)
```

#### Per Lineage

```{r}
#count per lineage
gen_bidirection_counts <- genetic_0.1_hits %>%
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%
  count(Direction, Sup.Lineage)

print(gen_bidirection_counts)
```

##### Graph

Graphing the bidirectionality by proportion in super lineages

```{r}
gen_bidirection <- ggplot(gen_bidirection_counts, aes(x = Direction, y = n, fill = Direction)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.5, size = 3,color = "black", fontface = "bold") +
  facet_wrap(~Sup.Lineage, scales = "free_y") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c("Positive (+)" = "#21908CFF", "Negative (-)" = "#440154FF")) +
  theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      legend.position = c(0.85, 0.2), #Positioned in empty space
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12, face = "bold"),
      legend.key.size = unit(1.2, "cm"),
      legend.key.width = unit(1, "cm"),
      strip.background = element_rect(fill = "white"), 
      strip.text = element_text(face = "bold"),
      axis.title.y = element_text(size = 12,face = "bold"),
      axis.title.x = element_text(size = 12,face = "bold"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    ) +
  labs(title = "Directionality by Lineage",
       y = "Number of Significant Genetic Hits")


#Saving plot
ggsave(
  filename = "genetic_hits/Gen_Direction_Lineage.png", 
  plot = gen_bidirection,
 path = output_path,
  width = 12,     
  height = 8
)
```

#### Per Protein

```{r}
genetic_0.1_hits %>%
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%
  group_by(trait2, Direction) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Direction, values_from = count, values_fill = 0) %>%
  mutate(Total = `Positive (+)` + `Negative (-)`,
         Prop_Positive = `Positive (+)` / Total) %>%
  arrange(desc(Total))
```

```{r}
genetic_0.1_hits %>%   
  filter(trait2 == "CXCL10") %>%   
  mutate(Direction = if_else(rhoG < 0, 
                             "Negative (-)",
                             "Positive (+)")) %>%   
  group_by(Sup.Lineage, Direction) %>%  
  tally() %>%   
  arrange(Sup.Lineage, desc(n)) 
```

#### Panelling

```{r}

gen_panels <-  (gen_bidirection) /(gen_top15_prot + ip_shared_prot_plot_gen)+
    plot_layout(heights = c(1, 1))

gen_panels <- gen_bidirection + gen_top15_prot + ip_shared_prot_plot_gen +
  
  plot_layout(design = spacer) +
  
  plot_annotation(  
    tag_levels = 'A'
  ) &
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.tag.position = c(0.08, 0.99),
    plot.tag = element_text(size = 18, face = "bold"))

ggsave(
  filename = "genetic_hits/Genetic_Panels.png",
  plot = gen_panels,
  path = output_path,
  width = 20,
  height = 18
)
```

### 6.6 Hubs Network Graph

To visualise hubs and IP - Protein connections, can filter how many hubs we want in the network graph.

```{r}
#Lookup data 
lineage_info_gen <- genetic_0.1_hits %>%
  select(trait1, Sub.Lineage) %>%
  distinct(trait1, .keep_all = TRUE)

#defining hub list 
hub_data_gen <- genetic_0.1_hits %>%
  group_by(trait1) %>%
  filter(n() >= 7) %>% #Select the number of Hubs +
   mutate(n_count = n()) %>%
  ungroup()%>%
  select(trait1, trait2, Sub.Lineage, n_count)

```

```{r}
#Graphing

#Building the node lists 
hub_nodes_gen <- data.frame(
  name = unique(c(hub_data_gen$trait1, hub_data_gen$trait2))) %>%
  #Counting how many times IPs share a protein for the hubs
  left_join(hub_data_gen %>% 
  distinct(trait1, n_count) %>% 
  rename(name = trait1, count = n_count), by = "name") %>%
  #Categorising
  mutate(type = ifelse(is.na(count), "Protein", "IP"),
  colour_group = ifelse(type == "Protein", "Protein", paste(count, "Proteins"))) %>%
  #Adding lineage info
  left_join(lineage_info_gen, by = c("name" = "trait1")) %>%
  mutate(shape_map = ifelse(type == "Protein", "Protein", Sub.Lineage))

#Network Graph 
#Combining the edges and nodes
hub_graph_gen <- graph_from_data_frame(hub_data_gen %>% select(trait1, trait2), directed = FALSE, vertices = hub_nodes_gen)

#Assigning colours to them 
#Need automated colour and shapes for IPs but keep protein the same 
hub_nodes_gen$colour_group <- factor(
  hub_nodes_gen$colour_group,levels = c("Protein", 
                                        sort(unique(
                                          hub_nodes_gen$colour_group[
                                            hub_nodes_gen$colour_group != "Protein"]))))

#Unique 
unique_conn_gen <- sort(unique(hub_nodes_gen$colour_group[hub_nodes_gen$colour_group != "Protein"]))

#Final colours for each + black for protein
final_colours_gen <- c("Protein" = "black", setNames(viridis_pal(option = "viridis")(length(unique_conn_gen)), unique_conn_gen))

#Shape for aesthetics
unique_lineages_gen <- sort(unique(hub_nodes_gen$shape_map))

#start at shape 15 square
shape_sequence_gen <- seq(from = 15, length.out = length(unique_lineages_gen)) 

#Lineage values for ggraph input
lineage_vals_gen <- setNames(shape_sequence_gen, unique_lineages_gen)


#Graph for visuals
ggraph(hub_graph_gen, layout = "fr") +
  #Arc Edges for visuals
  geom_edge_arc(alpha = 0.5, color = "grey70", strength = 0.1) +
  #Node
  geom_node_point(aes(color = colour_group, 
                      shape = shape_map), 
                  size = 4,
                  alpha = 0.9) +
  #Label
  geom_node_text(aes(label = name,
                 filter = type == "Protein"),
                 repel = TRUE, 
                 size = 3,
                 fontface ="bold",
                 box.padding = 0.4,
                 point.padding = 0.4)+
  #Colours
  scale_color_manual(values = final_colours_gen,breaks = unique_conn_gen
  ) + 
  #Size based on hub
  scale_shape_manual(values = lineage_vals_gen) +
#Aesthetic theme
  theme_void() +
  theme(
     plot.background = element_rect(fill = "white", color = NA),
     panel.background = element_rect(fill = "white", color = NA),
     plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
     plot.subtitle = element_text(hjust = 0.5, size = 11),
     #space for legend
     plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
     legend.position = "right",
     legend.box = "vertical",
     legend.text = element_text(size = 10),
   legend.title = element_text(size = 10, face = "bold"))+
  labs(title = "Network of Genetically Significant Hubs",
       color = "IP Hub Connectivity",
      shape = "Lineages by Shape") +
  guides(edge_width = "none")

#Saving plot
ggsave(
  filename = "genetic_hits/Gen_Network_Hubs.png", 
  path = output_path,
  width = 12,     
  height = 8
)
```

### 6.7 Static IP-Protein Pairs Network Graph

Filtered for \|rhoG\| at a threshold and IPs connected to 2 + proteins.

```{r}
#Lookup data 
lineage_info_gen <- genetic_0.1_hits %>%
  select(trait1, Sub.Lineage) %>%
  distinct(trait1, .keep_all = TRUE)

#FILTERING
hub_data_gen <- genetic_0.1_hits %>%
  filter(abs(rhoG) >= 0.3) %>% 
  group_by(trait1) %>%
  filter(n() >= 2) %>% 
  mutate(n_count = n()) %>%
  ungroup() %>%
  mutate(
    abs_rhoG = abs(rhoG),
    edge_sign = ifelse(rhoG >= 0, "Positive", "Negative") #For connection 
  )

#Node List
hub_nodes_gen <- data.frame(
  name = unique(c(hub_data_gen$trait1, hub_data_gen$trait2))
) %>%
  #Defining proteins from IPs
  left_join(lineage_info_gen, by = c("name" = "trait1")) %>%
  mutate(
    type = ifelse(name %in% lineage_info_gen$trait1, "IP", "Protein"),
    shape_map = ifelse(type == "Protein", "Protein", Sub.Lineage)
  )


#Edges
edge_data <- hub_data_gen %>% select(trait1, trait2, rhoG, abs_rhoG, edge_sign)

hub_graph_gen <- graph_from_data_frame(
  d = edge_data, 
  directed = FALSE, 
  vertices = hub_nodes_gen
)

#Leiden Clustering
#abs_rhoG as weights so strong negative correlations cluster together
leiden_clusters <- cluster_leiden(
  hub_graph_gen, 
  objective_function = "modularity", 
  weights = E(hub_graph_gen)$abs_rhoG
)

V(hub_graph_gen)$module <- as.character(leiden_clusters$membership)

#Colour - IPs are coloured by modules, proteins remain
V(hub_graph_gen)$colour_group <- ifelse(
  V(hub_graph_gen)$type == "Protein", 
  "Protein", 
  paste("Module", V(hub_graph_gen)$module)
)

graph_nodes <- igraph::as_data_frame(hub_graph_gen, what = "vertices")

#Colours, shapes, lines
unique_modules <- sort(unique(graph_nodes$colour_group[graph_nodes$colour_group != "Protein"]))

final_colours_gen <- c(
  "Protein" = "black", 
  setNames(viridis_pal(option = "viridis")(length(unique_modules)), unique_modules)
)

#Shapes - keep proteins same, IPs shapes for lineage
unique_lineages_gen <- sort(unique(graph_nodes$shape_map[graph_nodes$shape_map != "Protein"]))

shape_sequence_gen <- seq(from = 15, length.out = length(unique_lineages_gen)) 
lineage_vals_gen <- setNames(shape_sequence_gen, unique_lineages_gen)
#Proteins have a small dot
lineage_vals_gen["Protein"] <- 20 

#Lines
edge_line_vals <- c("Positive" = "solid", "Negative" = "dotted")

#Graphing
#Fruchterman-Reingold (FR) layout 
ggraph(hub_graph_gen, layout = "fr", weights = abs_rhoG) +
  #Edges dotted for negative, positive for sold 
  geom_edge_arc(
    aes(edge_linetype = edge_sign), 
    alpha = 0.5, 
    color = "grey60", 
    strength = 0.1
  ) +
  scale_edge_linetype_manual(values = edge_line_vals) +
  
  #Nodes - module as colours, shapes as lineage
  geom_node_point(
    aes(color = colour_group, shape = shape_map), 
    size = 4,
    alpha = 0.9
  ) +
  
  #Only label proteins
  geom_node_text(
    aes(label = name, filter = type == "Protein"),
    repel = TRUE, 
    size = 3,
    fontface = "bold",
    box.padding = 0.4,
    point.padding = 0.4
  ) +
  
  scale_color_manual(values = final_colours_gen, breaks = unique_modules) + 
  scale_shape_manual(values = lineage_vals_gen) +
  
  #Aesthetics
  theme_void() +
  theme(
     plot.background = element_rect(fill = "white", color = NA),
     panel.background = element_rect(fill = "white", color = NA),
     plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
     plot.subtitle = element_text(hjust = 0.5, size = 11),
     plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
     legend.position = "right",
     legend.box = "vertical",
     legend.text = element_text(size = 10),
     legend.title = element_text(size = 10, face = "bold")
  ) +
  labs(
    title = "Network of Genetically Significant Hubs (Leiden Clustered)",   
    subtitle = "Based on Leiden clustering. Filtered for |ρG| >0.3 and IPs associated to more than one protein. Fruchterman-Reingold (FR) layout was used.",
    color = "Leiden Modules",
    shape = "Lineages by Shape",
    edge_linetype = "rhoG Direction"
  ) +
  guides(edge_width = "none")


#Saving plot
ggsave(
  filename = "genetic_hits/Static_Gen_IP_Protein_Network_Hubs.png", 
  path = output_path,
  width = 20,     
  height = 15
)

```

## 7. Environmental Hits Analysis

Investigating environmental_hits. Environmental markers - reactive in response externally. Likely not realted to ageing.

### 7.1 Summary IP to Protein

```{r}
environmental_hits %>%
  summarise(
    total_rows = n(),
    unique_IPs = n_distinct(trait1),
    unique_proteins = n_distinct(trait2)
  )

```

```{r}
environmental_hits %>%
  summarise (
    mean_h2r1 = mean(h2r1),
    sd_h2r1   = sd(h2r1),
    max_h2r1 = max(h2r1),
    mean_h2r2 = mean(h2r2),
    sd_h2r2= sd(h2r2),
    max_h2r2 = max(h2r2),
    mean_rhoG = mean(abs(rhoG)),
    sd_rhoG= sd(abs(rhoG)),
    mean_rhoE = mean(abs(rhoE)),
    sd_rhoE= sd(abs(rhoE))
  )
```

### 7.2 Protein Distribution

```{r}

ip_per_protein_env <- environmental_hits %>%   
  distinct(trait1, trait2) %>% #Should be unique but sanity checks   
  count(trait2, sort = TRUE, name = "n_IPs")  

print(ip_per_protein_env)

```

```{r}
env_top15_prot <- ip_per_protein_env %>%   
  slice_max(n_IPs, n = 15, with_ties = FALSE) %>%   
  ggplot(aes(x = reorder(trait2, n_IPs), y = n_IPs)) +   geom_col(aes(fill = n_IPs), 
                                                                  alpha = 0.8, color = "white", 
                                                                  linewidth = 0.3) +    #Numerical values next to bar 
  geom_text(aes(label = n_IPs), hjust = -0.2, size =3, color = "black", fontface = "bold")+   
  #Colour scheme   
  scale_fill_viridis_c(option = "viridis")+   
  #Axis   
  coord_flip() +   
  scale_y_continuous(limits = c(0, 500),                    expand = expansion(mult = c(0, 0.1)))+   
  #Labelling   
  labs(title = "Top 15 Proteins by Number of IP Connections",        
       x = "Protein",
       y = "Number of IPs ") +   
  theme_minimal() +   
  theme(     
    axis.text.y = element_text(size = 10),          
    axis.title.y = element_text(size = 12,face = "bold"),           axis.text.x = element_text(size = 10),     
    axis.title.x = element_text(size = 12,face = "bold"), 
    plot.title = element_text(size = 14, face = "bold",hjust = 0.5, margin = margin(b = 10)),     
    legend.position = "none",     
    axis.line = element_line(color = "black"),    
    panel.grid.major.y = element_blank() #No vertical lines   
    )  

#Saving plot 
ggsave(   
  filename = "environmental_hits/Env_Top_15_Protein_Distribution.png",
  plot = env_top15_prot,
  path = output_path,   
  width = 10,        
  height = 8 )
```

### 7.3 IP Shared Protein Distribution

```{r}
#Counting how many share proteins 
ip_env_sharing <- environmental_hits %>%   
  group_by(trait1) %>%   
  summarise(protein_count = n()) %>%   
  arrange(desc(protein_count))

#Table 
shared_protein_ip_env <- table(ip_env_sharing$protein_count) 

shared_protein_ip_env 
```

```{r}
#DF conversion of the table 
ip_shared_prot_df_env <- as.data.frame(shared_protein_ip_env)
colnames(ip_shared_prot_df_env ) <- c("Num_Proteins", "Count_IPs")   

ip_shared_prot_plot_env <- ggplot(ip_shared_prot_df_env, 
                                  aes(x = Num_Proteins, y = Count_IPs)) +  
  geom_col(aes(fill = Count_IPs), 
           alpha = 0.8, color = "white",
           linewidth = 0.3) +    
  #Numerical values next to bar   
  geom_text(aes(label = Count_IPs), hjust = -0.2, size =3, color = "black", fontface = "bold")+   
  #Colour scheme   
  scale_fill_viridis_c(option = "viridis")+   
  #Axis   
  coord_flip() +    
  labs(title = "Protein Connectivity to IP Distribution",        
       x = "Number of Proteins Connected to a Single IP", y = "Number of IPs ") +  
  theme_minimal() +   
  theme(     
    axis.text.y = element_text(size = 8),          
    axis.title.y = element_text(size = 10,face = "bold"),     
    axis.text.x = element_text(size = 8),     
    axis.title.x = element_text(size = 10,face = "bold"),     
    plot.title = element_text(size = 12, face = "bold",hjust = 0.5, margin = margin(b = 10)),     
    legend.position = "none",     
    axis.line = element_line(color = "black"),     
    panel.grid.major.y = element_blank() #No vertical lines   
    )  

plot(ip_shared_prot_plot_env) 


#Saving plot 
ggsave(   
  filename = "environmental_hits/Env_IP_Prot_Distribution.png",    
  plot = ip_shared_prot_plot_env,
  path = output_path,   
  width = 10,        
  height = 8 )
```

### 7.4 Lineage Distribution

```{r}
#Lineage count
environmental_hits %>%
  count(Lineage, sort = TRUE)
```

```{r}
#Sup lineage 
environmental_hits  %>%
  group_by(Sup.Lineage) %>%
  summarise(unique_ip_count = n_distinct(trait1)) %>%
  arrange(desc(unique_ip_count))
```

### 7.5 Bidirectionality

#### Overall

```{r}
#overall count 
environmental_hits %>%   
  mutate(Direction = if_else(rhoE < 0, "Negative (-)", "Positive (+)")) %>%   
  count(Direction)
```

Obtaining the mean rhoG for both directions

```{r}
environmental_hits %>%
  mutate(direction = ifelse(rhoE > 0, "Positive", "Negative")) %>%
  group_by(direction) %>%
  summarise(
    count = n(),
    mean_rhoE = mean(rhoE, na.rm = TRUE),
    sd_rhoE = sd(rhoE, na.rm = TRUE)
  )
```

#### Per Lineage

```{r}
#count per lineage 
env_bidirection_counts <- environmental_hits %>%   
  mutate(Direction = if_else(rhoE < 0, "Negative (-)", "Positive (+)")) %>%   
  count(Direction, Sup.Lineage)  

print(env_bidirection_counts)
```

##### Graph

Graphing the bidirectionality by proportion in super lineages

```{r}
env_bidirection<- ggplot(env_bidirection_counts, 
       aes(x = Direction, y = n, fill = Direction)) +   
  geom_col() +   
  geom_text(aes(label = n), vjust = -0.5, size = 3,color = "black", fontface = "bold") +   
  facet_wrap(~Sup.Lineage, scales = "free_y") +   
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +   
  scale_fill_manual(values = c("Positive (+)" = "#21908CFF", "Negative (-)" = "#440154FF")) +   
  theme_bw() +     
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14, margin = margin(b = 10)),  
    legend.position = c(0.85, 0.2), #Positioned in empty space       
    legend.text = element_text(size = 12),       
    legend.title = element_text(size = 12, face = "bold"),       
    legend.key.size = unit(1.2, "cm"),       
    legend.key.width = unit(1, "cm"),       
    strip.background = element_rect(fill = "white"),        
    strip.text = element_text(face = "bold"),       
    axis.title.y = element_text(size = 12,face = "bold"),      
    axis.title.x = element_text(size = 12,face = "bold"),   
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    ) +   
  labs(title = "Directionality by Lineage",        
       y = "Number of Significant Genetic Hits")   


#Saving plot 
ggsave(   
  filename = "environmental_hits/Env_Direction_Lineage.png",
  plot = env_bidirection,
  path = output_path,   
  width = 12,        
  height = 8 )
```

### Paneling

```{r}

env_panels <-  (env_bidirection) /(env_top15_prot + ip_shared_prot_plot_env)+
    plot_layout(heights = c(1, 1))

env_panels <- env_bidirection + env_top15_prot + ip_shared_prot_plot_env +
  
  
   plot_layout(design = spacer) +
  
  plot_annotation(  
    tag_levels = 'A'
  ) &
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.tag.position = c(0.08, 0.99),
    plot.tag = element_text(size = 18, face = "bold"))


ggsave(
  filename = "environmental_hits/Environmental_Panels.png",
  plot = env_panels,
  path = output_path,
  width = 20,
  height = 18
)
```

### 7.6 Network graph with the Hubs

To visualise hubs and IP - Protein connections, can filter how many hubs we want in the network graph.

```{r}
#Lookup data  
lineage_info_env<- environmental_hits %>%   
  select(trait1, Sub.Lineage) %>%   
  distinct(trait1, .keep_all = TRUE)  

#defining hub list  
hub_data_env <- environmental_hits %>%   
  group_by(trait1) %>%   
  filter(n() >= 4) %>% #Select the number of Hubs +    
  mutate(n_count = n()) %>%   
  ungroup()%>%   
  select(trait1, trait2, Sub.Lineage, n_count) 
```

```{r}
#Graphing 
#Building the node lists  
hub_nodes_env <- data.frame(   
  name = unique(c(hub_data_env$trait1, hub_data_env$trait2))) %>%   
  #Counting how many times IPs share a protein for the hubs   
  left_join(hub_data_env %>%    
              distinct(trait1, n_count) %>%    
              rename(name = trait1, count = n_count), by = "name") %>%   
  #Categorising   
  mutate(type = ifelse(is.na(count), "Protein", "IP"),   
         colour_group = ifelse(type == "Protein", "Protein", paste(count, "Proteins"))) %>%   
  #Adding lineage info   
  left_join(lineage_info_env, by = c("name" = "trait1")) %>%   
  mutate(shape_map = ifelse(type == "Protein", "Protein", Sub.Lineage))  

#Number of unique lineages 
unique(hub_nodes_env$shape_map)
```

```{r}
#Network Graph  
#Combining the edges and nodes 
hub_graph_env <- graph_from_data_frame(hub_data_env %>% 
                                         select(trait1, trait2), directed = FALSE, vertices = hub_nodes_env)  

#Assigning colours to them  
#Need automated colour and shapes for IPs but keep protein the same
hub_nodes_env$colour_group <- factor(hub_nodes_env$colour_group,levels = c("Protein", sort(unique(hub_nodes_env$colour_group[hub_nodes_env$colour_group != "Protein"])))) 

#Unique  
unique_conn_env <- sort(unique(hub_nodes_env$colour_group[hub_nodes_env$colour_group != "Protein"])) 

#Final colours for each + black for protein 
final_colours_env <- c("Protein" = "black", setNames(viridis_pal(option = "viridis")(length(unique_conn_env)), unique_conn_env)) 

#Shape for aesthetics 
lineage_vals_env <- c(
  "Protein"                = 17, #Triangle
  "CD8 T cells"            = 15, #Square
  "Monocytes"              = 16, #Circle
  "14-/11c+ (nodim)/1c+"   = 8,  #Star
  "14-/mDC/1c+"            = 25, #Inverted triangle
  "B cells/D"              = 18  #Diamond
)



#Graph for visuals 
ggraph(hub_graph_env, layout = "fr") +   
  #Arc Edges for visuals   
  geom_edge_arc(alpha = 0.5, color = "grey70", strength = 0.1) +   
  #Node   
  geom_node_point(aes(color = colour_group,                        
                      shape = shape_map),                    
                  size = 4,                   
                  alpha = 0.9) +   
  #Label  
  geom_node_text(aes(label = name,                  
                     filter = type == "Protein"),                 
                 repel = TRUE,                   
                 size = 3,                  
                 fontface ="bold",                  
                 box.padding = 0.5,                  
                 point.padding = 0.5)+   
  #Colours   
  scale_color_manual(values = final_colours_env,breaks = unique_conn_env 
  ) + 
  #Size based on hub   
  scale_shape_manual(values = lineage_vals_env) + 
  #Aesthetic theme   
  theme_void() +   
  theme(      
    plot.background = element_rect(fill = "white", color = NA), 
    panel.background = element_rect(fill = "white", color = NA),     plot.title = element_text(hjust = 0.5,face = "bold", size = 14),      
    plot.subtitle = element_text(hjust = 0.5, size = 11),      
    #space for legend      
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),      
    legend.position = "right",     
    legend.box = "vertical",      
    legend.text = element_text(size = 10),   
    legend.title = element_text(size = 10, face = "bold"))+   
  labs(title = "Network of Environmentally Significant Hubs",   
       color = "IP Hub Connectivity",       
       shape = "Lineages by Shape") +   
  guides(edge_width = "none")  

#Saving plot 
ggsave(   
  filename = "environmental_hits/Env_Network_Hub.png",    
  path = output_path,   
  width = 12,        
  height = 8 )
```

### 7.7 Static IP-Protein Network Graph

```{r}

#Filterings thresholds
rho_threshold <- 0.3     
protein_sharing <- 2     

lineage_info_env <- environmental_hits %>%
  select(trait1, Sub.Lineage) %>%
  distinct(trait1, .keep_all = TRUE)

#FILTERING
hub_data_env <- environmental_hits %>%
  filter(abs(rhoE) >= rho_threshold) %>% 
  group_by(trait1) %>%
  filter(n() >= protein_sharing) %>% 
  mutate(n_count = n()) %>%
  ungroup() %>%
  mutate(
    abs_rhoE = abs(rhoE),
    edge_sign = ifelse(rhoE >= 0, "Positive", "Negative") 
  )

#Nodes
hub_nodes_env <- data.frame(
  name = unique(c(hub_data_env$trait1, hub_data_env$trait2))
) %>%
  left_join(lineage_info_env, by = c("name" = "trait1")) %>%
  mutate(
    type = ifelse(name %in% lineage_info_env$trait1, "IP", "Protein"),
    shape_map = ifelse(type == "Protein", "Protein", Sub.Lineage)
  )

#Edges
edge_data_env <- hub_data_env %>% select(trait1, trait2, rhoE, abs_rhoE, edge_sign)

hub_graph_env <- graph_from_data_frame(
  d = edge_data_env, 
  directed = FALSE, 
  vertices = hub_nodes_env
)

#Leiden Clustering
#|rhoE| as weights
leiden_clusters_env <- cluster_leiden(
  hub_graph_env, 
  objective_function = "modularity", 
  weights = E(hub_graph_env)$abs_rhoE
)

V(hub_graph_env)$module <- as.character(leiden_clusters_env$membership)

#IPs are coloured by modules, proteins remain
V(hub_graph_env)$colour_group <- ifelse(
  V(hub_graph_env)$type == "Protein", 
  "Protein", 
  paste("Module", V(hub_graph_env)$module)
)

graph_nodes_env <- igraph::as_data_frame(hub_graph_env, what = "vertices")

#Colours, shapes, lines
unique_modules_env <- sort(unique(graph_nodes_env$colour_group[graph_nodes_env$colour_group != "Protein"]))

final_colours_env <- c(
  "Protein" = "black", 
  setNames(viridis_pal(option = "viridis")(length(unique_modules_env)), unique_modules_env)
)

# Shapes - keep proteins same, IPs shapes for lineage
unique_lineages_env <- sort(unique(graph_nodes_env$shape_map[graph_nodes_env$shape_map != "Protein"]))

shape_sequence_env <- seq(from = 15, length.out = length(unique_lineages_env)) 
lineage_vals_env <- setNames(shape_sequence_env, unique_lineages_env)
#protein shape is dot
lineage_vals_env["Protein"] <- 20 

#Lines
edge_line_vals <- c("Positive" = "solid", "Negative" = "dotted")

#Graphing
# Fruchterman-Reingold (FR) layout 
ggraph(hub_graph_env, layout = "fr", weights = abs_rhoE) +
  #Edges
  geom_edge_arc(
    aes(edge_linetype = edge_sign), 
    alpha = 0.5, 
    color = "grey60", 
    strength = 0.1
  ) +
  scale_edge_linetype_manual(values = edge_line_vals) +
  
  #Nodes - module as colours, shapes as lineage
  geom_node_point(
    aes(color = colour_group, shape = shape_map), 
    size = 4,
    alpha = 0.9
  ) +
  
  #Labelling - only proteins
  geom_node_text(
    aes(label = name, filter = type == "Protein"),
    repel = TRUE, 
    size = 3,
    fontface = "bold",
    box.padding = 0.4,
    point.padding = 0.4
  ) +
  
  scale_color_manual(values = final_colours_env, breaks = unique_modules_env) + 
  scale_shape_manual(values = lineage_vals_env) +
  
  #Aesthetics
  theme_void() +
  theme(
     plot.background = element_rect(fill = "white", color = NA),
     panel.background = element_rect(fill = "white", color = NA),
     plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
     plot.subtitle = element_text(hjust = 0.5, size = 11),
     plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
     legend.position = "right",
     legend.box = "vertical",
     legend.text = element_text(size = 10),
     legend.title = element_text(size = 10, face = "bold")
  ) +
  labs(
    title = "Network of Environmentally Significant Hubs (Leiden Clustered)",   
    subtitle = paste0("Based on Leiden clustering. Filtered for |ρE| >0.3 and IPs associated to more than one protein. Fruchterman-Reingold (FR) layout was used ", protein_sharing, " proteins. FR layout."),
    color = "Leiden Modules",
    shape = "Lineages by Shape",
    edge_linetype = "rhoE Direction"
  ) +
  guides(edge_width = "none")

#Saving
ggsave(
  filename = "environmental_hits/Static_Env_IP_Protein_Network_Hubs.png", 
  path = output_path, 
  width = 20,     
  height = 15
)

```

### 7.8 Heatmap

```{r}
environmental_matrix <- conv_matrix (environmental_hits, "rhoE")

ip_names_env <- colnames(t(environmental_matrix))

anno_env <- environmental_hits %>%
  select(trait1, Lineage) %>%
  distinct() %>%
  filter(trait1 %in% ip_names_env) %>%
  column_to_rownames("trait1")

lineage_env <- anno_env$Lineage

col_env = colorRamp2(c(-0.6, 0, 0.6), 
                         c("#440154FF", "#e8e8e8", "#21908CFF"))

env_heatmap <- Heatmap(t(environmental_matrix), 
          name = "ρE", 
          col = col_env,
          column_split = lineage_env,
          column_title_side = "bottom",
          column_title_gp = gpar(fontsize = 8, fontface = "bold"),
          column_title_rot = 90,
          cluster_rows = TRUE,
          cluster_columns = TRUE,
          show_column_dend = FALSE,
          show_column_names = FALSE,
          show_row_names = TRUE,
          row_names_gp = gpar(fontsize = 8),
          na_col = "grey90",
          border = FALSE)

draw(env_heatmap, column_title = "Environmental Architecture of IP-Protein Pairs")

```

## 8. Both Genetic and Environmentally Correlated

All hits are positive. Only CXCL9 protein has IP associations.

Overall CXCL9 had the phenotype pP correlation.

Deriving from genetic pG and environmental pE but rhoG is stronger/more dominant than rhoE.

### 8.1 Summary IP to Protein

```{r}
both_hits %>%
  summarise(
    total_rows = n(),
    unique_IPs = n_distinct(trait1),
    unique_proteins = n_distinct(trait2)
  )
```

```{r}
#Obtaining average of IPs and Protein as they are aLL CD8 T cell lineage with CXCL9
both_hits %>%
  summarise (
    mean_h2r1 = mean(h2r1),
    sd_h2r1   = sd(h2r1),
    max_h2r1 = max(h2r1),
    mean_h2r2 = mean(h2r2),
    sd_h2r2= sd(h2r2),
    max_h2r2 = max(h2r2),
    mean_rhoG = mean(abs(rhoG)),
    sd_rhoG= sd(abs(rhoG)),
    mean_rhoE = mean(abs(rhoE)),
    sd_rhoE= sd(abs(rhoE))
  )
```

### 8.3 Lineage

```{r}
both_hits %>%   
  group_by(trait2) %>%   
  summarise(     
    n_IPs = n_distinct(trait1),
    associated_lineages = paste(unique(Sub.Lineage), 
                                collapse = ", ")) %>%   
  arrange(desc(n_IPs))
```

### 8.5 Bidirectionality

```{r}
both_hits %>%
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%
  count(Direction)
```

### 8.6 Graph

```{r}
#Plotting with different graph
both_dual <- both_hits %>%
  mutate(IP = reorder(trait1, abs(rhoG))) %>%
  select(IP, Protein = trait2, rhoG, rhoE)

#For the graph mean + sd visuals
stats_rhoG <- both_hits %>%
  summarise(mean = mean(rhoG), sd = sd(rhoG)) %>%
  mutate(min = mean - sd, max = mean + sd)
stats_rhoE <- both_hits %>%
  summarise(mean = mean(rhoE), sd = sd(rhoE)) %>%
  mutate(min = mean - sd, max = mean + sd)

#Dumbell graph to show both rhoG and rhoE
ggplot(both_dual, aes(x = reorder(IP, abs(rhoG)))) +
  
  #Mean +sd
  geom_rect(aes(xmin = -Inf, xmax = Inf, 
                ymin = stats_rhoG$min, ymax = stats_rhoG$max), 
            fill = "#440154FF", alpha = 0.02) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, 
                ymin = stats_rhoE$min, ymax = stats_rhoE$max), 
            fill = "#21908CFF", alpha = 0.02) +
  geom_hline(yintercept = stats_rhoG$mean, color = "#440154FF", linetype = "solid", alpha = 0.4) +
  geom_hline(yintercept = stats_rhoE$mean, color = "#21908CFF", linetype = "solid", alpha = 0.4) +
  #label for mean +sd
  annotate("text", x = 1, y = stats_rhoG$mean, label = "MEAN", 
           angle = 90, size = 3, color = "#440154FF", vjust = -1, hjust = -0.2,fontface = "bold") +
  annotate("text", x = 1, y = stats_rhoG$max+0.01, label = "STDEV", 
           angle = 90, size = 3, color = "#440154FF", vjust = -1, hjust = -0.2,fontface = "bold") +
           
  
  geom_segment(aes(xend = IP, y = rhoG, yend = rhoE), color = "grey85", size = 1) +
  geom_point(aes(y = rhoG, color = "Genetic (ρG)"), size = 5) +
  geom_point(aes(y = rhoE, color = "Environmental (ρE)"), size = 5) +
  scale_y_continuous(expand = c(0, 0), limits = c(0.2, 0.55)) + #x axis
  coord_flip() +
  scale_color_manual(values = c("#21908CFF","#440154FF"))+
  labs(
    title = "Both Genetic and Environmental Correlations",
    x = "CD8 T Cells IPs", y = "Correlation Coefficient (ρ) with CXCL9",
    color = "Correlation Type"
  ) +
  theme_minimal() +
    theme(
    legend.position = "bottom",
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 11),
    plot.title = element_text(face = "bold", hjust = 0.5,margin = margin(b = 20)),
    axis.title = element_text(size = 11, face = "bold"),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 9),
    axis.line = element_line(color = "black"),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank())


#Saving plot
ggsave(
  filename = "both_hits/Both_Gen_Env_Dumbbell.png", 
  path = output_path,
  width = 10,     
  height = 8
)


print(stats_rhoG$mean)
print(stats_rhoG$sd)
print(stats_rhoE$mean)
print(stats_rhoE$sd)

```

CXCL9 more genetically correlated than environmental counterpart though both significant. Heterogeneity of immune cell influence. MEAN and SD purely for trend, no biological context. Effect size are significantly higher in genetic than environmental . They are all from CD8 T cell lineage.

## 9. Genetic -\> Environmental

### 9.1 Summary IP to Protein

Number of IPs and Proteins

```{r}
gen_env_0.1 %>%
  summarise(
    total_rows = n(),
    unique_IPs = n_distinct(trait1),
    unique_proteins = n_distinct(trait2)
  )
```

```{r}
gen_env_0.1 %>%
  summarise (
    mean_h2r1 = mean(h2r1),
    sd_h2r1   = sd(h2r1),
    max_h2r1 = max(h2r1),
    mean_h2r2 = mean(h2r2),
    sd_h2r2= sd(h2r2),
    max_h2r2 = max(h2r2),
    mean_rhoG = mean(abs(rhoG)),
    sd_rhoG= sd(abs(rhoG)),
    mean_rhoE = mean(abs(rhoE)),
    sd_rhoE= sd(abs(rhoE))
    
  )

```

### 9.2 Protein Distribution

```{r}
ip_per_protein_gen_env <- gen_env_0.1%>%
  distinct(trait1, trait2) %>% #Should be unique but sanity checks
  count(trait2, sort = TRUE, name = "n_IPs")

print(ip_per_protein_gen_env)
```

Finding out which IP lineage is paired to TNFB

```{r}
gen_env_0.1 %>%
  filter(trait2 == "TNFB") %>%
  select(trait1, Sub.Lineage, trait2) %>% 
  distinct()
```

### 9.3 IP Shared Protein Distribution

```{r}
#Counting how many share proteins
ip_gen_env_sharing <- gen_env_0.1 %>%
  group_by(trait1) %>%
  summarise(protein_count = n()) %>%
  arrange(desc(protein_count))

#Table
shared_protein_ip_gen_env <- table(ip_gen_env_sharing$protein_count)

shared_protein_ip_gen_env
```

Finding out which IPs share the two proteins.

```{r}
gen_env_0.1 %>%
  group_by(trait1) %>%
  mutate(protein_count = n()) %>%  
  filter(protein_count == 2) %>%
  ungroup()
```

### 9.4 Lineage

```{r}
gen_env_0.1 %>%
  group_by(trait2) %>%
  summarise(
    n_IPs = n_distinct(trait1),                
    associated_lineages = paste(unique(Sub.Lineage), collapse = ", ") 
  ) %>%
  arrange(desc(n_IPs))
```

### 9.5 Bidirectionality

```{r}
gen_env_0.1 %>%
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%
  count(Direction)
```

```{r}
#count per lineage 
gen_env_0.1 %>%   
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)")) %>%    count(Direction, Sup.Lineage, trait2)  

```

### 9.6 Summary Graph

```{r}
#Add in direction
gen_env_0.1_data <- gen_env_0.1 %>%
  mutate(Direction = if_else(rhoG < 0, "Negative (-)", "Positive (+)"))


ggplot(gen_env_0.1_data, aes(x = trait2, fill = Sub.Lineage)) +
  geom_bar( position = "stack", alpha = 0.8) +
  #Panel
  facet_wrap(~ Direction, scales = "free_y") +
  #Label
  geom_text(stat = "count",
            aes(label = after_stat(count)), 
            position = position_stack(vjust = 0.5),
            size = 4,
            color = "black", 
            fontface = "bold") +   

  labs(
    title = "Distribution of Genetically Environmental Associations",
    x = "Proteins",
    y = "Number of IP Associations",
    fill = "Lineage"
  ) +
  theme_bw() +     
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14, margin = margin(b = 10)),       
    legend.position = "bottom" ,
    legend.text = element_text(size = 12),       
    legend.title = element_text(size = 12, face = "bold"),       
    legend.key.size = unit(1.2, "cm"),       
    legend.key.width = unit(1, "cm"),       
    strip.background = element_rect(fill = "white"),        
    strip.text = element_text(face = "bold"),       
    axis.title.y = element_text(size = 12,face = "bold"),      
    axis.title.x = element_text(size = 12,face = "bold"),
      panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
    ) +   
  scale_fill_viridis_d()

#Saving plot
ggsave(
  filename = "gen_env_hits/Gen_Env_Summary_Bar_Graph.png", 
  path = output_path,
  width = 12,     
  height = 8
)
```

### 9.7 Heatmap

```{r}
gen_env_matrix <- conv_matrix (gen_env_0.1, "rhoE")

ip_names_gen_env <- colnames(t(gen_env_matrix))

anno_gen_env<- gen_env_0.1 %>%
  select(trait1, Lineage) %>%
  distinct() %>%
  filter(trait1 %in% ip_names_gen_env) %>%
  column_to_rownames("trait1")

lineage_gen_env <- anno_gen_env$Lineage

col_gen_env = colorRamp2(c(-0.6, 0, 0.6), 
                          c("#440154FF", "#e8e8e8", "#21908CFF"))


gen_env_heatmap<-Heatmap(t(gen_env_matrix),
          name = "rhoE",
          col = col_gen_env,
          column_split = lineage_gen_env,
          column_title_side = "bottom",
          column_title_gp = gpar(fontsize = 8, fontface = "bold"),
          column_title_rot = 90,
          cluster_rows = FALSE, 
          cluster_columns = FALSE,
          show_column_dend = FALSE,
          show_column_names = FALSE,
          show_row_names = TRUE,
          row_names_gp = gpar(fontsize = 8),
          na_col = "grey90",
          border = FALSE
          )

draw(gen_env_heatmap, 
     column_title = "Genetically with Environmental IP-Protein Correlations ")

```

### 9.8 Plot Graph

```{r}
ggplot(gen_env_0.1, aes(x = rhoE, y = rhoG, color = trait2)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = " Genetically Envrionmental Correlations ",
    subtitle = "IP–Protein associations coloured by protein",
    x = "Environmental Correlation (rhoE)",
    y = "Genetic Correlation (rhoG)",
    color = "Protein"
  )
```

## 10. Environmental -\> Genetic

### 10.1 Summary IP to Protein

```{r}
env_gen %>%
  summarise(
    total_rows = n(),
    unique_IPs = n_distinct(trait1),
    unique_proteins = n_distinct(trait2)
  )
```

```{r}
env_gen  %>%
  summarise (
    mean_h2r1 = mean(h2r1),
    sd_h2r1   = sd(h2r1),
    max_h2r1 = max(h2r1),
    mean_h2r2 = mean(h2r2),
    sd_h2r2= sd(h2r2),
    max_h2r2 = max(h2r2),
    mean_rhoG = mean(abs(rhoG)),
    sd_rhoG= sd(abs(rhoG)),
    mean_rhoE = mean(abs(rhoE)),
    sd_rhoE= sd(abs(rhoE))
  )
```

### 10.2 Protein Distribution

```{r}
ip_per_protein_env_gen <- env_gen%>%   
  distinct(trait1, trait2) %>% #Should be unique but sanity checks   
  count(trait2, sort = TRUE, name = "n_IPs")  

print(ip_per_protein_env_gen)
```

### 10.3 IP Shared Protein Distribution

```{r}
#Counting how many share proteins 
ip_env_gen_sharing <- env_gen %>%   
  group_by(trait1) %>%   
  summarise(protein_count = n()) %>%   
  arrange(desc(protein_count))  

#Table 
shared_protein_ip_env_gen<- table(ip_env_gen_sharing$protein_count)

shared_protein_ip_env_gen
```

Finding out which IPs share the two proteins.

```{r}
env_gen%>%   
  group_by(trait1) %>%   
  mutate(protein_count = n()) %>%     
  filter(protein_count == 2) %>%  
  ungroup()
```

All monocytes

### 10.4 Lineage

```{r}
env_gen %>%   
  group_by(trait2) %>%   
  summarise(     
    n_IPs = n_distinct(trait1),
    associated_lineages = paste(unique(Sub.Lineage), 
                                collapse = ", ")) %>%   
  arrange(desc(n_IPs))
```

### 10.5 Bidirectionality

```{r}
env_gen %>%   
  mutate(Direction = if_else(rhoE < 0, "Negative (-)", "Positive (+)")) %>%   
  count(Direction)
```

```{r}
#count per lineage  
env_gen %>%      
  mutate(Direction = if_else(rhoE < 0, "Negative (-)", "Positive (+)")) %>%    
  count(Direction, Sup.Lineage, trait2)   
```

### 10.6 Summary Graph

```{r}
#All positive
env_gen_data <- ggplot(env_gen, aes(x = trait2, fill = Sub.Lineage)) + 
  geom_bar(position = "stack", alpha = 0.8) +
  #Label   
  geom_text(stat = "count",             
            aes(label = after_stat(count)),             
            position = position_stack(vjust = 0.5),             
            size = 4,             
            color = "black",              
            fontface = "bold") +
  labs(     
    title = "Distribution of Environmentally Genetic Associations",     x = "Proteins",     
    y = "Number of IP Associations",     
    fill = "Lineage"
    ) +   
  theme_classic() +        
  theme(     
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14, margin = margin(b = 10)),            
    legend.position = "bottom" ,     
    legend.text = element_text(size = 12),            
    legend.title = element_text(size = 12, face = "bold"),
    legend.key.size = unit(1.2, "cm"),            
    legend.key.width = unit(1, "cm"),            
    strip.background = element_rect(fill = "white"),             
    strip.text = element_text(face = "bold"),            
    axis.title.y = element_text(size = 12,face = "bold"),       
    axis.title.x = element_text(size = 12,face = "bold"),    
      panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
    ) +      
  scale_fill_manual(values = c("#440154FF", "#21908CFF"))

#Saving plot 
ggsave(   
  filename = "env_gen_hits/Env_Gen_Summary_Bar_Graph.png",    
  path = output_path,   
  width = 8,        
  height = 8 )
```

### 10.7 Heatmap

```{r}
env_gen_matrix <- conv_matrix(env_gen, "rhoE")

ip_names_env_gen <- colnames(t(env_gen_matrix))

anno_env_gen <- env_gen %>%
  select(trait1, Lineage) %>%
  distinct() %>%
  filter(trait1 %in% ip_names_env_gen) %>%
  column_to_rownames("trait1")

lineage_env_gen <- anno_env_gen$Lineage

col_env_gen = colorRamp2(c(-0.6, 0, 0.6), 
                          c("#440154FF", "#e8e8e8", "#21908CFF"))

env_gen_heatmap<-Heatmap(t(env_gen_matrix),
          name = "rhoE",
          col = col_env_gen,
          column_split = lineage_env_gen,
          column_title_side = "bottom",
          column_title_gp = gpar(fontsize = 8, fontface = "bold"),
          column_title_rot = 90,
          cluster_rows = FALSE, 
          cluster_columns = FALSE,
          show_column_dend = FALSE,
          show_column_names = FALSE,
          show_row_names = TRUE,
          row_names_gp = gpar(fontsize = 8),
          na_col = "grey90",
          border = FALSE
          )

draw(env_gen_heatmap, 
     column_title = "Environmental then Genetic IP-Protein Correlations ")
```

All positive correlations, few overlaps.

### 10.8 Plot Graph

```{r}
ggplot(env_gen, aes(x = rhoE, y = rhoG, color = trait2)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = " Environmentally Correlated with Genetic Correlations ",
    subtitle = "IP–Protein associations coloured by protein",
    x = "Environmental Correlation (rhoE)",
    y = "Genetic Correlation (rhoG)",
    color = "Protein"
  )
```
