---
title: "protein_protein_analysis"
format: html
editor: visual
---

# Protein Protein Genetic Correlations

With the SOLAR analysis results performed for all proteins protein associations, identify the genetic and environmental correlations. The SOLAR ran for protein-protein associations used the twins to identify whether they share the same proteins

## 1. Set Up

### 1.1 Load Libraries

```{r, message = FALSE}
library(tidyr)
library(dplyr)
library(viridis) #Colour blind friendly palette
#Network graphs
library(igraph)
library(ggraph)
library(tidygraph)
```

### 1.2 Load data

Loading the collated csv from the solar analysis results

```{r}
prot_data <- read.delim("~/Twins_Project/data/protein_protein_solar/prot_summary_heritability_COMPLETE.csv")
```

### 1.3 Save path

Defining an output path where I will save the visualisations generated to.

```{r}
output_path <- "~/Twins_Project/visualisations"
```

## 2. Filtering

Using FDR filtering

```{r}
#Use FDR adjustment
prot_data_fdr <- prot_data %>%
  mutate(
    rhoG_FDR = p.adjust(rhoG_p0, method = "fdr"),
    rhoE_FDR = p.adjust(rhoE_p, method = "fdr")
  )

#Filter FDR <0.05 for genetic hits
prot_gen_0.1_hits <- prot_data_fdr %>%
  filter(h2r1 > 0.1 & h2r2 > 0.1 & rhoG_FDR < 0.05)
```

## 3. Graphing

#### 3.1 Network Graphs

```{r}
#Building the edges and nodes from the genetic protein hits
prot_edges <- prot_gen_0.1_hits %>%
  filter(abs(rhoG) > 0.5) %>%  #filtered 0.5 for visuals
  mutate(weight = abs(rhoG)) %>%
  select(from = trait1, to = trait2, weight, rhoG)

#Matrix to proteins as nodes and correlations as edges
g_prot_s <- graph_from_data_frame(prot_edges, directed = FALSE)

#Clustering Louvain used to find modules
cl_prot_s <- cluster_louvain(g_prot_s, weights = E(g_prot_s)$weight)

#Building the graph 
#Tidygraph
tg_prot <- tbl_graph(
  nodes = data.frame(
    name = V(g_prot_s)$name, 
    module = membership(cl_prot_s)  #from the clustering results
  ),
  edges = prot_edges %>%          
    rename(weight = weight),       
  directed = FALSE
)%>%
  activate(nodes) %>% #for proteins not the edges
  mutate(Centrality = centrality_degree()) #Proteins correlated with lots more other proteins -> number of edges connected to each node
#allows to idnetify 'hub' proteins with lots of connections and 'specialist' proteins connected to another or few


#Set colours
n_mods <- length(unique(V(tg_prot)$module))
viridis_colours <- viridis(n_mods)

#Plotting with ggraph as better visually
#fr fruchterman-reingold -> related proteins closer together
ggraph(tg_prot, layout = "fr") +
  #Nodes and edges
  geom_edge_link(aes(width = weight, edge_colour = rhoG),alpha = 0.8) +
  geom_node_point(aes(color = as.factor(module), size = Centrality), alpha = 1)+
  #Colour
scale_edge_colour_gradient(low = "black", high = "grey80") +
  scale_color_manual(values = viridis_colours, name = "Module")+
  geom_node_text(aes(label = name), repel = TRUE, size = 2.5, 
                 max.overlaps = 12) +
  #Aesthetics, layouts
  theme_void() +
  theme(
   plot.background = element_rect(fill = "white", color = NA),
   panel.background = element_rect(fill = "white", color = NA),
   plot.title = element_text(hjust = 0.5,face = "bold", size = 14),
   plot.subtitle = element_text(hjust = 0.5, size = 11),
   #space for legend
   plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"),
   legend.position = "right",
   legend.box = "vertical",
   legend.text = element_text(size = 6),
   legend.title = element_text(size = 8, face = "bold"))+
  labs(title = "Protein-Protein Genetic Correlation Network",
       subtitle = "Based on SOLAR analysis results",
       edge_color = "rhoG",
       color = "Module") +
  guides(edge_width = "none")

#Saving plot
ggsave(
  filename = "Protein_protein_solar_networkgraph.png", 
  path = output_path,
  width = 10,     
  height = 6
)
```

The lines/edges represent genetic corerelation RhoG (pG) -\> shared genetic influence / relationship between two proteins.

The size of the points represent centrality which is how many connections/neighbours a proteins has -\> how integrated the protein is in the network.

Colours represent modules -\> how groups of proteins are highly correlated with each other.

SCF and CCL20 are the only protein pair to have a negative rhoG however, when searching, they should have a positive correlations so the results seem speculative...

```{r}
#ADDS IN CIRCLES AROUND BUT DONT REALLY NEED AS WE ALREADY HAVE COLOURS IN PLACE FOR THE MODULES
library(ggforce) 


n_mods <- length(unique(V(tg_prot)$module))
viridis_colours <- viridis(n_mods)

ggraph(tg_prot, layout = "fr") +
  
  geom_mark_hull(aes(x = x, y = y,fill = as.factor(module), label = as.factor(module)), 
                 concavity = 5,   #Higher concavity makes tighter hulls
                 radius = unit(0.3, "cm"),  #Smoothing
                 alpha = 0.05,    
                 colour = "grey80", 
                 expand = unit(0.5, "cm"), 
                 label.size = 3) + 
  
  
  geom_edge_link(aes(width = weight, edge_colour = rhoG), alpha = 0.5) +
  geom_node_point(aes(color = as.factor(module), size = Centrality), alpha = 0.8) +
  
  scale_edge_colour_gradient(low = "black", high = "grey90") +
  

  scale_fill_manual(values = viridis_colours, guide = "none") + 
  scale_color_manual(values = viridis_colours, name = "Module") +
  
  geom_node_text(aes(label = name), repel = TRUE, size = 1.8, max.overlaps = 12) +
  theme_void() +
  labs(
    title = "Protein-Protein Genetic Correlation Network",
    subtitle = "Based on SOLAR analysis results (Grouping Hulls Added)",
    edge_colour = "rhoG",
    color = "Module",
    size = "Centrality"
  ) +
  guides(edge_width = "none")
```

If I make this interactive, can adjust the filtering of rhoG \>0.5 but will risk overcrowded visualisations.
